### 2018-02-05 Replication - P values/notes_seminar.tex
Notes on McNemar's/Fisher's
Skeleton of results for McNemar's Test with initial thoughts on Fisher's Method

\documentclass{article}

%\usepackage[margin=1.0in]{geometry}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage[hidelinks]{hyperref}

	\addtolength{\oddsidemargin}{-.875in}
	\addtolength{\evensidemargin}{-.875in}
	\addtolength{\textwidth}{1.75in}

	\addtolength{\topmargin}{-.875in}
	\addtolength{\textheight}{1.75in}

%\def\labelitemi{--}
%\doublespacing

\title{\textsc{Properties of Methods for Assessing Replication}}
\author{Hedges Working Group}

\begin{document}

\maketitle

Let there be $k$ pairs of studies each containing an original study and a replication study. Each pair of studies can be characterized by a vector of true effect parameters $(\theta_{1,i}, \theta_{2,i})$ that are unobserved and have zero sampling variance. Each pair of studies is further associate with estimates of these true effects: $(T_{1,i}, T_{2,i})$. These estimates have sampling variances $V_{1,i}$ and $V_{2,i}$ which are assumed to be known. Let $T_{j,i} \sim N(\theta_{j,i}, V_{j,i})$ for $j = 1,2$ and $i=1,2,\ldots, k$. Further, we assume that $T_{j,i}$ are independent.
\\

For the purposes of exploring the proposed methods of assessing replication in the Open Science Collaboration article (2015), we assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of $0.05$. Thus, we add another dimension by which we can characterize the study pairs: their p-values. Let $p_{1,i}$ denote the p-value of the \textit{i}\textsuperscript{th} original study and $p_{2,i}$ denote the p-value of the \text{i}\textsuperscript{th} replication study.

\section{Assessing replication via significance and p-values}

The authors propose four different tests for assessing replication based on significance and p-values. These methods include Fisher's method, McNemar's test, the Wilcoxon signed-rank test, and the \textit{t} test for dependent samples. In what follows, we derive properties of each of these tests as they pertain to a particular framework for assessing replication.

The authors assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of $0.05$. For some of the proposed methods, the resulting p-values are classified as either significant ($p_{j,i} \leq 0.05$) or non-significant ($p_{j,i} > 0.05$).

\subsection{McNemar's test}

McNemar's test tests the hypothesis that the proportion of statistically significant results among original studies is equal to the proportion of statistically significant results among the replication studies. Applying this test requires categorizing each $p_{1,i}$ and $p_{2,i}$ as either significant (1) or non-significant (0). That is, each pair of studies is associated with a vector of p-values $(p_{1,i}, p_{2,i})$ that can take one of the following pairs of values: $(0,0), (0,1), (1,0), (1,1)$. This information is generally summarized in a $2\times 2$ contingency table containing counts of occurrences of these four pair values in the data.

\begin{table}[h!]
\centering
\begin{tabular}{l c c }
& Rep Sig & Rep Non-Sig \\
Orig Sig & \# (1,1) & \#(1, 0) \\
Orig Non-Sig & \# (0, 1) & \# (0,0) \\
\end{tabular}
\end{table}

The null hypothesis for McNemar's test is $H_0: P((1,1) \cap (1,0)) = P((1,1) \cap (0,1))$ which can conveniently be rewritten as $H_0: P(p_{1,i} \leq 0.05) = P(p_{2,i} \leq 0.05) $ for any study $i$, randomly drawn from the $k$ studies. Note that this null hypothesis can be translated or mapped onto our replication framework which does not include p-values. That is, by noting that $p_{j,i} \leq 0.05 \equiv \frac{|T_{j,i}|}{\sqrt{V_{j,i}}} \geq 1.96$ we can rewrite the null hypothesis as follows:

\begin{align}
\begin{split}
H_0:  P \Big(\frac{| \tilde{T}_{1,i} |}{\sqrt{\tilde{V}_{1,i}}} \geq 1.96 \Big) &= P \Big(\frac{| \tilde{T}_{2,i} | }{\sqrt{\tilde{V}_{2,i}}} \geq 1.96 \Big) \\
\equiv \\
H_0 : \sum\limits_{i=1}^k P \Big ( \frac{T_{1i}}{V_{1i}} \geq 1.96 | i \Big ) P(i) &= \sum\limits_{i=1}^k P \Big ( \frac{T_{2i}}{V_{2i}} \geq 1.96 | i \Big ) P(i)
\end{split} 
\end{align}

Thus, we can reformulate the null hypothesis for this particular test as a comparison of the average power of the original studies to the average power of the replicate studies.

\section{Fisher's Method}

The authors also propose assessing replication by applying Fisher's method on only the ``nonsignificant \textit{P} values of the replication studies.'' Using our notation, the hypotheses for this particular test can be defined as follows:

\begin{align}
\begin{split}
H_0&: \theta_{2,1} = \theta_{2,2} = \cdots = \theta_{2,k} \\
H_1&: \text{at least one of } \theta_{2,i} \text{ is not } = 0 \\
\end{split}
\end{align}

When the null hypotheses are true, the test statistic $-2\sum\limits_{i=1}^k ln(p_{2,i})$ follows a $\chi^2$ distribution with $2k$ degrees of freedom. This result depends on the fact that $p_{2,i} \sim U(0,1) \forall i$.
\\

Given the constraints presented by the authors (i.e. including only those p-values that are $> 0.05$), the proposed test statistic is not appropriate. In essence, $-ln(p_{2,i}) \sim TEXP(1,3)$ which cannot possibly have a chi-squared distribution once scaled since the support for this statistic has an upper bound. That being said, we can derive a test statistic that \textit{does} follow a chi-squared distribution. If we let $p_{2,i} \sim U(0.05,1)$, then $- ln\Big ( \frac{X-0.05}{0.95} \Big ) \sim Exp(1)$. It follows that $-2 \sum\limits_{i=1}^k ln \Big ( \frac{p_{2,i} - 0.05}{0.95} \Big ) \sim \chi^2_{2k}$ where $k$ is the number of tests being combined. The method as proposed by the author's will always underestimate the size of the test statistic, making it more likely that using this method, you will fail to reject the null hypothesis than when using the corrected test statistic.

\end{document}

### 2018-02-12 significance/significance.Rmd
added files and filled out readme

---
title: "Metrics for Assessing Replication: Statistical Significance"
author: "STAT 461"
date: "Winter 2018"
output: pdf_document
---

The Open Science Collaboration (2015), when evaluating if findings have been replicated argue, "There is no single standard for evaluating replication success." 
They analyzed replicate pairs in a variety of ways.

# Correspondence of Significance

\begin{quote}
Assuming a two-tailed test and significance or a level of 0.05, all test results of original and replication studies were classified as statistically significant ($p \leq 0.05$) and nonsignificant ($p > 0.05$). However, original studies that interpreted nonsignificant $p$ values as significant were coded as significant (four cases, all with $p$ values $< 0.06$).
\end{quote}

The quote above suggests one metric of assessing replication is to determine if findings are statistically significant or not. 
The OSC is not alone in considering this metric (see Steiner and Wong, 2016; Valentine, 2012)
By this standard, the OSC concluded that over 60% of their attempted replications failed.

However, it is not clear how exactly to interpret these results. 
For one, the difference in between statistically significant and nonsignificant findings may not reflect meaningful scientific differences in effects.
As well, the properties of this as an inference procedure---e.g., error rates---are necessary to understand the OSC findings.
How often, for instance, might this analysis construe negligible scientific differences between studies as nonreplication?

Using a statistical framework motivated by meta-analysis, this vignette attemts to clarify subjective ideas of replication as implied by comparing statistical significance.
It describes the types of definitions of replication such analyses might correspond to, and delineates some of the statistical properties in those scenarios.

# Statistical Model and Analysis

We can represent the true underlying scientific effect parameter in study $i$ as $\theta_i$, which we refer to as the study's results.
This is the value that would be observed if study $i$ had perfect precision (zero measurement and sampling variance).
We do not actually observe $\theta_i$, but instead observe an estimate $T_i$ that has some variance $v_i$---due, for example, to sampling.
Four simplifying assumptions used in this vignette are that $T_i$ are independent, unbiased, normally distributed, with known sampling variance $v_i$: 
\[
T_i \stackrel{indep}{\sim} N(\theta_i, v_i)
\]
In general, we may let $i$ range from 1 (the original study) up to an arbitrary number $k > 1$ (to reflect $k-1$ replication attempts), but to evaluate the analysis by the OSC, we set $k=2$. 
Thus $i=1$ corresponds to the original study and $i=2$ refers to the (single) replicate study.

We would argue that replication can be defined in terms of underlying effect parameters $\theta_i$. 
They may vary due to differences in sample compositions---studies $i$ and $j$ sample from different populations---or experimental contexts---minor variations in the way studies $i$ and $j$ carry out the experiment.
However, if both are ignoreable (or nearly so), then the effect parameters ought to be (nearly) the same.
Conversely, estimates $T_i$ may vary due to random chance; that is, even if studies share the same underlying effect $\theta$, estimates will differ due to variation from sampling. 
Thus, as a guiding principle, we would argue that coherent scientific and statistical definitions of replication should involve similarty between the $\theta_i$.

Comparison of statistical significance involves computing $p$-values for the replicates and determining if both are either nonsignificant $p > 0.05$ or significant $p \leq 0.05$.
As stated above, this is the totality of this analysis, and we can summarize this metric as 
\begin{equation}
\mathbf{1}\{p_1 \leq 0.05 \wedge p_2 \leq 0.05\} + 
\mathbf{1}\{p_1 > 0.05 \wedge p_2 > 0.05\}
\label{eq:metsig}
\end{equation}
Under the model (and simplifying assumptions), we can rewrite this as 
\begin{equation}
\mathbf{1}\left\{
            \frac{\vert T_1 \vert}{\sqrt{v_1}} \geq 1.96 \wedge 
            \frac{\vert T_2 \vert}{\sqrt{v_2}} \geq 1.96
          \right\} + 
\mathbf{1}\left\{
            \frac{\vert T_1 \vert}{\sqrt{v_1}} < 1.96 \wedge 
            \frac{\vert T_2 \vert}{\sqrt{v_2}} < 1.96
          \right\}
\label{eq:metsigT}
\end{equation}

A type I error occurs when one study is significant, and the other is not (assuming that the studies replicate). 
Then, the probability of a false positive is given by
\begin{align*}
P\left[\text{Type I Error}\right] 
  &= P\left[
        \frac{\vert T_1 \vert}{\sqrt{v_1}} \geq 1.96 \wedge 
        \frac{\vert T_2 \vert}{\sqrt{v_2}} < 1.96
     \right] + 
     P\left[
        \frac{\vert T_1 \vert}{\sqrt{v_1}} > 1.96 \wedge 
        \frac{\vert T_2 \vert}{\sqrt{v_2}} \geq 1.96
     \right] \\
  &= P\left[
      \frac{\vert T_1 \vert}{\sqrt{v_1}} \geq 1.96
     \right] 
    P\left[
      \frac{\vert T_2 \vert}{\sqrt{v_2}} < 1.96
     \right] + 
    P\left[
      \frac{\vert T_1 \vert}{\sqrt{v_1}} > 1.96
     \right]
    P\left[
      \frac{\vert T_2 \vert}{\sqrt{v_2}} \geq 1.96
     \right] \\
  &= \left[
        \Phi\left(-1.96 - \frac{\theta_1}{\sqrt{v_1}}\right) + 
        1 - \Phi\left(1.96 - \frac{\theta_1}{\sqrt{v_1}}\right)
     \right]
     \left[
        \Phi\left(1.96 - \frac{\theta_2}{\sqrt{v_2}}\right) - 
        \Phi\left(-1.96 - \frac{\theta_2}{\sqrt{v_2}}\right)
     \right]
     + \\&\quad\quad
     \left[
        \Phi\left(1.96 - \frac{\theta_1}{\sqrt{v_1}}\right) - 
        \Phi\left(-1.96 - \frac{\theta_1}{\sqrt{v_1}}\right)
     \right]
     \left[
        \Phi\left(-1.96 - \frac{\theta_2}{\sqrt{v_2}}\right) + 
        1 - \Phi\left(1.96 - \frac{\theta_2}{\sqrt{v_2}}\right)
     \right]
\end{align*}
Note that when both studies replicate exactly, when $\theta_1 = \theta_2 = 0$, so that both studies have null effects, then the expression above reduces to
\[
\left[
  \Phi\left(-c_{\alpha}\right) + 
  1 - \Phi\left(c_{\alpha}\right)
\right]
\left[
  \Phi\left(c_{\alpha}\right) - 
  \Phi\left(-c_{\alpha}\right)
\right] =
2 (\alpha - \alpha^2)
\]
Thus, if both significance tests are at the $\alpha = 0.05$ level, then the type I error rate in this case is 0.095.

In general, if the studies replicate exactly, so that $\theta_1 = \theta_2 = \theta$ and the sampling variances are identical so that $v_1 = v_2 = v$, then the expression can be written as 
\[
2[\eta - \eta^2]
\]

```{r, echo=F, message=F, warning=F}
library(ggplot2)
typeIerror = function(theta1, theta2, v1, v2, alpha=0.05, side='both'){
  if(side == 'both'){
    calpha = qnorm(1 - alpha/2)
  } else if(side == 'upper'){
    calpha = qnorm(1 - alpha)
  } else if (side == 'lower'){
    calpha = qnorm(alpha)
  } else {
    print("Must specify side as 'both', 'upper', or 'lower'. 
            Using default 'both' for 2-sided test.")
    calpha = qnorm(1 - alpha/2)
  }

  return(
    (pnorm(-calpha - theta1/sqrt(v1)) + 1 - pnorm(calpha - theta1/sqrt(v1))) *
    (pnorm(calpha - theta2/sqrt(v2)) + pnorm(-calpha - theta2/sqrt(v2))) +
      (pnorm(calpha - theta1/sqrt(v1)) + pnorm(-calpha - theta1/sqrt(v1))) *
      (pnorm(-calpha - theta2/sqrt(v2)) + 1 - pnorm(calpha - theta2/sqrt(v2)))
  )
}
tmp = setNames(data.frame(expand.grid(10:100, 10:100)), c('n1', 'n2'))
tmp$err0 = sapply(1:nrow(tmp), 
                  FUN=function(i) 
                    typeIerror(theta1 = 0, 
                               theta2 = 0, 
                               v1 = 2/tmp$n1[i], 
                               v2 = 2/tmp$n2[i]))
tmp$err2 = sapply(1:nrow(tmp), 
                  FUN=function(i) 
                    typeIerror(theta1 = 0.2, 
                               theta2 = 0.2, 
                               v1 = 2/tmp$n1[i], 
                               v2 = 2/tmp$n2[i]))
tmp$err5 = sapply(1:nrow(tmp), 
                  FUN=function(i) 
                    typeIerror(theta1 = 0.5, 
                               theta2 = 0.5, 
                               v1 = 2/tmp$n1[i], 
                               v2 = 2/tmp$n2[i]))
tmp$err8 = sapply(1:nrow(tmp), 
                  FUN=function(i) 
                    typeIerror(theta1 = 0.8, 
                               theta2 = 0.8, 
                               v1 = 2/tmp$n1[i], 
                               v2 = 2/tmp$n2[i]))

ggplot(tmp) + geom_point(aes(n1, n2, color=err2)) + 
  scale_color_gradient(low='white', high='red')

ggplot(tmp) + geom_point(aes(n1, n2, color=err5)) + 
  scale_color_gradient(low='white', high='red')

ggplot(tmp) + geom_point(aes(n1, n2, color=err8)) + 
  scale_color_gradient(low='white', high='red')

tmp2 = setNames(data.frame(expand.grid(10:100, 0.01*0:100)), c('n', 'theta'))
tmp2$err = sapply(1:nrow(tmp2), 
                  FUN=function(i) 
                    typeIerror(theta1 = tmp2$theta[i], 
                               theta2 = tmp2$theta[i], 
                               v1 = 2/tmp2$n[i], 
                               v2 = 2/tmp2$n[i]))
ggplot(tmp2) + geom_point(aes(n, theta, color=err)) +
  scale_color_gradient(low='white', high='red')
```

However, others have noted that statistical significance may not correspond to replication if the significant estimates do not share the same sign. 
That is, if study 1 has a positive significant estimate and study 2 has a negative and significant estimate, one likely would not conclude they replicate.
Thus, an alternative to (\ref{eq:metsig}) could be
\begin{equation}
\mathbf{1}\{p_1 \leq 0.05 \wedge p_2 \leq 0.05\}\mathbf{1}\{\text{sign}(T_1) = \text{sign}(T_2)\} + 
\mathbf{1}\{p_1 > 0.05 \wedge p_2 > 0.05\}
\label{eq:metsigsign}
\end{equation}
These are related analyses that have slightly different properties, which we illustrate below.

### 2018-03-05 notation.Rmd
added intro for significance metric and a note about notation

---
title: "Notation"
output: pdf_document
---

# Statistical Model for Replication Studies

Meta-analysis represents the "results" of an experiment by an effect parameter, which we denote as $\theta$.
This "true" effect reflects what the researcher would observe if they has no uncertainty in their estimate; for instance, if they had an infinite sample size.
Effect parameters are typically on a standard scale, such as a standardized mean difference, Fisher-transformed correlation coefficient, or log-odds ratio.

In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
For the RPP and RPE programs, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$.
These estimates have variances $v_1$ and $v_2$.
The estimates and their variances are what get reported in typical statistical analyses of experiments.

Two common assumptions are that the estimates are normally distributed with known variances $v_i$; that is:
\[
T_i \stackrel{indep}{\sim} N(\theta_i, v_i), \quad i = 1, \ldots, k
\]
This assumption is almost exactly true for effect sizes such as Fisher's $z$-transformed correlation coefficient and standardized mean differences, and for others remains a useful large-sample approximation.

For the empirical evaluations in question, we can define replication as the similarity of underlying effect parameters.
In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication.
However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
Thus, a more practical definition would say studies replicate if $\theta_1 \approx \theta_2$.
For further discussion on this distinction see Hedges and Schauer (2018).

### 2018-04-10 Ttest-pvals/rep.R
T test and pvalues folder added

@@ -1,0 +1,86 @@
+# thetas are different
+## theta1=0.8 theta2=-0.8 with same n
+
+# N is the number of times repeated
+N<-10000
+# both n1 and n2 have the same sample size of 100 
+# since their estimates are the same size but in opposite directions
+
+n1<-100
+n2<-100
+ts<-c()
+for (k in 1:N){
+  # creates two vectors of estimates from original to replicate
+  # pulling 100 random draws still not sure how many to pull?????
+  t1<- rnorm(100, 0.8,sqrt(4/n1)) #not sure about the variance ?????
+  t2<-rnorm(100,-0.8, sqrt(4/n2))
+      # create the z score and then pvalues for each of the estimates
+      # still questioning the variance of this? 
+      z1<-t1/sqrt(4/n1)
+      p1<-2*pnorm(-abs(z1))
+      z2<-t2/sqrt(4/n2)
+      p2<-2*pnorm(-abs(z2))
+  # take the difference of the two groups of pvalues
+  diff<-p1-p2
+  # get the t stat for the pair t test
+  t<-mean(diff)/(sd(diff)/10)
+  # store in vector
+  ts[k]<-t
+}
+
+hist(ts)
+# 2.64 is what was written down in notes but is it 1.98???
+sum(abs(ts) > 1.98) # number of t stats that are over the critical value
+sum(abs(ts) > 1.98)/N # the probability that the test would conclude that the means are different
+
+
+
+## B where is theta1 is 0.8 and theta2 is 0.2
+## 16n1=n2 since they have different effect sizes small and large
+n1<- 52
+n2<-2*393
+ts<-c()
+for (k in 1:N){
+  t1<- rnorm(100, 0.8,sqrt(4/n1))
+  t2<-rnorm(100,0.2, sqrt(4/n2))
+ 
+  z1<-t1/sqrt(4/n1)
+  p1<-2*pnorm(-abs(z1))
+  z2<-t2/sqrt(4/n2)
+  p2<-2*pnorm(-abs(z2))
+  
+  diff<-p1-p2
+  t<-mean(diff)/(sd(diff)/10)
+  ts[k]<-t
+}
+max(ts)
+sum(ts > 1.98)
+sum(ts > 1.98)/N
+hist(ts)
+
+
+
[…]

### 2018-04-11 src/metrics_funs.R
Added some code for simulations.

@@ -1,0 +1,83 @@
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+### REPLICATION METRICS COMMON FUNCTIONS LIBRARY
+###
+### This script brings together common functions used in assessing the 
+### properties of replication metrics used by the RPP and RPE.
+###
+### It includes some functions for running simulations, computing p-values, etc.
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+
+##----------------------------------------##
+## Common Libraries
+##----------------------------------------##
+library(dplyr); library(distr); library(ggplot2)
+
+
+##----------------------------------------##
+## Simulating k studies
+##----------------------------------------##
+simulate_studies <- function(theta, v, k=NULL){
+  #############################################
+  # TAKES: theta; a vector of means
+  #         v; a vector of sampling variances
+  #         k; an integer for the # of studies (optional)
+  #
+  # RETURNS: data frame of (T, v, theta, v, p-value, power) 
+  #          drawn from normal dist.
+  #############################################
+  
+  # Resolve arguments
+  if(is.null(k)){
+    
+    k = length(theta)
+  
+  } else {
+    
+    if(length(theta) == 1){
+      theta = rep(theta, k)
+    }
+    
+    if(length(v) == 1){
+      v = rep(v, k)
+    }
+  
+  }
+  
+  # Check input
+  if(length(theta) > 1 & length(theta) != k){ # Invalid 
+    
+    print("Error: theta does not have length k.")
+    return(NULL)
+
+  } else if(length(v) > 1 & length(v) != k){ # Invalid
+    
+    print("Error: v does not have length k.")
+    return(NULL)
+    
+  } else if((length(theta) > 1 & length(v) > 1) & length(theta) != length(v)){ # Invalid
+    
+    print("Error: v and theta have different lengths.")
+    return(NULL)
+    
+  } else { # Valid: run simluation and return values
+    
+    ###-----Run simluation, compute properties, and return it as a DF.
+    return(
+      data.frame(
+              t = rnorm(k, theta, sqrt(v)), # draws (T)
+              v = v, # variances
+              theta = theta, # means
[…]

### 2018-04-11 significance/significance.Rmd

@@
 ---
-title: "Metrics for Assessing Replication: Statistical Significance"
-author: "STAT 461"
-date: "Winter 2018"
+title: "Correspondence in Significance"
 output: pdf_document
@@
 
-The Open Science Collaboration (2015), when evaluating if findings have been replicated argue, "There is no single standard for evaluating replication success." 
-They analyzed replicate pairs in a variety of ways.
+Programs of research that examine the replicability of research findings have become more common in psychology. 
+However, there is some ambiguity over how to assess the findings of such programs; that is, how do we know if a finding has been replicated?
+The Open Science Collaboration (2015) argue, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.
+Subsequent programs have used some of these same analysis methods (e.g., Camerer, et al., 2016; Schweinsberg, et al., 2016).
 
-# Correspondence of Significance
+It is not entirely clear what these methods say about replication.
+For one, an actual definition of replication remains somewhat elusive---or flexible---across these methods. 
+Saying that studies get "the same result" is somewhat subjective, and the OSC attempted to operationalize this by describing several different scenarios of studies getting "the same result".
+Moreover, interpreting the results of the OSC analyses also requires knowledge of the properties of the methods they (and others) use.
+For instance, how sensitive are these analyses? 
+The irony is that answering this second question, in some ways, requires an answer to the former: how do we define replication?
 
-\begin{quote}
-Assuming a two-tailed test and significance or a level of 0.05, all test results of original and replication studies were classified as statistically significant ($p \leq 0.05$) and nonsignificant ($p > 0.05$).
-However, original studies that interpreted nonsignificant $p$ values as significant were coded as significant (four cases, all with $p$ values $< 0.06$).
-\end{quote}
+In this paper, we clarify subjective notions about replication using a meta-analytic framework. 
+Given this definition, we then evaluate various analysis methods to determine their properties.
 
-The quote above suggests one metric of assessing replication is to determine if findings are statistically significant or not. 
-The OSC is not alone in considering this metric (see Steiner and Wong, 2016; Valentine, 2012)
-By this standard, the OSC concluded that over 60% of their attempted replications failed.
+# Defining Successful Replication: Meta-analytic Perspectives
 
-However, it is not clear how exactly to interpret these results. 
-For one, the difference in between statistically significant and nonsignificant findings may not reflect meaningful scientific differences in effects.
-As well, the properties of this as an inference procedure---e.g., error rates---are necessary to understand the OSC findings.
-How often, for instance, might this analysis construe negligible scientific differences between studies as nonreplication?
+Meta-analysis represents the "results" of an experiment by an effect parameter, denoted $\theta$.
+This "true" effect reflects what the researcher would observe if they had an infinite sample size.
+Effect parameters are typically on a standard scale, such as a standardized mean difference, Fisher-transformed correlation coefficient, or log-odds ratio.
 
-Using a statistical framework motivated by meta-analysis, this vignette attemts to clarify subjective ideas of replication as implied by comparing statistical significance.
-It describes the types of definitions of replication such analyses might correspond to, and delineates some of the statistical properties in those scenarios.
+In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
+For the RPP and RPE programs, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
+However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
+The estimates and their variances are what get reported in typical statistical analyses of experiments.
 
-# Statistical Model and Analysis
+We argue that one useful definition of replication involves the similarity not of estimates $T_i$, but of actual true effects $\theta_i$.
+The $T_i$ might differ because of both differences in the underlying true effects, but also due to random chance.
+The $\theta_i$ might differ, however, because studies sample participants from different populations, or if there are differences in experimental contexts and conditions between experiments.
+Thus, it may make sense to define replication in terms of differences in experiments rather than statistical noise.
+Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2017).
 
-We can represent the true underlying scientific effect parameter in study $i$ as $\theta_i$, which we refer to as the study's results.
-This is the value that would be observed if study $i$ had perfect precision (zero measurement and sampling variance).
-We do not actually observe $\theta_i$, but instead observe an estimate $T_i$ that has some variance $v_i$---due, for example, to sampling.
-Four simplifying assumptions used in this vignette are that $T_i$ are independent, unbiased, normally distributed, with known sampling variance $v_i$: 
-\[
-T_i \stackrel{indep}{\sim} N(\theta_i, v_i)
-\]
-In general, we may let $i$ range from 1 (the original study) up to an arbitrary number $k > 1$ (to reflect $k-1$ replication attempts), but to evaluate the analysis by the OSC, we set $k=2$. 
-Thus $i=1$ corresponds to the original study and $i=2$ refers to the (single) replicate study.
+In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication.
+However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
+Thus, a more practical definition would say studies replicate if $\theta_1 \approx \theta_2$.
+For further discussion on this distinction see Hedges and Schauer (2018).
 
-We would argue that replication can be defined in terms of underlying effect parameters $\theta_i$. 
-They may vary due to differences in sample compositions---studies $i$ and $j$ sample from different populations---or experimental contexts---minor variations in the way studies $i$ and $j$ carry out the experiment.
-However, if both are ignoreable (or nearly so), then the effect parameters ought to be (nearly) the same.
-Conversely, estimates $T_i$ may vary due to random chance; that is, even if studies share the same underlying effect $\theta$, estimates will differ due to variation from sampling. 
-Thus, as a guiding principle, we would argue that coherent scientific and statistical definitions of replication should involve similarty between the $\theta_i$.
+The key distinction here, is that the definition of replication involves the parameters $\theta_i$. 
+Analyses of replication involve the estimates $T_i$ and their variances $v_i$, which we use make inferences about how similar the $\theta_i$ are.
+This paradigm allows us to evaluate properties of analysis methods.
 
-Comparison of statistical significance involves computing $p$-values for the replicates and determining if both are either nonsignificant $p > 0.05$ or significant $p \leq 0.05$.
-As stated above, this is the totality of this analysis, and we can summarize this metric as 
-\begin{equation}
-\mathbf{1}\{p_1 \leq 0.05 \wedge p_2 \leq 0.05\} + 
-\mathbf{1}\{p_1 > 0.05 \wedge p_2 > 0.05\}
-\label{eq:metsig}
-\end{equation}
-Under the model (and simplifying assumptions), we can rewrite this as 
-\begin{equation}
-\mathbf{1}\left\{
-            \frac{\vert T_1 \vert}{\sqrt{v_1}} \geq 1.96 \wedge 
-            \frac{\vert T_2 \vert}{\sqrt{v_2}} \geq 1.96
-          \right\} + 
-\mathbf{1}\left\{
-            \frac{\vert T_1 \vert}{\sqrt{v_1}} < 1.96 \wedge 
-            \frac{\vert T_2 \vert}{\sqrt{v_2}} < 1.96
-          \right\}
-\label{eq:metsigT}
-\end{equation}
-
-A type I error occurs when one study is significant, and the other is not (assuming that the studies replicate). 
-Then, the probability of a false positive is given by
-\begin{align*}
-P\left[\text{Type I Error}\right] 
-  &= P\left[
-        \frac{\vert T_1 \vert}{\sqrt{v_1}} \geq 1.96 \wedge 
-        \frac{\vert T_2 \vert}{\sqrt{v_2}} < 1.96
-     \right] + 
-     P\left[
-        \frac{\vert T_1 \vert}{\sqrt{v_1}} > 1.96 \wedge 
-        \frac{\vert T_2 \vert}{\sqrt{v_2}} \geq 1.96
-     \right] \\
-  &= P\left[
-      \frac{\vert T_1 \vert}{\sqrt{v_1}} \geq 1.96
-     \right] 
-    P\left[
-      \frac{\vert T_2 \vert}{\sqrt{v_2}} < 1.96
-     \right] + 
-    P\left[
-      \frac{\vert T_1 \vert}{\sqrt{v_1}} > 1.96
-     \right]
-    P\left[
-      \frac{\vert T_2 \vert}{\sqrt{v_2}} \geq 1.96
-     \right] \\
-  &= \left[
-        \Phi\left(-1.96 - \frac{\theta_1}{\sqrt{v_1}}\right) + 
-        1 - \Phi\left(1.96 - \frac{\theta_1}{\sqrt{v_1}}\right)
-     \right]
-     \left[
-        \Phi\left(1.96 - \frac{\theta_2}{\sqrt{v_2}}\right) - 
-        \Phi\left(-1.96 - \frac{\theta_2}{\sqrt{v_2}}\right)
-     \right]
-     + \\&\quad\quad
-     \left[
-        \Phi\left(1.96 - \frac{\theta_1}{\sqrt{v_1}}\right) - 
-        \Phi\left(-1.96 - \frac{\theta_1}{\sqrt{v_1}}\right)
-     \right]
-     \left[
-        \Phi\left(-1.96 - \frac{\theta_2}{\sqrt{v_2}}\right) + 
-        1 - \Phi\left(1.96 - \frac{\theta_2}{\sqrt{v_2}}\right)
-     \right]
-\end{align*}
-Note that when both studies replicate exactly, when $\theta_1 = \theta_2 = 0$, so that both studies have null effects, then the expression above reduces to
-\[
-\left[
-  \Phi\left(-c_{\alpha}\right) + 
-  1 - \Phi\left(c_{\alpha}\right)
-\right]
-\left[
-  \Phi\left(c_{\alpha}\right) - 
-  \Phi\left(-c_{\alpha}\right)
-\right] =
-2 (\alpha - \alpha^2)
-\]
-Thus, if both significance tests are at the $\alpha = 0.05$ level, then the type I error rate in this case is 0.095.
-
-In general, if the studies replicate exactly, so that $\theta_1 = \theta_2 = \theta$ and the sampling variances are identical so that $v_1 = v_2 = v$, then the expression can be written as 
-\[
-2[\eta - \eta^2]
-\]
-
-```{r, echo=F, message=F, warning=F}
-library(ggplot2)
-typeIerror = function(theta1, theta2, v1, v2, alpha=0.05, side='both'){
-  if(side == 'both'){
-    calpha = qnorm(1 - alpha/2)
-  } else if(side == 'upper'){
-    calpha = qnorm(1 - alpha)
-  } else if (side == 'lower'){
-    calpha = qnorm(alpha)
-  } else {
-    print("Must specify side as 'both', 'upper', or 'lower'. 
-            Using default 'both' for 2-sided test.")
-    calpha = qnorm(1 - alpha/2)
-  }
-
-  return(
-    (pnorm(-calpha - theta1/sqrt(v1)) + 1 - pnorm(calpha - theta1/sqrt(v1))) *
-    (pnorm(calpha - theta2/sqrt(v2)) + pnorm(-calpha - theta2/sqrt(v2))) +
-      (pnorm(calpha - theta1/sqrt(v1)) + pnorm(-calpha - theta1/sqrt(v1))) *
-      (pnorm(-calpha - theta2/sqrt(v2)) + 1 - pnorm(calpha - theta2/sqrt(v2)))
-  )
-}
-tmp = setNames(data.frame(expand.grid(10:100, 10:100)), c('n1', 'n2'))
-tmp$err0 = sapply(1:nrow(tmp), 
-                  FUN=function(i) 
-                    typeIerror(theta1 = 0, 
-                               theta2 = 0, 
-                               v1 = 2/tmp$n1[i], 
-                               v2 = 2/tmp$n2[i]))
-tmp$err2 = sapply(1:nrow(tmp), 
-                  FUN=function(i) 
-                    typeIerror(theta1 = 0.2, 
-                               theta2 = 0.2, 
-                               v1 = 2/tmp$n1[i], 
-                               v2 = 2/tmp$n2[i]))
-tmp$err5 = sapply(1:nrow(tmp), 
-                  FUN=function(i) 
-                    typeIerror(theta1 = 0.5, 
-                               theta2 = 0.5, 
-                               v1 = 2/tmp$n1[i], 
-                               v2 = 2/tmp$n2[i]))
-tmp$err8 = sapply(1:nrow(tmp), 
-                  FUN=function(i) 
-                    typeIerror(theta1 = 0.8, 
-                               theta2 = 0.8, 
-                               v1 = 2/tmp$n1[i], 
-                               v2 = 2/tmp$n2[i]))
-
-ggplot(tmp) + geom_point(aes(n1, n2, color=err2)) + 
-  scale_color_gradient(low='white', high='red')
-
-ggplot(tmp) + geom_point(aes(n1, n2, color=err5)) + 
-  scale_color_gradient(low='white', high='red')
-
-ggplot(tmp) + geom_point(aes(n1, n2, color=err8)) + 
-  scale_color_gradient(low='white', high='red')
-
-tmp2 = setNames(data.frame(expand.grid(10:100, 0.01*0:100)), c('n', 'theta'))
-tmp2$err = sapply(1:nrow(tmp2), 
-                  FUN=function(i) 
-                    typeIerror(theta1 = tmp2$theta[i], 
-                               theta2 = tmp2$theta[i], 
-                               v1 = 2/tmp2$n[i], 
-                               v2 = 2/tmp2$n[i]))
-ggplot(tmp2) + geom_point(aes(n, theta, color=err)) +
-  scale_color_gradient(low='white', high='red')
-```
-
-However, others have noted that statistical significance may not correspond to replication if the significant estimates do not share the same sign. 
-That is, if study 1 has a positive significant estimate and study 2 has a negative and significant estimate, one likely would not conclude they replicate.
-Thus, an alternative to (\ref{eq:metsig}) could be
-\begin{equation}
-\mathbf{1}\{p_1 \leq 0.05 \wedge p_2 \leq 0.05\}\mathbf{1}\{\text{sign}(T_1) = \text{sign}(T_2)\} + 
-\mathbf{1}\{p_1 > 0.05 \wedge p_2 > 0.05\}
-\label{eq:metsigsign}
-\end{equation}
-These are related analyses that have slightly different properties, which we illustrate below.
+For this paper, we interpret the OSC analyses in the tradition of statistical hypothesis testing.
+Each analysis can be seen of a test of the null hypothesis $H_0: \theta_1 = \theta_2$, the metric used can be seen as a test statistic, and determinations that a finding failed to replicate correspond to a rejection of this null hypothesis.
+Note that this is not the only structure we might infer here, and is not even the only way to operationalize these methods as hypothesis tests.

### 2018-04-11 src/metrics_funs.R
Test on simulation function

@@ -78,6 +78,6 @@ simulate_studies <- function(theta, v, k=NULL){
   
   }
 }
-simulate_studies(.2, 3/393, 2)
+simulate_studies(rnorm(2, 0, 10), 3/393, 2)
 
 

### 2018-04-11 src/metrics_funs.R
Updated code library w/working and new functions.

@@ -11,16 +11,48 @@
 ###-----------------------------------------------------------------------------###
 ###-----------------------------------------------------------------------------###
 
-##----------------------------------------##
+##-----------------------------------------------##
 ## Common Libraries
-##----------------------------------------##
+##-----------------------------------------------##
 library(dplyr); library(distr); library(ggplot2)
 
 
-##----------------------------------------##
+##-----------------------------------------------##
+## Power and p-value functions
+##-----------------------------------------------##
+power_fun <- function(theta, v){
+  
+  #############################################
+  # TAKES: theta; a vector of means
+  #         v; a vector of sampling variances
+  #
+  # RETURNS: power of design (vector)
+  #############################################
+  
+  return((1 - pnorm(1.96, theta/sqrt(v))) + pnorm(-1.96, theta/sqrt(v))) 
+  
+}
+
+pval_fun <- function(t, v){
+  
+  #############################################
+  # TAKES: t; a vector of estimates
+  #         v; a vector of sampling variances
+  #
+  # RETURNS: p-value of design (vector)
+  # 
+  # * Assumes normal dist. w/known variance
+  #############################################
+  
+  return(2 * (1 - pnorm(abs(t)/sqrt(v), 0, 1))) 
+  
+}
+
+##-----------------------------------------------##
 ## Simulating k studies
-##----------------------------------------##
+##-----------------------------------------------##
 simulate_studies <- function(theta, v, k=NULL){
+  
   #############################################
   # TAKES: theta; a vector of means
   #         v; a vector of sampling variances
@@ -66,18 +98,25 @@ simulate_studies <- function(theta, v, k=NULL){
   } else { # Valid: run simluation and return values
     
     ###-----Run simluation, compute properties, and return it as a DF.
+    draws = rnorm(k, theta, sqrt(v))
     return(
       data.frame(
-              t = rnorm(k, theta, sqrt(v)), # draws (T)
+              t = draws, # draws (T)
               v = v, # variances
               theta = theta, # means
-              power = (1 - pnorm(1.96, theta/sqrt(v))) + pnorm(-1.96, theta/sqrt(v)), # power of each study
-              p = 2 * (1 - pnorm(abs(t)/sqrt(v), 0, 1)) # p-value of each estimate
+              power = power_fun(theta, v), # power of each study
+              p = pval_fun(t, v) # p-value of each estimate
       )
     )
   
   }
 }
[…]

### 2018-04-11 notation.Rmd
Added stuff to notation, and added data.

@@
 For further discussion on this distinction see Hedges and Schauer (2018).
+
+# Notation Dictionary
+
+\begin{tabular}{l p{5cm}}
+$\theta_i$ & Effect parameter of study $i$, also called the study results.\\[1.1ex]
+$T_i$ & Estimate of study $i$\\[1.1ex]
+$v_i$ & Variance of estimate for study $i$\\[1.1ex]
+$p_i$ & $p$-value for study $i$\\[1.1ex]
+$1 - \beta_i$ & Power for study $i$\\[1.1ex]
+$\theta_{ij}$ & Effect parameter of study $i$ for finding $j$\\[1.1ex]
+$T_{ij}$ & Estimate of study $i$ for finding $j$\\[1.1ex]
+$v_{ij}$ & Variance of estimate for study $i$ for finding $j$\\[1.1ex]
+$p_{ij}$ & $p$-value for study $i$ for finding $j$\\[1.1ex]
+$1 - \beta_{ij}$ & Power for study $i$ for finding $j$\\[1.1ex]
+\end{tabular}

### 2018-04-11 Fishers_method/power_sims.R
Added code to Fisher's method sims.

@@ -14,7 +14,7 @@ power_sims<-function(N,m,delta,n){
     for (j in 1:m){
       p1[j]<-0
       while (p1[j]<=0.05) { #throw away p-values<=0.05
-        p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(2/n))))) #calculate p-values for the false negatives
+        p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(2/n)))/sqrt(2/n))) #calculate p-values for the false negatives
       }
     }
     #test statistic for Fisher's method, with transformation for truncating p-values
@@ -24,7 +24,7 @@ power_sims<-function(N,m,delta,n){
   print(power)
 }
 n<-36
-delta<-0.5
+delta<-1
 power_sims(100000,1,delta,n)
 power_sims(100000,2,delta,n)
 power_sims(100000,3,delta,n)

### 2018-04-11 Fishers_method/power_sims.R
added comment to Fishers method sims

@@ -3,7 +3,7 @@ power_sims<-function(N,m,delta,n){
   # TAKES: N; number of simulations 
   #        m; number of false negatives
   #        delta; effect size under alternative hypothesis, on scale of cohen's d
-  #        n; sample size 
+  #        n; median total sample size across studies
   # RETURNS: power of Fisher's test to reject
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
   ###############################################################################
@@ -14,7 +14,7 @@ power_sims<-function(N,m,delta,n){
     for (j in 1:m){
       p1[j]<-0
       while (p1[j]<=0.05) { #throw away p-values<=0.05
-        p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(2/n))))) #calculate p-values for the false negatives
+        p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(4/n))))) #calculate p-values for the false negatives
       }
     }
     #test statistic for Fisher's method, with transformation for truncating p-values
@@ -23,7 +23,7 @@ power_sims<-function(N,m,delta,n){
   power<-sum(T>155.4047)/N
   print(power)
 }
-n<-36
+n<-76
 delta<-0.5
 power_sims(100000,1,delta,n)
 power_sims(100000,2,delta,n)

### 2018-04-11 Ttest-pvals/simttest.R
Simulations of pvalues for t test

@@ -1,0 +1,37 @@
+#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+source("./src/metrics_funs.R")
+### Case 1 different thetas but same power
+N<- 1000
+tmatrix<-matrix(data = NA, nrow = N, ncol = 3)
+for (k in 1:N){
+  
+  t1<-rep(.2, 100)
+  sim_1_small<-simulate_studies(t1, 4/393) # 393 sample size needed for 80% power is small 
+  t2<-rep(.5, 100)
+  sim_1_med<-simulate_studies(t2, 4/64) # 64 is samples size needed for 80% power if medium
+  t3<-rep(.8,100)
+  sim_1_large<-simulate_studies(t3, 4/25) # 25 is samples size needed for 80% power if medium
+  
+  #difference in each pvalue
+  diff_smallmed<-sim_1_small$p - sim_1_med$p 
+  diff_smalllarge<-sim_1_small$p - sim_1_large$p 
+  diff_medlarge<-sim_1_med$p - sim_1_large$p 
+  
+  # find t statistics
+  tsmallmed<- mean(diff_smallmed)/(sd(diff_smallmed)/10)
+  tsmalllarge<- mean(diff_smalllarge)/(sd(diff_smalllarge)/10)
+  tmedlarge<- mean(diff_medlarge)/(sd(diff_medlarge)/10)
+  
+  tmatrix[k,1]<-tsmallmed
+  tmatrix[k,2]<-tsmalllarge
+  tmatrix[k,3]<-tmedlarge
+  
+}
+case1_smallmed<-tmatrix[,1]
+case1_smalllarge<-tmatrix[,2]
+case1_medlarge<-tmatrix[,3]
+sum(tmatrix[,1] > 1.98)/N
+sum(tmatrix[,2] > 1.98)/N
+sum(tmatrix[,3] > 1.98)/N
+
+  

### 2018-04-12 Ttest-pvals/simttest.R
Simulations of some cases of pvalues on the t-test

@@ -1,15 +1,15 @@
 #setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 source("./src/metrics_funs.R")
 ### Case 1 different thetas but same power
-N<- 1000
+N<- 1000 
 tmatrix<-matrix(data = NA, nrow = N, ncol = 3)
 for (k in 1:N){
   
-  t1<-rep(.2, 100)
+  t1<-rep(.2, 100) # Small theta values
   sim_1_small<-simulate_studies(t1, 4/393) # 393 sample size needed for 80% power is small 
-  t2<-rep(.5, 100)
+  t2<-rep(.5, 100) # medium theta values
   sim_1_med<-simulate_studies(t2, 4/64) # 64 is samples size needed for 80% power if medium
-  t3<-rep(.8,100)
+  t3<-rep(.8,100) # large theta values
   sim_1_large<-simulate_studies(t3, 4/25) # 25 is samples size needed for 80% power if medium
   
   #difference in each pvalue
@@ -18,6 +18,7 @@ for (k in 1:N){
   diff_medlarge<-sim_1_med$p - sim_1_large$p 
   
   # find t statistics
+  # single t statistics for given disparity
   tsmallmed<- mean(diff_smallmed)/(sd(diff_smallmed)/10)
   tsmalllarge<- mean(diff_smalllarge)/(sd(diff_smalllarge)/10)
   tmedlarge<- mean(diff_medlarge)/(sd(diff_medlarge)/10)
@@ -27,11 +28,11 @@ for (k in 1:N){
   tmatrix[k,3]<-tmedlarge
   
 }
-case1_smallmed<-tmatrix[,1]
-case1_smalllarge<-tmatrix[,2]
-case1_medlarge<-tmatrix[,3]
-sum(tmatrix[,1] > 1.98)/N
-sum(tmatrix[,2] > 1.98)/N
-sum(tmatrix[,3] > 1.98)/N
 
+sum(abs(tmatrix[,1]) > 1.98)/N # prob of test concluding that they are different for small and meduim differences
+sum(abs(tmatrix[,2]) > 1.98)/N # prob of test concluding that they are different for small and large differences
+sum(abs(tmatrix[,3]) > 1.98)/N # prob of test concluding that they are different for medium and large differences
+
+
+## Case 2 same thetas with different powers
   

### 2018-04-14 Ttest-pvals/simttest.R
Simulation of p-values with t test

@@ -1,16 +1,16 @@
-#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+.llllllllllllllllllllllllllllllllll..#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 source("./src/metrics_funs.R")
 ### Case 1 different thetas but same power
-N<- 1000 
+N<- 100000 
 tmatrix<-matrix(data = NA, nrow = N, ncol = 3)
 for (k in 1:N){
   
   t1<-rep(.2, 100) # Small theta values
-  sim_1_small<-simulate_studies(t1, 4/393) # 393 sample size needed for 80% power is small 
+  sim_1_small<-simulate_studies(t1, 4/(2*393)) # 393 sample size needed for 80% power is small 
   t2<-rep(.5, 100) # medium theta values
-  sim_1_med<-simulate_studies(t2, 4/64) # 64 is samples size needed for 80% power if medium
+  sim_1_med<-simulate_studies(t2, 4/(2*64)) # 64 is samples size needed for 80% power if medium
   t3<-rep(.8,100) # large theta values
-  sim_1_large<-simulate_studies(t3, 4/25) # 25 is samples size needed for 80% power if medium
+  sim_1_large<-simulate_studies(t3, 4/(2*25)) # 25 is samples size needed for 80% power if medium
   
   #difference in each pvalue
   diff_smallmed<-sim_1_small$p - sim_1_med$p 

### 2018-04-18 McNemar's Test/Notes on McNemar's Test.Rmd
McNemar's Test as a means of assessing replication.

---
title: "Properties of Methods for Assessing Replication"
author: "Sarah Peko-Spicer"
date: "February 7, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A framework for replication
Let there be $k$ pairs of studies each containing an original study and a replication study. Each pair of studies can be characterized by a vector of true effect parameters $(\theta_{1,j}, \theta_{2,j})$ that are unobserved with zero sampling variance. Each pair of studies is further associated with estimates $(T_{1,j}, T_{2,j})$ of these true effects. These estimates have sampling variances $V_{1,j}$ and $V_{2,j}$ which are assumed to be known. Let $T_{i,j} \sim N(\theta_{i,j}, V_{i,h})$ for $i=1,2$ and $j=1,2,\ldots, k$. Further, assume that $T_{i,j}$ are independent. Generally speaking, two studies replicate if $\theta_{1,j} \approx \theta_{2,j}$.\footnote{We observe \textit{strict} replication when $\theta_{1,j} = \theta_{2,j}$.}

## Assessing replication via significance and p-values

The Open Science Foundation (OSF) propose four different test for assessing replication based on significance and p-values. These methods include Fish'er method, McNemar's test, the Wilcoxon signed-rank test, and the \textit{t} test for dependent samples. In deriving properties for each of these tests, we assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of 0.05. Let $p_{1,j}$ denote the p-value of the i\textsuperscript{th} original study and let $p_{2,j}$ denote the p-value of the i\textsuperscript{th} replication study.

### The Umbrella of the Sign Test

Before diving into the technical details, it may be worth making a quick observation about the relationship between three of these tests. As shown in the diagram below, three of these tests can be thought of as an extension on the sign test. If we chose to test for differences in pairs of p-values using the sign test, we would let $\rho = P(p_{1,i} > p_{2,i})$ and test $H_0: \rho = 0.5$. In this case, we don't make use of the actual p-values but rather the \textit{sign} of the difference in p-values. Suppose we want to use a little more information. We might observe that p-values can be ranked, which would call for use of the Wilcoxon signed-rank test. Alternatively, we might observe that we can categorize p-values as being either significant or non-significant. In this case, our observations are binary and we can use McNemar's test. Finally, we might make full use of the "information" contained in the p-values, acknowledging that they are continuous variables and assuming that the difference in p-values is approximately normally distributed. In this case, we might gravitate towards the t test.

This framework for the relationship for these tests may, or may not, prove to be useful in understanding the properties of these tests in the context of replication. It could be that this relationship can help explain congruence between the tests. It could also be the case that since these tests are using different amounts of information about the p-values, that they may produce results that do not necessarily confirm one another.

 ```{r, echo=FALSE}
par(mar=c(1,1,1,1), mfrow=c(1,1))

names <- c("Sign Test", "Wilcoxon Test", "t Test", "McNemar's Test")
M <- matrix(nrow=4, ncol=3, byrow=T, data=0)

M[2,1] <- "Ranked"
M[3,1] <- "Normal"
M[4,1] <- "Binary"
plotmat(M, pos = c(1, 3), curve=0, name = names, lwd = 1,
        box.lwd = 2, cex.txt = 0.6, dtext=-0.9, box.size = 0.15,
        box.type = "square", box.prop = 0.5)

```

### McNemar's Test

McNemar's test is typicaly used to test the difference in proportions in paired data. In this case, the test is used to determine if there is a statistically significant difference between the proportion of statistically significant results among original studies and the proportion of statistically significant results among replication studies. Applying this test requires categorizing each $p_{j,i}$ as either significant (1) or non-significant (0). That is, each pair of studies is associated with a vector of p-values $(p_{1,i}, p_{2,i})$ that can take one of the following pairs of vlaues: (0,0), (0,1), (1,0), (1,1). This information is generally summarized in a 2 $\times$ 2 contingency table containing counts of occurrences in the data of these four pairs of values.

The null hypothesis for this test is one of marginal homogeneity. That is, under the null hypothesis, the marginal probabilities for each outcome are the same. In our case, this can be stated as

$$
\begin{aligned}
H_0 &: P(p_{1,i} \leq 0.05 \cap p_{2,i} \leq 0.05) + P(p_{1,i} \leq 0.05 \cap p_{2,i} > 0.05) \\
&= P(p_{1,i} \leq 0.05 \cap p_{2,i} \leq 0.05) + P(p_{1,i} > 0.05 \cap p_{2,i} \leq 0.05) 
\end{aligned}
$$

This can be be further simplified to

$$
\begin{aligned}
H_0 &: P(p_{1,i} \leq 0.05) = P(p_{2,i} \leq 0.05) \\
H_1 &: P(p_{1,i} \leq 0.05) \neq P(p_{2,i} \leq 0.05) 
\end{aligned}
$$

In other words, under the null hypothesis, if we were to draw a study pair at random, the probability of obtaining a significant result in the original study  is equal to the probability of obtaining a significant result in the replication study. Taking this observation into consideration and noting that $p_{j,i} \leq 0.05 \equiv \frac{|T_{j,i}|}{\sqrt{V_{j,i}}} \geq 1.96$, we can rewrite the null hypothesis as follows.

$$
\begin{aligned}
H_0 &: \sum\limits_{i=1}^k P \Big(\frac{|T_{1,i}|}{\sqrt{V_{1,i}}} \geq 1.96 |i \Big)P(i) = \sum\limits_{i=1}^k P \Big(\frac{|T_{2,i}|}{\sqrt{V_{2,i}}} \geq 1.96 |i \Big)P(i) \\
\end{aligned}
$$
So, we can reformulate this test of difference in paired proportions into a test of difference in average power. A rejection of the null hypothesis leads to the conclusion that there is a statistically significant difference between the average power of the original studies and the average power of the replication studies. Given that the power of any of the $2k$ studies depends on a variety of factors including sample size, true effect size, and population variance, it should be clear that a rejection of the McNemar null does not immediately imply that, on average, the studies do not replicate. Conceivably, the true effect sizes can be quite similar within study pairs but differences in sample size or variance between original and replication studies would lead to a rejection of the McNemar null. This suggests that equality in average power is perhaps not the best metric for assessing replication, or even rates of replication.

It is also worth exploring the power of McNemar's test, particularly in terms of its relationship to the power of the $2k$ studies. The power for McNemar's test is given by

$$
\begin{aligned}
\beta_M = \Phi \Big ( \frac{(p_{10} - p_{01})\sqrt{k} - z_{1-\alpha/2}\sqrt{p_{10} + p_{01}}}{\sqrt{p_{10} + p_{01} - (p_{10} - p_{01})^2}} \Big)
\end{aligned}
$$
Noting that $p_{10} - p_{01} = (p_{10} + p_{11}) - (p_{01} + p_{11})$ and that $p_{01} + p_{10} = (p_{10} + p_{11}) + (p_{01} + p_{11}) - 2p_{11}$, we can rewrite the power function as

$$
\begin{aligned}
\Phi \Big ( \frac{(p_{1+} - p_{+1})\sqrt{k} - z_{1-\alpha/2}\sqrt{p_{+1} + p_{1+} - 2p_{11}}}{\sqrt{p_{+1} + p_{1+} - 2p_{11} - (p_{1+} - p_{+1})^2}} \Big)
\end{aligned}
$$
where $p_{1+} = P(p_{1,i} \leq 0.05)$ and $p_{2+} = P(p_{2,i} \leq 0.05)$ for a randomly drawn $i$.

[Something to note, but for right now, I won't be making use of it...]

Back to $\beta_M$: If we assume that the original and replication studies are independent of one another, then $p_{10} = \frac{1}{k} \sum\limits_{j=1}^k (1-\beta_{1j})\beta_{2j} $ and $p_{01} = \frac{1}{k}\sum\limits_{j=1}^k \beta_{1j}(1-\beta_{2j})$ where $\beta_{1j}$ is the power of the j\textsuperscript{th} original study and $\beta_{2j}$ is the power of the j\textsuperscript{th} replication study.

One thing we may want to consider is what happens as the replication sample sizes grow large. If $n_{2j} \rightarrow \infty$ $\forall j$, then $\beta_{2j} \rightarrow 0$ (or power goes to 1) implying that $p_{10} \rightarrow \frac{1}{k} \sum\limits_{j=1}^k \beta_{1j}$ and $p$

### 2018-04-18 McNemar's Test/mcnemar_sims.R

@@ -1,0 +1,186 @@
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+### REPLICATION METRICS - MCNEMARS TEST FUNCTIONS
+###
+### This script includes functions and simulations related to the use of 
+### McNemar's Test as a means for assessing replication.
+###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+
+
+##-----------------------------------------------##
+## Sourced Scripts
+##-----------------------------------------------##
+source("./src/metrics_funs.R")
+
+##-----------------------------------------------##
+## McNemar's Test Function
+##-----------------------------------------------##
+
+mcnemar_fun <- function(theta1, theta2, n1, n2, k){
+  
+  #############################################
+  # TAKES: theta1; a vector of means (original)
+  #        theta2; a vector of means (replication)
+  #        n1; sample size (original)
+  #        n2; sample size (replication)
+  #        k; number of study pairs
+  #
+  # RETURNS: p-value of McNemar's Test
+  #############################################
+  
+  
+  orig.studies <- simulate_studies(rep(theta1, k), 2/n1) # simulate k original studies
+  rep.studies  <- simulate_studies(rep(theta2, k), 2/n2) # simulate k replication studies
+  
+  # number of study pairs for which original finding is significant and replication finding is not
+  b <- length(which(orig.studies$p <=.05 & rep.studies$p > .05)) 
+  
+  # number of study pairs for which original finding is not significant and replication finding is
+  c <- length(which(orig.studies$p > .05 & rep.studies$p <= .05))
+  
[…]

### 2018-04-18 Fishers_method/power_sims.R
table with delta values 0.2, 0.5, 0.8

@@ -30,4 +30,9 @@ power_sims<-function(N,M,delta,n){
 
 M<-c(1,2,3,4,5,10,32,64)
 n<-38
-power_sims_table<-data.frame(M,power_0.5)
+power_0.2<-power_sims(100000,M,0.2,n)
+power_0.5<-power_sims(100000,M,0.5,n)
+power_0.8<-power_sims(100000,M,0.8,n)
+
+power_sims_table<-data.frame(M,power_0.2,power_0.5,power_0.8)
+saveRDS(power_sims_table,"Fisher_power_sims.Rds")

### 2018-04-18 Ttest-pvals/simttest.R
Finished Case 1 and Case 2 presented with tables for each.

@@ -110,7 +110,38 @@ n3<-393
 
 thetasmall<-ttestsim(N, theta1, theta2, theta3, n1, n2, n3)
 
-tab2[1,1]<-sum(abs(thetasmall[,1]) > 1.98)/N # prob of test concluding that they are different for small and meduim differences
-tab2[1,2]<-sum(abs(thetasmall[,2]) > 1.98)/N # prob of test concluding that they are different for small and large differences
-tab2[1,3]<-sum(abs(thetasmall[,3]) > 1.98)/N # prob of test concluding that they are different for medium and large differences
+tab2[1,1]<-sum(abs(thetasmall[,1]) > 1.98)/N # prob of test concluding that they are different for 40 and 60 power
+tab2[1,2]<-sum(abs(thetasmall[,2]) > 1.98)/N # prob of test concluding that they are different for 40 and 80 power
+tab2[1,3]<-sum(abs(thetasmall[,3]) > 1.98)/N # prob of test concluding that they are different for 60 and 80 power
 
+# theta 0.5 with power 40,60,80
+theta1<-0.5
+theta2<-0.5
+theta3<-0.5
+
+n1<-24
+n2<-40
+n3<-63
+
+thetamed<-ttestsim(N, theta1, theta2, theta3, n1, n2, n3)
+
+tab2[2,1]<-sum(abs(thetamed[,1]) > 1.98)/N # prob of test concluding that they are different for 40 and 60 power
+tab2[2,2]<-sum(abs(thetamed[,2]) > 1.98)/N # prob of test concluding that they are different for 40 and 80 power
+tab2[2,3]<-sum(abs(thetamed[,3]) > 1.98)/N # prob of test concluding that they are different for 60 and 80 power
+
+# theta 0.8 with power 40,60,80
+theta1<-0.8
+theta2<-0.8
+theta3<-0.8
+
+n1<-10
+n2<-16
+n3<-25
+
+thetalarge<-ttestsim(N, theta1, theta2, theta3, n1, n2, n3)
+
+tab2[3,1]<-sum(abs(thetalarge[,1]) > 1.98)/N # prob of test concluding that they are different for 40 and 60 power
+tab2[3,2]<-sum(abs(thetalarge[,2]) > 1.98)/N # prob of test concluding that they are different for 40 and 80 power
+tab2[3,3]<-sum(abs(thetalarge[,3]) > 1.98)/N # prob of test concluding that they are different for 60 and 80 power
+
+tab2

### 2018-04-18 Notes on McNemar's Test.Rmd

--- 2018-04-18 McNemar's Test/Notes on McNemar's Test.Rmd
@@
 Let there be $k$ pairs of studies each containing an original study and a replication study.
-Each pair of studies can be characterized by a vector of true effect parameters $(\theta_{1,j}, \theta_{2,j})$ that are unobserved with zero sampling variance.
-Each pair of studies is further associated with estimates $(T_{1,j}, T_{2,j})$ of these true effects.
-These estimates have sampling variances $V_{1,j}$ and $V_{2,j}$ which are assumed to be known.
-Let $T_{i,j} \sim N(\theta_{i,j}, V_{i,h})$ for $i=1,2$ and $j=1,2,\ldots, k$.
-Further, assume that $T_{i,j}$ are independent.
-Generally speaking, two studies replicate if $\theta_{1,j} \approx \theta_{2,j}$.\footnote{We observe \textit{strict} replication when $\theta_{1,j} = \theta_{2,j}$.}
+Each pair of studies can be characterized by a vector of true effect parameters $(\theta_{1,i}, \theta_{2,i})$ that are unobserved with zero sampling variance.
+Each pair of studies is further associated with estimates $(T_{1,i}, T_{2,i})$ of these true effects.
+These estimates have sampling variances $V_{1,i}$ and $V_{2,i}$ which are assumed to be known.
+Let $T_{j,i} \sim N(\theta_{j,i}, V_{j,i})$ for $j=1,2$ and $i=1,2,\ldots, k$.
+Further, assume that $T_{j,i}$ are independent.
+Generally speaking, two studies replicate if $\theta_{1,i} \approx \theta_{2,i}$.
 
@@
 
-The Open Science Foundation (OSF) propose four different test for assessing replication based on significance and p-values.
-These methods include Fish'er method, McNemar's test, the Wilcoxon signed-rank test, and the \textit{t} test for dependent samples.
-In deriving properties for each of these tests, we assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of 0.05.
-Let $p_{1,j}$ denote the p-value of the i\textsuperscript{th} original study and let $p_{2,j}$ denote the p-value of the i\textsuperscript{th} replication study.
+The authors propose four different tests for assessing replication based on significance and p-values.
+These methods include Fisher's method, McNemar's test, the Wilcoxon signed-rank test, and the \textit{t} test for dependent samples.
+In what follows, we derive properties of each of these tests and discuss how they function within our proposed framework for replication.
 
-### The Umbrella of the Sign Test
-
-Before diving into the technical details, it may be worth making a quick observation about the relationship between three of these tests.
-As shown in the diagram below, three of these tests can be thought of as an extension on the sign test.
-If we chose to test for differences in pairs of p-values using the sign test, we would let $\rho = P(p_{1,i} > p_{2,i})$ and test $H_0: \rho = 0.5$.
-In this case, we don't make use of the actual p-values but rather the \textit{sign} of the difference in p-values.
-Suppose we want to use a little more information.
-We might observe that p-values can be ranked, which would call for use of the Wilcoxon signed-rank test.
-Alternatively, we might observe that we can categorize p-values as being either significant or non-significant.
-In this case, our observations are binary and we can use McNemar's test.
-Finally, we might make full use of the "information" contained in the p-values, acknowledging that they are continuous variables and assuming that the difference in p-values is approximately normally distributed.
-In this case, we might gravitate towards the t test.
-
-This framework for the relationship for these tests may, or may not, prove to be useful in understanding the properties of these tests in the context of replication.
-It could be that this relationship can help explain congruence between the tests.
-It could also be the case that since these tests are using different amounts of information about the p-values, that they may produce results that do not necessarily confirm one another.
-
- ```{r, echo=FALSE}
-par(mar=c(1,1,1,1), mfrow=c(1,1))
-
-names <- c("Sign Test", "Wilcoxon Test", "t Test", "McNemar's Test")
-M <- matrix(nrow=4, ncol=3, byrow=T, data=0)
-
-M[2,1] <- "Ranked"
-M[3,1] <- "Normal"
-M[4,1] <- "Binary"
-plotmat(M, pos = c(1, 3), curve=0, name = names, lwd = 1,
-        box.lwd = 2, cex.txt = 0.6, dtext=-0.9, box.size = 0.15,
-        box.type = "square", box.prop = 0.5)
-
-```
+The authors assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of 0.05.
+Let $p_{1,i}$ denote the p-value of the i\textsuperscript{th} original study and let $p_{2,i}$ denote the p-value of the i\textsuperscript{th} replication study.
 
@@
 where $p_{1+} = P(p_{1,i} \leq 0.05)$ and $p_{2+} = P(p_{2,i} \leq 0.05)$ for a randomly drawn $i$.
-
-[Something to note, but for right now, I won't be making use of it...]
-
-Back to $\beta_M$: If we assume that the original and replication studies are independent of one another, then $p_{10} = \frac{1}{k} \sum\limits_{j=1}^k (1-\beta_{1j})\beta_{2j} $ and $p_{01} = \frac{1}{k}\sum\limits_{j=1}^k \beta_{1j}(1-\beta_{2j})$ where $\beta_{1j}$ is the power of the j\textsuperscript{th} original study and $\beta_{2j}$ is the power of the j\textsuperscript{th} replication study.
-
-One thing we may want to consider is what happens as the replication sample sizes grow large.
-If $n_{2j} \rightarrow \infty$ $\forall j$, then $\beta_{2j} \rightarrow 0$ (or power goes to 1) implying that $p_{10} \rightarrow \frac{1}{k} \sum\limits_{j=1}^k \beta_{1j}$ and $p$

### 2018-04-22 Ttest-pvals/simttest.R
bootstrap of t test simulations

@@ -1,5 +1,25 @@
-.llllllllllllllllllllllllllllllllll..#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+### REPLICATION METRICS - T-TEST FUNCTIONS
+###
+### This script includes functions and simulations related to the use of 
+### T-Test as a means for assessing replication.
+###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+
+
+##-----------------------------------------------##
+## Sourced Scripts
+##-----------------------------------------------##
 source("./src/metrics_funs.R")
+
+
+##-----------------------------------------------##
+## T-Test Function
+##-----------------------------------------------##
 
 ttestsim<-function(N, theta1,theta2, theta3, powern1, powern2, powern3){
   
@@ -46,8 +66,11 @@ ttestsim<-function(N, theta1,theta2, theta3, powern1, powern2, powern3){
   
 }
 
-### Case 1 different thetas but same power
-N<- 100000 
+##-----------------------------------------------##
+## Case 1 different thetas but same power
+##-----------------------------------------------##
+
+N<- 10000
 tab<-matrix(data = NA, nrow = 3, ncol = 3, dimnames = list(c("Power 80", "Power 60", "Power 40"), c("Org. 0.2 Rep. 0.5", "Org. 0.2 Rep. 0.8", "Org. 0.5 Rep. 0.8")))
 # 80 % power
 theta1<-0.2
@@ -97,7 +120,11 @@ tab[3,3]<-sum(abs(power40diff[,3]) > 1.98)/N # prob of test concluding that they
 
 tab
 
+
+##-----------------------------------------------##
 ## Case 2 same thetas with different powers
+##-----------------------------------------------##
+
[…]

### 2018-04-22 Fishers_method/power_sims.R
update power_sims to take in vector of n test for "upper bound" on power with largest studies having non-null effects

@@ -1,24 +1,25 @@
 power_sims<-function(N,M,delta,n){
-  ###############################################################################
+  #########################################################################################
   # TAKES: N; number of simulations 
-  #        M; vector of number of false negatives
+  #        M; vector of number of non-null effects
   #        delta; effect size under alternative hypothesis, on scale of cohen's d
-  #        n; median total sample size across studies
+  #        n; vector of treatment/control sample size across studies (total sample size/2)
   # RETURNS: power of Fisher's test to reject
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
-  ###############################################################################
-  T<-c()
-  Power<-matrix()
+  #########################################################################################
+  T<-c() #empty list for Fisher's test statistic 
+  Power<-matrix() #empty matrix to store results
   for(k in 1:length(M)){
-    print(paste("m=",M[k]))
     for (i in 1:N){
+      #print(i) #uncomment to show progress for lengthy simulations
       p0<-runif(64-M[k],0.05,1) #draws p-values for the true negatives
       p1<-c() 
       for (j in 1:M[k]){
         p1[j]<-0
         while (p1[j]<=0.05) { #throw away p-values<=0.05
-          p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(2/n)))/sqrt(2/n))) #calculate p-values for the false negatives
+          p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(2/n[j])))/sqrt(2/n[j])))
         }
+        #print(n[j]) #uncomment to show progress for lengthy simulations 
       }
       #test statistic for Fisher's method, with transformation for truncating p-values
       T[i]<--2*sum(log((p0-0.05)/0.95))-2*sum(log((p1-0.05)/0.95)) 
@@ -28,11 +29,22 @@ power_sims<-function(N,M,delta,n){
   return(Power)
 }
 
-M<-c(1,2,3,4,5,10,32,64)
-n<-38
-power_0.2<-power_sims(100000,M,0.2,n)
-power_0.5<-power_sims(100000,M,0.5,n)
-power_0.8<-power_sims(100000,M,0.8,n)
[…]

### 2018-04-25 bootsims.R
Properly structured code for bootstrap of t test

@@ -1,3 +1,26 @@
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+### REPLICATION METRICS - T-TEST FUNCTIONS
+###
+### This script includes functions and simulations related to the use of 
+### T-Test as a means for assessing replication using the bootstrap method.
+###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+###-----------------------------------------------------------------------------###
+
+
+##-----------------------------------------------##
+## Sourced Scripts
+##-----------------------------------------------##
+source("./Ttest-pvals/simttest.R")
+
+
+##-----------------------------------------------##
+## T-Test Function Bootstrap Case 1
+##-----------------------------------------------##
+
 case1tab<-function(N, theta, powern1, powern2, powern3, rowNames, colNames){
   
   #############################################
@@ -74,6 +97,10 @@ case1means
 case1sd
 
 
+
+##-----------------------------------------------##
+## T-Test Function Bootstrap Case 2
+##-----------------------------------------------##
 
 
 case2tab<-function(N, theta, powern1, powern2, powern3, rowNames, colNames){

### 2018-04-25 Ttest-pvals/simttest.R
updated cases of confusion in power.

@@ -178,7 +178,7 @@ tab2
 
 
 tab3<-matrix(data = NA, nrow = 3, ncol = 3, dimnames = list(c( "0.2", "0.5", "0.8"), c("Power 50/Power 60", "Power 50/Power 90", "Power 60/Power 90")))
-# theta 0.2 with power 40,60,80
+# theta 0.2 with power 50,60,90
 theta1<-0.2
 theta2<-0.2
 theta3<-0.2

### 2018-05-21 significance/significance.Rmd
Updated significance writeup

@@
 
-Programs of research that examine the replicability of research findings have become more common in psychology. 
-However, there is some ambiguity over how to assess the findings of such programs; that is, how do we know if a finding has been replicated?
-The Open Science Collaboration (2015) argue, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.
-Subsequent programs have used some of these same analysis methods (e.g., Camerer, et al., 2016; Schweinsberg, et al., 2016).
+```{r, echo=F, warning=F, message=F}
+source("./code/error_rates.R")
+```
 
-It is not entirely clear what these methods say about replication.
-For one, an actual definition of replication remains somewhat elusive---or flexible---across these methods. 
-Saying that studies get "the same result" is somewhat subjective, and the OSC attempted to operationalize this by describing several different scenarios of studies getting "the same result".
-Moreover, interpreting the results of the OSC analyses also requires knowledge of the properties of the methods they (and others) use.
-For instance, how sensitive are these analyses? 
-The irony is that answering this second question, in some ways, requires an answer to the former: how do we define replication?
+Programs of research that examine the replicability of research findings have become more common in the social sciences. 
+On its face, assessing replication seems trivial: just repeat an experiment and check that you get the same results.
+However analyses of these programs reveal that we might mean several different things when we refer to "the same results."
+Indeed, the Open Science Collaboration (OSC) (2015) argued, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.
 
-In this paper, we clarify subjective notions about replication using a meta-analytic framework. 
-Given this definition, we then evaluate various analysis methods to determine their properties.
+Despite immediate criticism of the analyses (see Gilbert, Pettigrew, and King, 2015), the core set of OSC analyses has become something of a standard in replication research.
+Camerer, et al. (2016) conducted an almost identical research design with the same analytic methods on findings in experimental economics. 
+Schweinsberg, et al. (2016), while conducting several---i.e., more than one---replicates of the same findings, aggregated replicate results so that they could use the same analyses as the OSC.
+Thus, the idea that replication is impossible to assess has persisted across multiple high-profile replication research efforts.
 
-# Defining Successful Replication: Meta-analytic Perspectives
+Given their prevalence in the replication literature, the analyses conducted by the OSC have a few important aspects worth exploring.
+The first involves the definitions of replication that they imply. 
+For instance, comparing $p$-values versus effect estimates may suggest very different ways we may think about studies' "results" and whether they are "the same."
+The second are the properties of the methods used to assess definitions of replication.
+For most replication studies, authors seek to make determinations about whether a finding was successfully replicated.
+Doing so amid uncertainty---such as in effect estimates---runs into similar issues of null hypothesis tests. 
+Thus, unambiguous interpretation of the results of replication research, can depend on the properties of the methods authors use to make these determinations.
 
-Meta-analysis represents the "results" of an experiment by an effect parameter, denoted $\theta$.
+In this paper, we explore methods for assessing replication as proposed by the OSC.
+We argue that definitions of replication ought to involve effect parameters, rather than $p$-values. 
+Using this logic, we then frame the set of analyses common in replication research as null hypothesis tests about replication. 
+The properties of these tests are computed under some simplifying assumptions, and shown that there are conditions where they may be suboptimal.
+
+# Model & Framework
+
+One approach to assessing replication proceeds from standard meta-analytic models. 
+Meta-analysis represents the results of an experiment by an effect parameter, denoted $\theta$.
 This "true" effect reflects what the researcher would observe if they had an infinite sample size.
@@
 In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
-For the RPP and RPE programs, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
+For programs such as that of the OSC, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
 However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
@@
 The $T_i$ might differ because of both differences in the underlying true effects, but also due to random chance.
-The $\theta_i$ might differ, however, because studies sample participants from different populations, or if there are differences in experimental contexts and conditions between experiments.
+The same can be said for the resulting $p$-values---$p_1$ and $p_2$.
+The $\theta_i$ might differ, however, because studies sample participants from different populations, or due to differences in contexts and conditions between experiments.
 Thus, it may make sense to define replication in terms of differences in experiments rather than statistical noise.
-Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2017).
+Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2018).
 
@@
 However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
-Thus, a more practical definition would say studies replicate if $\theta_1 \approx \theta_2$.
+Thus, a more practical definition might posit that studies replicate if $\theta_1 \approx \theta_2$.
 For further discussion on this distinction see Hedges and Schauer (2018).
-
-The key distinction here, is that the definition of replication involves the parameters $\theta_i$. 
-Analyses of replication involve the estimates $T_i$ and their variances $v_i$, which we use make inferences about how similar the $\theta_i$ are.
-This paradigm allows us to evaluate properties of analysis methods.
+The key distinction here, is that the \textit{definition} of replication involves the parameters $\theta_i$, while \textit{analyses} involve the estimates $T_i$ and their variances $v_i$
 
@@
 Note that this is not the only structure we might infer here, and is not even the only way to operationalize these methods as hypothesis tests.
+
+# Correspondence in Statistical Significance
+
+One metric used by the OSC to determine if studies successfully replicated to to check that both the original and replicate study corresponded in sign and statistical significance.
+That is, findings were deemed to have replcated if both studies found significant effects in the same direction---i.e., both positive---or if both studies found null effects.
+This type of metric would seem to answer an important question about replication, particularly of published research which more often involves statistically significant findings.
+Indeed, Fisher (1935) noted that "we may say that a phenomenon is experimentally demonstrable when we know how to conduct an experiment which will rarely fail to give us statistically significant results."
+
+However, it has long been noted that differences between significant and nonsignificant findings need not correspond to differences in underlying effect parameters (e.g., Gelman & Stern, 2005). 
+Thus, it is entirely possible that $\theta_1 = \theta_2 > 0$, so that the studies replicate exactly, but only $T_1$ is statistically significant.
+Alternatively, it could happen that $\theta_1$ and $\theta_2$ are markedly different, but both studies involve statistically significant estimates.
+
+To get a sense of the properties of this test, assume that $\theta_1 = \theta_2 = \theta$, or that the studies replicate exactly.
+If this were true, we would want this test to infrequently reject this null hypothesis; in statistical terms, we would want the type I error rate to be low.
+The null hypothesis would be rejected if $T_1$ is significant and $T_2$ is either not significant or significant and has an opposite direction (or visa versa). 
+Thus, the type I error rate is given by
+<!-- \begin{align} -->
+<!-- \begin{split} -->
+<!-- & P[T_1/\sqrt{v_1} > 1.96]P[T_2/\sqrt{v_2} \leq 1.96] +  -->
+<!-- P[T_1/\sqrt{v_1} \leq 1.96]P[T_2/\sqrt{v_2} > 1.96] \\ +  -->
+<!-- & P[T_1/\sqrt{v_1} < -1.96]P[T_2/\sqrt{v_2} \geq -1.96] +  -->
+<!-- P[T_1/\sqrt{v_1} \geq -1.96]P[T_2/\sqrt{v_2} < -1.96] -->
+<!-- \end{split} -->
+<!-- \end{align} -->
+<!-- Under the model, this can be written as  -->
+\begin{align*}
+\text{level }
+& = [1 - \Phi\left(1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(1.96 - \theta/\sqrt{v_2}\right)\\ 
+& + \Phi\left(1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(1.96 - \theta/\sqrt{v_2}\right)]\\ 
+& + \Phi\left(-1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)]\\
+& + [1 - \Phi\left(-1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)
+\end{align*}
+
+When $\theta_1 = \theta_2 = 0$, so that both studies involve null effects, then the type I error rate is 9.5\%.
+However, when $\theta_1 = \theta_2 \neq 0$, the frequency of type I errors depends on the power of each study to detect a non-null effect.
+Figure \ref{fig:sig:falsepos} shows the false positive rate of the test as a function of the power of study 1, and the relative size of study 2 to study 1. 
+For instance, if study 1 had 60\% power to detect $\theta_1$ and study 2 had 60\% power to detect $\theta_2$---so that the studies are the same size---then the false positive rate will be just under 50\%!
+
+In Figure \ref{fig:sig:falsepos}, we see that the false positive rates are nearly all above the nominal 5\%. 
+This is not necessarily an artifact of the scale of the graph. 
+Indeed, if study 2 had an infinite sample size, so that $v_2 \to 0$ and $\theta_2/\sqrt{v_2} \to 0$, then the false positive rate of this test for replication converges to one minus the power of the initial study:
+\[
+\text{Error rate } \to \beta_1
+\]
+where $1 - \beta_1$ is the power of study 1.
+Thus, unless study 1 has at least 95\% power, determining if studies replicated according to correspondence of sign and statisitical significance can result in type I error rates well above 5\%.
+
+```{r, echo=F, message=F, fig.cap = "False Positive Rate of Test for Replication.
+This plot displays the false positive rate of determining replication according to correspondence of sign and statistical significance as a function of the size of study 1 and study 2.
+Each colored line corresponds to study 2 being a multiple of the size of study 1.
+The dashed line corresponds to a 5\\% false positive rate.
+\\label{fig:sig:falsepos}"}
+falsepos_plt
+```
+
+Conversely, if we assume $\theta_1 \neq \theta_2$, then we would want the test to reject the null hypothesis frequently---that is, we would want a low type II error rate.

### 2018-05-22 significance/significance.Rmd
Edited significance section

@@
 \begin{align*}
-\text{level }
+P[\text{Type I Error}]
 & = [1 - \Phi\left(1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(1.96 - \theta/\sqrt{v_2}\right)\\ 
@@
 \end{align*}
+\noindent
+where $\Phi$ is the standard normal distribution function.
 
@@
 Conversely, if we assume $\theta_1 \neq \theta_2$, then we would want the test to reject the null hypothesis frequently---that is, we would want a low type II error rate.
+To conclude that the studies do not replicate would require one finding to be significant and in one direction (e.g., positive), and the other to be either nonsignificant or significant but in the opposite direction (e.g., negative).
+Under the model, the probability of concluding that the studies do replicate when $\theta_1 \neq \theta_2$ can be written as:
+\begin{align*}
+P[\text{Type II Error}]
+& = [1 - \Phi\left(1.96 - \theta_1/\sqrt{v_1}\right)][1 - \Phi\left(1.96 - \theta_2/\sqrt{v_2}\right)]\\ 
+& + [\Phi\left(1.96 - \theta_1/\sqrt{v_1}\right) - \Phi\left(-1.96 - \theta_1/\sqrt{v_1}\right)]
+    [\Phi\left(1.96 - \theta_2/\sqrt{v_2}\right) - \Phi\left(-1.96 - \theta_2/\sqrt{v_2}\right)]\\ 
+& + \Phi\left(-1.96 - \theta_1/\sqrt{v_1}\right) \Phi\left(-1.96 - \theta_2/\sqrt{v_2}\right)
+\end{align*}
+
+When one of the effect parameters, say $\theta_2$ is equal to zero, then this can be written in terms of the power of the other study:
+\[
+P[\text{Type II Error}]
+ = .05 + 0.95 \beta_1
+\]
+where $1 - \beta_1$ is the power of study 1. 
+Thus, if study 1 has low power, so that $\beta_1$ is large, then this test would be likely to conclude that the studies successfully replicated. 
+For instance, if study 1 has power 80\%, then this test would conclude that the studies do not replicate 24\% of the time. 
+However, if the power were lower, say 60\%, then the test would make this error 43\% of the time.
+
+If both of the results are nonzero but the same direction (e.g., both positive), then the error rate will be large whenever both studies have high (or low) power. 
+For instance, if $\theta_1$ is much larger than $\theta_2$, but both have power 80\%, then this test will conclude that the studies replicate almost 68\% of the time! 
+Similarly, if both studies have low power then there is a higher probability that both will return nonsignificant results, and hence we would conclude that the studies replicated despite $\theta_1 \neq \theta_2$.
+In fact, the only way for the false negative rate to decrease is for one study to have very high power, and the other to have very low power.
+
+In summary, the properties of this metric will depend heavily on the power of each study to detect a non-null effect, but in a way that can be counterintuitive.
+If the studies replicate exactly, then this test will falsely conclude that they do not a large percentage of the time unless both either have very high power (e.g., >90\%) or very low power (e.g., < 10\%).
+Moreover, it will be difficult to find scenarios where the type I error rate is below 9.5\%, and realistically could be much higher than that.
+Conversely, if the studies involve very different effect parameters, the test will often conclude that studes replicate. 
+If effects are in the same direction, the power of this test will decrease in the sample size of each individual experiment. 
+Only if one effect is positive and the other is null or negative will the error rate decrease with sample size.
+In other words, there are fairly plausible scenarios where collecting more data---having larger studies---can \textit{increase} the type I or type II error rates of this test!

### 2018-05-22 significance/significance.Rmd
fixed typos

@@
 In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
-For programs such as that of the OSC, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
+For programs such as that of the OSC, this involves $\theta_1$ for the original study and $\theta_2$ for the replicate.
 However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
@@
 One metric used by the OSC to determine if studies successfully replicated to to check that both the original and replicate study corresponded in sign and statistical significance.
-That is, findings were deemed to have replcated if both studies found significant effects in the same direction---i.e., both positive---or if both studies found null effects.
+That is, findings were deemed to have replicated if both studies found significant effects in the same direction---i.e., both positive---or if both studies found null effects.
 This type of metric would seem to answer an important question about replication, particularly of published research which more often involves statistically significant findings.
@@
 where $1 - \beta_1$ is the power of study 1.
-Thus, unless study 1 has at least 95\% power, determining if studies replicated according to correspondence of sign and statisitical significance can result in type I error rates well above 5\%.
+Thus, unless study 1 has at least 95\% power, determining if studies replicated according to correspondence of sign and statistical significance can result in type I error rates well above 5\%.
 
@@
 Moreover, it will be difficult to find scenarios where the type I error rate is below 9.5\%, and realistically could be much higher than that.
-Conversely, if the studies involve very different effect parameters, the test will often conclude that studes replicate. 
+Conversely, if the studies involve very different effect parameters, the test will often conclude that studies replicate. 
 If effects are in the same direction, the power of this test will decrease in the sample size of each individual experiment. 

### 2018-05-30 significance/significance.Rmd
Edits to forumlas and power section.

@@
 
+A common assumption in meta-analysis is that the $T_i$ are independent, unbiased, normally distributed, and have a known variance, that is:
+\[
+T_i \stackrel{indep}{\sim} N(\theta_i, v_i)
+\]
+This is true for effect sizes such as $z$-transformed correlations, and is a large sample approximation for other effect sizes such as standardized and raw mean differences and log-odds ratios.
+Further, in this paper, we assume that within-study hypothesis tests involve $H_0: \theta_i = 0$, and are carried out at the $\alpha = 0.05$ level.
+
 We argue that one useful definition of replication involves the similarity not of estimates $T_i$, but of actual true effects $\theta_i$.
@@
 Note that this is not the only structure we might infer here, and is not even the only way to operationalize these methods as hypothesis tests.
+However, it is in the tradition of statistical hypothesis tests about heterogeneity across effects.
 
@@
 However, it has long been noted that differences between significant and nonsignificant findings need not correspond to differences in underlying effect parameters (e.g., Gelman & Stern, 2005). 
-Thus, it is entirely possible that $\theta_1 = \theta_2 > 0$, so that the studies replicate exactly, but only $T_1$ is statistically significant.
+Thus, it is entirely possible that $\theta_1 = \theta_2$, so that the studies replicate exactly, but only $T_1$ is statistically significant.
 Alternatively, it could happen that $\theta_1$ and $\theta_2$ are markedly different, but both studies involve statistically significant estimates.
@@
 & = [1 - \Phi\left(1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(1.96 - \theta/\sqrt{v_2}\right)\\ 
-& + \Phi\left(1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(1.96 - \theta/\sqrt{v_2}\right)]\\ 
-& + \Phi\left(-1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)]\\
-& + [1 - \Phi\left(-1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)
+&\quad + \Phi\left(1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(1.96 - \theta/\sqrt{v_2}\right)]\\ 
+&\quad + \Phi\left(-1.96 - \theta/\sqrt{v_1}\right) [1 - \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)]\\
+&\quad + [1 - \Phi\left(-1.96 - \theta/\sqrt{v_1}\right)] \Phi\left(-1.96 - \theta/\sqrt{v_2}\right)
 \end{align*}
@@
 Conversely, if we assume $\theta_1 \neq \theta_2$, then we would want the test to reject the null hypothesis frequently---that is, we would want a low type II error rate.
-To conclude that the studies do not replicate would require one finding to be significant and in one direction (e.g., positive), and the other to be either nonsignificant or significant but in the opposite direction (e.g., negative).
+To conclude that the studies do replicate would require both findings to be significant and in the same direction (e.g., positive), or to both be nonsignificant.
 Under the model, the probability of concluding that the studies do replicate when $\theta_1 \neq \theta_2$ can be written as:
@@
 P[\text{Type II Error}]
- = .05 + 0.95 \beta_1
+ = 0.025 + 0.925 \beta_1
 \]
@@
 Thus, if study 1 has low power, so that $\beta_1$ is large, then this test would be likely to conclude that the studies successfully replicated. 
-For instance, if study 1 has power 80\%, then this test would conclude that the studies do not replicate 24\% of the time. 
-However, if the power were lower, say 60\%, then the test would make this error 43\% of the time.
+For instance, if study 1 has power 80\%, then this test would conclude that the studies do not replicate 21\% of the time. 
+However, if the power were lower, say 60\%, then the test would make this error 39.5\% of the time.
 

### 2018-05-30 effect sizes/effect_sizes.Rmd
Started effect size analyses

---
title: "Effect Sizes"
output: pdf_document
---

# Confidence Interval Comparison

Another approach used by the OSC and others to determine if a replication was successful involves confidence intervals. 
They determined that studies replicated if the estimate from the initial experiment $T_1$ is contained in a 95\% confidence interval for the replication effect parameter $\theta_2$. 
The logic behind this can be seen in two different ways. 
On the one hand, it can be conceived of as determining if the effect parameter in the replicate $\theta_2$ is different from the estimate of the initial study $T_1$. 
Note that they will be equal with probability zero under the model.

Another way to think about this analysis is that it tests whether the underlying effect parameters are the same.
To see this, note that $T_1$ being contained in the 95\% confidence interval for $\theta_2$ is equivalent to 
\[
S = |T_1 - T_2|/\sqrt{v_2} \leq 1.96
\]
This is conceptually identical to testing a null hypothesis of $H_0: \theta_1 = \theta_2$ with test statistic $S$ being compared to a standard normal distribution.
However, under $H_0$, $S$ does not follow a standard normal distribution.

### 2018-05-30 Ttest-pvals/writeup.Rmd
Write up for t-test of pvalues start

---
title: "Write Up: OSF T-Test"
author: "Mena Whalen"
output: pdf_document
---

1. OSF Article
A) steps list of what they did
B) why they did it

2. Theory
why is it bad
case 1 to case 2
3. Results sims
explain table and results

### OSF Article
 In the OSF article the authors used a paired t-test to compare the p-values from the original study with its replicate study for the 100 studies. The authors omit one case where the p-values were unavailable for both original and replicate. The test statistic for the paired t-test looks like $t=\frac{\bar{X}_{Diff}}{\frac{SD_{Diff}}{\sqrt k}}$ where $\bar{X}_{Diff}$ is the mean of all the differences between original and replicate p-values, $SD_{Diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of pairs. The t-test is testing the hypothesis that these two groups are different from one another, meaning that the p-value from the first experiment done is the same p-value as the replicate study done. A paired t-test is used in this case since each study is being compared with its replicate so they are assumed to be controlling the particular experiment being conducted within each of the different groups, original and replicate.

 For the 99 pairs available the dependent t-test was done, finding a mean difference to be -0.2738438 showing that the replicate p-values were larger than the original. This leads to a test statistic of -8.2068 and less than 0.0001 p-value. This would lead to saying that the pairs are not the same or that their difference is equal to 0, thus the original studies p-values are not the same and the replicate studies p-values. Leading back to that the original studys' estimate is not the same as the replicates' estimate.

### Theory
#pvalues what goes into it
#power
#hypothesis meaning (mult. hypothesis)
#case 1 and case 2 what happens
 Using the set up provided with $T_i \sim N(\theta_i, v_i)$, if a study is significant, meaning p-value is less than or equal to 0.05, then $\frac{|T_i|}{\sqrt{v_i}} \geq 1.96$ if testing $H_0:\theta_i =0$. We assume that the variance is known for each study and is $\frac{4}{n}$ where $n$ is the sample size. This means that the test statistic used for determining the p-value is dependent on the size of the estimate and the sample size. This p-value is a random variable coming from the given data, the distribution of which is dependent on the power of the test. The power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$. Given that each original study and its replicate study has possibly different estimates and sample sizes to calculate its p-value means that the groups could have different statistical power for each test.

 When comparing the p-values of two paired groups using the t-test incorporates each cases' estimate and the sample size of each test. In a single example, say study 1, there would be $p_1$ the p-value from the orginal and $p_2$ the p-value from the replicate study. This is equivalent to $p_1=\Phi^{-1}\Big(\frac{|T_1|}{\sqrt{v_1}}\Big)$ and similar can be shown for $p_2$, so subtracting the two leads to $\Phi^{-1}\Big(\frac{|T_1|}{\sqrt{v_1}}\Big)-\Phi^{-1}\Big(\frac{|T_2|}{\sqrt{v_2}}\Big)$. This is dependent on the power of each statistical test using the $T$'s and $n$'s for each group. This scaled for k studies, the mean of the paired difference of p-values would be 
\begin{align}
\frac{1}{k} \sum_i ^k \Big[ \Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big)-\Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big) \Big]
\end{align}
This mean takes into account both the estimate and the sample size, thus the power to detect if the p-values are different from one another depends upon the power of both the original and the replicate studies.

 In the current situation the t-test is testing if the group of original studies are similar to the group of replicate studies over all studies using p-values to assess similarity. When examining all studies together this takes away from a single study being replicated. Each study is dependent upon the power of both original and replicate being able to detect an effect. This is not the same thing as testing if an individual study did replicate which would relate back in hypothesis test form to 
 \begin{align}
 H_0: \theta_1 = \theta_2 \hspace{1cm} H_A: \theta_1 \neq \theta_2
 \end{align}
where $\theta_1$ is the parameter from the original and $\theta_2$ from the replicate.

### 2018-05-30 effect sizes/effect_sizes.Rmd
Updated effect size stuff for CIs

@@
 However, under $H_0$, $S$ does not follow a standard normal distribution.
+Instead, it follows a normal distribution with mean zero and variance $1 + v_1/v_2$.
+
+If $\theta_1 = \theta_2$, then we would like this test to conclude that the studies replicate.
+This would require $S$ to be smaller than 1.96 in magnitude. 
+The probability that $S$ exceeds 1.96 would be the Type I error rate. 
+The exact error rate is given by:
+\[
+1 - \Phi\left(\frac{1.96}{\sqrt{1 + v_1/v_2}}\right) + \Phi\left(\frac{-1.96}{\sqrt{1 + v_1/v_2}}\right)
+\]
+If study 1 is much larger than study 2, so that $v_1/v_2$ is close to zero, than this error rate will be close to nominal---that is, the Type I error rate will be close to 0.05.
+However, the error rate will be increasing as $v_1/v_2$ gets larger. 
+If both studies are the same size so that $v_1 = v_2$, then the error rate is 32.7\%. 
+If study 2 is twice the size of study 1, so that $v_1 = v_2/2$, then the error rate is 51.4\%.
+
+[INSERT FIGURE? error rate vs. v1/v2]
+
+If, however, $\theta_1 \neq \theta_2$ we would want the test to determine that the studies do not replicate, so that $S$ is at least as large as 1.96 in magnitude.

### 2018-06-06 Fishers_method/Fisher's method_050318.Rmd
Add fisher's method .Rmd and pdf

---
title: "Fisher’s Method"
author: "Katie Fitzgerald"
date: "05/03/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("~/Desktop/NU/Replication/replication_metrics")
```

## Fisher's test

Among the 100 replication studies conducted by the Open Science Collaboration (OSC), 64 of them found null effects, defined as having a p-value greater than 0.05. The OSC authors applied Fisher's method to this set of non-significant p-values to test the hypothesis that these 64 replication studies did in fact have "no evidential value." That is, they tested whether the null hypothesis of zero effect held for each study and that there were no false negatives among them. Therefore, their hypothesis for the Fisher's test could be formalized as follows, where $\theta_i$ is the treatment effect for study $i$:
\begin{center}
$H_0$: $\theta_1=\theta_2=...=\theta_{64}=0$

$H_a$: at least one $\theta_i\neq 0,\quad i=1,...,64$
\end{center}
Often, Fisher's method is used to combine the results of several independent tests on the same overall hypothesis that one $\Theta=0$. For example, if researchers are interested in whether one particular treatment has a true effect, they could conduct *k* studies and pool the results to give an indication if the true $\Theta=0$. In that scenario, if the null hypothesis for each of the *k* studies is true (i.e. $H_0$ above holds), the p-values follow a uniform distribution on [0,1]. The test statistic for Fisher's method is $$X^2=-2\sum_{i=0}^{k} ln(p_i),$$ which follows a $\chi^2$ distribution with $2k$ degrees of freedom when $H_0$ is true. Low p-values give a larger test statistic, leading to a rejection that $\Theta=0$.

In the OSC scenario, however, the authors are conducting Fisher's test only among the 64 replication studies that did not have significant results. By design then, the p-values have a U[0.05,1] distribution. The OSC makes an appropriate transformation, $p_i^*=\frac{p_i-0.05}{0.95},$ so that $p_i^*\sim U[0,1]$ and Fisher's method can be applied. Similarly, the transformed test statistic is  $$X^2=-2\sum_{i=0}^{k} ln(p_i^*)=-2\sum_{i=0}^{k} ln(\frac{p_i-0.05}{0.95}),$$ which follows a $\chi^2$ distribution with $2k$ degrees of freedom under $H_0$.

Among the 64 non-significant replicate p-values, the OSC finds a test statistic of $X^2=155.83$, which falls just beyond the critical value $\chi^2_{128}=155.40$ and therefore is significant with $p=0.048$. The authors acknowledge this suggests there is at least one replication finding that could be a false negative, but they conclude "nonetheless, the wide distribution of p-values suggests against insufficient power as the only explanation of failures to replicate." This statement seems to indicate the authors are looking for evidence that there are no false negatives among the non-significant replication studies, which is their null hypothesis.

While it is never advised to conduct a test in order to conclude the null hypothesis, this switched framework is especially problematic when the test is underpowered to reject $H_0$ because the Type II error rate will be large. Even though the authors were able to reject their null hypothesis, we think that in general Fisher's method might be underpowered to answer the question at hand, so we want to explore this more rigorously. The intuition is that if there is just one $\theta_i\neq0$, and the remaining $k-1$ studies have p-values from a U[0,1] distribution, the one non-null effect will need to be very large to skew the distribution enough to look differently from U[0,1]. Or alternatively, there would need to be several non-null effects to skew the distribution to a detectable degree. In the scenario where a researcher is combining the results of $k$ studies of the same treatment to determine if an overall $\Theta=0$, this type of conservative test may be appropriate. In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_i$ has no bearing on the 63 other $\theta_i's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies. We will use power simulations to determine if the intuition holds that it would take many and possibly large non-null effects for Fisher's test to reject $H_0$.

We make a few simplifying assumptions about the framework of these 64 studies in order to proceed. Assume each study $i$ is testing the presence of some treatment effect $\theta_i$ and has equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$). Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($j=1,...n_i$), and assume $Y_{ij}^t\sim N(\mu_i +\theta_i, \sigma^2)$ and $Y_{ij}^c\sim N(\mu_i, \sigma^2)$, where $\sigma^2$ is known. Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in this subset of 64 replicate studies are of these two types (89%).

Note then, $\theta_i$ is the difference in means between the treatment and control groups, and its estimate $T_i=\overline{Y_{i.}^t}-\overline{Y_{i.}^c}$ has variance $v_i=\frac{\sigma^2}{n_{ci}} +\frac{\sigma^2}{n_{ti}}=\frac{2\sigma^2}{n_i}.$ For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma}$. Note then that $\delta_i$ is estimated by $d_i=\frac{T_i}{\sigma}=\frac{T_i}{\sqrt{\frac{v_i}{2/n_i}}}=\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}}.$ Since under $H_0$, $T_i\sim N(\theta_i,v_i) \Rightarrow \frac{T_i}{\sqrt{v_i}} \sim N(\frac{\theta_i}{\sqrt{v_i}},1)$ and $\sqrt{\frac{2}{n_i}}$ is a constant, then we have  $Var(d_i)=Var(\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}})=\frac{2}{n_i}$. Therefore, $d_i\sim N(\delta_i,\frac{2}{n_i})$.

The hypotheses of each of the 64 studies can be represented on the scale of Cohen's d as follows: $$H_{0i}: \delta_{i}=0$$ $$H_{ai}: \delta_{i}\neq0,$$ and the p-value of this test can therefore be calculated as $2(1-\Phi(\frac{|d_i|}{\sqrt{2/n_i}})).$

In our simulations, we will assume that there are $m$ false negatives among the 64 studies, $m=1,...64$. Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis. That is, we draw a random variable $d_i$ from a $N(\delta_i,\frac{2}{n_i})$ distribution, where $\delta_i\neq0$ and compute its p-value. We continue drawing $d_i's$ until we obtain $m$ p-values greater than 0.05 (due to the restriction in the OSC scenario of only considering replicate studies with non-significant results). We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{i=0}^{k} ln(\frac{p_i-0.05}{0.95})$. We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{l=0}^{N}I_{\{X^2_l>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.

We first consider the case when $n_i=76/2=38$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups. The results of the power simulations for median sample size are given below. Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d. We find that there needs to be 10 large effects $\delta_i=0.8$ in order to achieve close to the standard 80% power (power=0.77426). If there is only one large effect, Fisher's test only has approximately 8% power to reject. Even when all 64 studies have small effects, there is only 70% power.  
```{r,echo=FALSE}
library(knitr)
table <- readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/power sims table with changing n.RDS")
table_median<-table[,1:4]
#colnames(table_median)[2]<
kable(table_median,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power to reject H0 with median sample size (n=38)")

```

In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first. That is, if $m=1$ we let the non-null effect be from the largest study, if $m=2$ we let the non-null effects be from the two largest studies, etc. Note, however that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic. For example, the largest study has $n_i=768703/2=384351.5$ which has 100% power to detect a large effect of $\delta_i=0.8$ (see tableXX in Appendix). In other words, it would not have been possible to obtain a p-value less than 0.05 with this sample size and true $\delta_i=0.8$. We therefore take the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$. TableXX gives the largest sample sizes used for each $\delta$.

$\delta$ | $n$ | Power of OSC replicate to detect
---------|-----|---------------------------------
0.2      | 745 | 0.9713
0.5      | 159 | 0.9938
0.8      | 100 | 0.9999

See TableXX in the Appendix for the full sample size vectors that were used to conduct the "best-case scenario" power simulations.

As shown in TableXX, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it. When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340)

```{r, echo=FALSE}
table <- readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/power_large_n.RDS")
kable(table,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power to reject H0 with large n")
```

##Asymptotic power
If all 64 have 0.2, need n=278 (total sample size) for 80% power, n=45 for d=0.5, n=18 for d=0.8.

##Appendix
```{r,echo=FALSE,warning=FALSE,message=FALSE}
#test upper bounds of power (by largest sample size study having false negatives)
rpp_data<-read.csv("~/Desktop/NU/Replication/replication_metrics/data/rpp_data.csv")
data_forFisher<- rpp_data[ !is.na(rpp_data$T_pval_USE..R.) & (rpp_data$T_pval_USE..R.>0.05)  , ]
n<-sort(data_forFisher$N..R.,decreasing=TRUE)/2

source("~/Desktop/NU/Replication/replication_metrics/src/metrics_funs.R")
replicate_power<-data.frame(n,power_delta_0.2=power_fun(0.2,2/n),power_delta_0.5=power_fun(0.5,2/n),power_delta_0.8=power_fun(0.8,2/n))
library(knitr)
kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("n","d=0.2","d=0.5","d=0.8"),caption="Power to reject H0 with median sample size (n=38)")

large_n_used<-readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/large_n_used.RDS")
kable(large_n_used,format="markdown",digits=1,col.names=c("d=0.2","d=0.5","d=0.8"))

```

### 2018-06-13 Ttest-pvals/writeup.Rmd
finish writing about t-test of p-values. present results

@@
 ---
-
-1. OSF Article
-A) steps list of what they did
-B) why they did it
-
-2. Theory
-why is it bad
-case 1 to case 2
-3. Results sims
-explain table and results
 
@@
 ### Theory
-#pvalues what goes into it
-#power
-#hypothesis meaning (mult. hypothesis)
-#case 1 and case 2 what happens
+
  Using the set up provided with $T_i \sim N(\theta_i, v_i)$, if a study is significant, meaning p-value is less than or equal to 0.05, then $\frac{|T_i|}{\sqrt{v_i}} \geq 1.96$ if testing $H_0:\theta_i =0$.
@@
  When comparing the p-values of two paired groups using the t-test incorporates each cases' estimate and the sample size of each test.
-In a single example, say study 1, there would be $p_1$ the p-value from the orginal and $p_2$ the p-value from the replicate study.
+In a single example, say study 1, there would be $p_1$ the p-value from the original and $p_2$ the p-value from the replicate study.
 This is equivalent to $p_1=\Phi^{-1}\Big(\frac{|T_1|}{\sqrt{v_1}}\Big)$ and similar can be shown for $p_2$, so subtracting the two leads to $\Phi^{-1}\Big(\frac{|T_1|}{\sqrt{v_1}}\Big)-\Phi^{-1}\Big(\frac{|T_2|}{\sqrt{v_2}}\Big)$.
@@
 \begin{align}
-\frac{1}{k} \sum_i ^k \Big[ \Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big)-\Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big) \Big]
+\frac{1}{k} \sum_j ^k \Big[ \Phi^{-1}\Big(\frac{|T_{1j}|}{\sqrt{v_{1j}}}\Big)-\Phi^{-1}\Big(\frac{|T_{2j}|}{\sqrt{v_{2j}}}\Big) \Big]
 \end{align}
@@
  In the current situation the t-test is testing if the group of original studies are similar to the group of replicate studies over all studies using p-values to assess similarity.
-When examining all studies together this takes away from a single study being replicated.
-Each study is dependent upon the power of both original and replicate being able to detect an effect.
-This is not the same thing as testing if an individual study did replicate which would relate back in hypothesis test form to
+When examining all studies together this relates back to a single study being replicated.
+Each study's statistical test is dependent upon the power of both original and replicate being able to detect an effect.
+This is not the same thing as testing if an individual study did replicate, which would relate back to a hypothesis test of the form
  \begin{align}
@@
 where $\theta_1$ is the parameter from the original and $\theta_2$ from the replicate.
+
+ The t-test has two types of errors that can occur, one where the effect sizes from original to replicate not the same but the test says that they are not significantly different meaning the p-values are the same and the other is when the effect sizes are the same but the test concludes that the p-values are significantly different from one another.
+The first instance, which we will call Case 1, can occur when $\theta_{1j} \neq \theta_{2j}$ but the powers are the same, $(1-\beta_{1j}) = (1-\beta_{2j})$.
+Power is determined by the effect size and the sample size, so when the effect size is small the sample size needs to be bigger to have higher power.
+If two effect sizes are different but they have the same power their test statistic from the original study's work and the replicate experiment would look identical, $\frac{T_1}{\sqrt{V_1}} = \frac{T_2}{\sqrt{V_2}}$, since we know the $T_i$'s are different from one another then the denominator  would have to be scaled to be the same.
+The smaller effect size would have a larger samples size making the variance become smaller and the larger effect size would have a smaller samples size resulting in a larger variance, meaning $cV_{small}=V_{big}$.
+Since both test statistics are similar then their p-values will be similar resulting in the paired t-test determining that p-values from original to replicate are not significantly different from one another.
+The power of this test goes to the $\alpha$ level of the test, normally 0.05.
+
+ The second type of error that can occur from using p-values, call is Case 2, is from the same effect sizes with different power.
+In a simple though experiment if we compared one studies original and replicate using the t-test of p-values the the test statistic would look like $\frac{{p_1-p_2}}{\sqrt{V_{p_1-p_2}}}$.
+Letting the original study be fixed, then the sample size of the replicate study is let to go off to $\infty$ then the variance of the replicate is 0 making the p-value, $p_2$ go to 0.
+This leaves the test statistic only having the p-value from the original study resulting $\frac{p_1}{V_{p_1}}$, this makes the power of the test go to 1.
+The test will rarely say that the two p-values are significantly similar to one another.
+Another way to examine it is look at is similar to the two test statistics from the two works as  $\frac{T_1}{\sqrt{V_1}} \neq \frac{T_2}{\sqrt{V_2}}$ since the numerator is the same but the denominator is different since each sample has different power meaning different samples sizes and thus different variances.
+This makes the p-values different from one another so the test will highly reject that they are the same, resulting in the power of the test going to 0.80.
+
+### Results
+ Using simulations to demonstrate Case 1 and Case 2 are presented in Table 1 and Table 2 below. 
+```{r tab, echo=FALSE}
+
+source("./bootsims.R")
+kable(case1means)
+
+```

### 2018-06-13 Ttest-pvals/writeup.Rmd
explaining results

@@
  Using simulations to demonstrate Case 1 and Case 2 are presented in Table 1 and Table 2 below.
+In each simulation the effect sizes used were small 0.2, medium 0.5, and large 0.8.
+The powers used were 40,60,80 to range all possible powers that can occur.
+In Case 1 the rows represent the power of both original and replicate studies and the columns are the different effect sizes of the studies.
+For example, section (1,1) is if the original study had effect size 0.2 and the replicate had effect size 0.5, or vise versa, and both had 80% power.
+All results show power around 0.05 or below given the discussion made previously about Case 1.
+Notice that as the power decreases, going down the rows, the power of the test increases being capped at 0.05.
+This is from a higher power being able to detect effects easier than a lower powers so when comparing two high power studies with different effect but similar power the p-values will be closer together.For Case 2, the rows are the effect sizes for both orginal and replicate study and the columns are different power of each study.
+The result in section (1,1) would be the power of the t-test of p-values from the effect size of 0.2 with one study having power 40 and the second study having power 60.
+All results in the table are over 80% power as dicussed in the precious section since the same sizes are the same but with different powers resulting in different p-values for each resulting in the test saying they studies are not the same.
+In the second column with power 40 and power 80 the results are almost 1 which is similar to the thought experiment presented eariler since the sample size of the 80% powered test is much larger then the sample size of the 40% powered test, making the higher powered test's variance go to 0.
+
 ```{r tab, echo=FALSE}

### 2018-06-15 Ttest-pvals/writeup.Rmd
other computer update

--- 2018-05-30 Ttest-pvals/writeup.Rmd
@@
 This scaled for k studies, the mean of the paired difference of p-values would be
-\begin{align}
-\frac{1}{k} \sum_i ^k \Big[ \Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big)-\Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big) \Big]
-\end{align}
+
+$\frac{1}{k} \sum_j ^k \Big[ \Phi^{-1}\Big(\frac{|T_{1j}|}{\sqrt{v_{1j}}}\Big)-\Phi^{-1}\Big(\frac{|T_{2j}|}{\sqrt{v_{2j}}}\Big) \Big]$
+
 This mean takes into account both the estimate and the sample size, thus the power to detect if the p-values are different from one another depends upon the power of both the original and the replicate studies.

### 2018-06-20 Fishers_method/asympotic power.R
update to take vectors of n and delta

@@ -1,7 +1,27 @@
-power<-function(n,delta){
+power_median<-function(n,delta){
   1-pnorm(155.4047, mean=64*n*(1/4)*(delta^2), sqrt(64*4*n*(1/4)*(delta^2)))
 }
 par(mfrow=c(3,1))
-plot(power(1:1000,0.2), type='l')
-plot(power(1:1000,0.5), type='l')
-plot(power(1:1000,0.8), type='l')
+plot(power_median(1:1000,0.2), type='l')
+plot(power_median(1:1000,0.5), type='l')
+plot(power_median(1:1000,0.8), type='l')
+
+s<-c()
+Power<-c()
+power<-function(n,delta,M){
+  for(j in 1:length(M)){
+    for(i in 1:M[j]){
+      s[i]<-n[i]*1/4*delta^2
+    }
+    Power[j]<-1-pnorm(155.4047,1/2*sum(s),sqrt(sum(s)))+pnorm(-155.4047,1/2*sum(s),sqrt(sum(s)))
+  }
+  return(Power)
+}
+n<-rep(76,64)
+M<-c(1,2,3,4,5,10,32,64) #vector of number of non-null effects
+power(n,0.2)
+power(n,0.5)
+power(n,0.8)
+x1<-power(large_n_used[,1],0.2,M)
+x2<-power(large_n_used[,2],0.5,M)
+x3<-power(large_n_used[,3],0.8,M)

### 2018-06-20 Fishers_method/Fishers_method_short.Rmd
add .Rmd for shortened version of Fisher write-up

--- 2018-06-06 Fishers_method/Fisher's method_050318.Rmd
@@
 author: "Katie Fitzgerald"
-date: "05/03/2018"
-output:
-  pdf_document: default
-  html_document:
-    df_print: paged
+date: "06/18/2018"
+output: pdf_document
 ---
@@
 
-Among the 100 replication studies conducted by the Open Science Collaboration (OSC), 64 of them found null effects, defined as having a p-value greater than 0.05.
-The OSC authors applied Fisher's method to this set of non-significant p-values to test the hypothesis that these 64 replication studies did in fact have "no evidential value." That is, they tested whether the null hypothesis of zero effect held for each study and that there were no false negatives among them.
+Among the 100 replication studies conducted by the Open Science Collaboration (OSC), 64 found null effects, defined as having a p-value greater than 0.05.
+The OSC authors applied Fisher's method to this set of non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
 Therefore, their hypothesis for the Fisher's test could be formalized as follows, where $\theta_i$ is the treatment effect for study $i$:
@@
 \end{center}
-Often, Fisher's method is used to combine the results of several independent tests on the same overall hypothesis that one $\Theta=0$.
-For example, if researchers are interested in whether one particular treatment has a true effect, they could conduct *k* studies and pool the results to give an indication if the true $\Theta=0$.
-In that scenario, if the null hypothesis for each of the *k* studies is true (i.e.
-$H_0$ above holds), the p-values follow a uniform distribution on [0,1].
-The test statistic for Fisher's method is $$X^2=-2\sum_{i=0}^{k} ln(p_i),$$ which follows a $\chi^2$ distribution with $2k$ degrees of freedom when $H_0$ is true.
-Low p-values give a larger test statistic, leading to a rejection that $\Theta=0$.
 
-In the OSC scenario, however, the authors are conducting Fisher's test only among the 64 replication studies that did not have significant results.
-By design then, the p-values have a U[0.05,1] distribution.
-The OSC makes an appropriate transformation, $p_i^*=\frac{p_i-0.05}{0.95},$ so that $p_i^*\sim U[0,1]$ and Fisher's method can be applied.
-Similarly, the transformed test statistic is  $$X^2=-2\sum_{i=0}^{k} ln(p_i^*)=-2\sum_{i=0}^{k} ln(\frac{p_i-0.05}{0.95}),$$ which follows a $\chi^2$ distribution with $2k$ degrees of freedom under $H_0$.
+Because the OSC authors are conducting Fisher's test only among the 64 replication studies that did not have significant results and thus $p_i \geq 0.05$ for all $i$, they use the following transformation of Fisher's test statistic  $$X^2=-2\sum_{i=1}^{k} ln(p_i^*)=-2\sum_{i=1}^{k} ln(\frac{p_i-0.05}{0.95}),$$ where under $H_0$, the $p_i^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$.
+Low p-values give a larger test statistic, leading to a rejection of $H_0$.
 
-Among the 64 non-significant replicate p-values, the OSC finds a test statistic of $X^2=155.83$, which falls just beyond the critical value $\chi^2_{128}=155.40$ and therefore is significant with $p=0.048$.
-The authors acknowledge this suggests there is at least one replication finding that could be a false negative, but they conclude "nonetheless, the wide distribution of p-values suggests against insufficient power as the only explanation of failures to replicate." This statement seems to indicate the authors are looking for evidence that there are no false negatives among the non-significant replication studies, which is their null hypothesis.
+We believe that Fisher's test is not well suited to assess replication.
+In the event that you do reject $H_0$, this result only tells you that at least one study was a false negative, but it does not till you *which* study did in fact have a true effect.
+Even still, it provides no information about the size or direction of that true effect and whether or not it replicates the original finding.
+Presumably, a finding of "no false negatives" would be most informative in assessing replication in this scenario, but this can never be validly concluded from Fisher's method since that would require concluding the null hypothesis.
+While it is never advised to conduct a test in order to conclude the null hypothesis, this switched framework is especially problematic when the test is underpowered to reject $H_0$ because the Type II error rate will be large.
 
-While it is never advised to conduct a test in order to conclude the null hypothesis, this switched framework is especially problematic when the test is underpowered to reject $H_0$ because the Type II error rate will be large.
-Even though the authors were able to reject their null hypothesis, we think that in general Fisher's method might be underpowered to answer the question at hand, so we want to explore this more rigorously.
+Even though the OSC was able to reject their null hypothesis ($X^2=155.83, p=0.048$), we think that in general Fisher's method is underpowered to answer the question at hand, so we want to explore this more rigorously.
 The intuition is that if there is just one $\theta_i\neq0$, and the remaining $k-1$ studies have p-values from a U[0,1] distribution, the one non-null effect will need to be very large to skew the distribution enough to look differently from U[0,1].
 Or alternatively, there would need to be several non-null effects to skew the distribution to a detectable degree.
-In the scenario where a researcher is combining the results of $k$ studies of the same treatment to determine if an overall $\Theta=0$, this type of conservative test may be appropriate.
+In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall $\Theta=0$, this type of conservative test may be appropriate.
 In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_i$ has no bearing on the 63 other $\theta_i's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
@@
 
-We make a few simplifying assumptions about the framework of these 64 studies in order to proceed.
-Assume each study $i$ is testing the presence of some treatment effect $\theta_i$ and has equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$).
-Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($j=1,...n_i$), and assume $Y_{ij}^t\sim N(\mu_i +\theta_i, \sigma^2)$ and $Y_{ij}^c\sim N(\mu_i, \sigma^2)$, where $\sigma^2$ is known.
-Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in this subset of 64 replicate studies are of these two types (89%).
+For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma_i}$, where $\theta_i$ is the mean difference between the treatment and control groups in study $i$ and $\sigma_i$ is the constant variance among the treatment and control populations.
+Assuming equal sample sizes in the treatment and control groups within study $i$ (i.e.
+$n_i^t=n_i^c=n_i$), $\delta_i$ is estimated by $d_i\sim N(\delta_i,\frac{2}{n_i})$ (See Appendix for proof and discussion of simplifying assumptions).
+Under this framework, Fisher's method can be represented as testing the hypotheses
 
-Note then, $\theta_i$ is the difference in means between the treatment and control groups, and its estimate $T_i=\overline{Y_{i.}^t}-\overline{Y_{i.}^c}$ has variance $v_i=\frac{\sigma^2}{n_{ci}} +\frac{\sigma^2}{n_{ti}}=\frac{2\sigma^2}{n_i}.$ For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma}$.
-Note then that $\delta_i$ is estimated by $d_i=\frac{T_i}{\sigma}=\frac{T_i}{\sqrt{\frac{v_i}{2/n_i}}}=\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}}.$ Since under $H_0$, $T_i\sim N(\theta_i,v_i) \Rightarrow \frac{T_i}{\sqrt{v_i}} \sim N(\frac{\theta_i}{\sqrt{v_i}},1)$ and $\sqrt{\frac{2}{n_i}}$ is a constant, then we have  $Var(d_i)=Var(\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}})=\frac{2}{n_i}$.
-Therefore, $d_i\sim N(\delta_i,\frac{2}{n_i})$.
+\begin{center}
+$H_0$: $\delta_1=\delta_2=...=\delta_{64}=0$
 
-The hypotheses of each of the 64 studies can be represented on the scale of Cohen's d as follows: $$H_{0i}: \delta_{i}=0$$ $$H_{ai}: \delta_{i}\neq0,$$ and the p-value of this test can therefore be calculated as $2(1-\Phi(\frac{|d_i|}{\sqrt{2/n_i}})).$
+$H_a$: at least one $\theta_i\neq 0,\quad i=1,...,64$
+\end{center} and the 64 p-values can be calculated as $p_i=2(1-\Phi(\frac{|d_i|}{\sqrt{2/n_i}})).$
 
-In our simulations, we will assume that there are $m$ false negatives among the 64 studies, $m=1,...64$.
-Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
-That is, we draw a random variable $d_i$ from a $N(\delta_i,\frac{2}{n_i})$ distribution, where $\delta_i\neq0$ and compute its p-value.
-We continue drawing $d_i's$ until we obtain $m$ p-values greater than 0.05 (due to the restriction in the OSC scenario of only considering replicate studies with non-significant results).
-We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{i=0}^{k} ln(\frac{p_i-0.05}{0.95})$.
-We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{l=0}^{N}I_{\{X^2_l>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.
-
-We first consider the case when $n_i=76/2=38$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
-The results of the power simulations for median sample size are given below.
+We first consider the power of Fisher's method when $n_i=76/2=38$ for all $i$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
+The results of the power simulations for median sample size are given in TableXX (See Appendix for code and details on how the power simulations were conducted).
 Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
-We find that there needs to be 10 large effects $\delta_i=0.8$ in order to achieve close to the standard 80% power (power=0.77426).
+We find that there needs to be 10 large effects in order to achieve close to the standard 80% power (power=0.77426).
 If there is only one large effect, Fisher's test only has approximately 8% power to reject.
-Even when all 64 studies have small effects, there is only 70% power.
+Even when all 64 studies have small effects, Fisher's method only has 70% power to reject.
 ```{r,echo=FALSE}
@@
 #colnames(table_median)[2]<
-kable(table_median,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power to reject H0 with median sample size (n=38)")
+kable(table_median,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"), caption="Power to reject H0 with median sample size (n=38)")
 
@@
 In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
-That is, if $m=1$ we let the non-null effect be from the largest study, if $m=2$ we let the non-null effects be from the two largest studies, etc.
-Note, however that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic.
-For example, the largest study has $n_i=768703/2=384351.5$ which has 100% power to detect a large effect of $\delta_i=0.8$ (see tableXX in Appendix).
-In other words, it would not have been possible to obtain a p-value less than 0.05 with this sample size and true $\delta_i=0.8$.
-We therefore take the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
-TableXX gives the largest sample sizes used for each $\delta$.
-
-$\delta$ | $n$ | Power of OSC replicate to detect
----------|-----|---------------------------------
-0.2      | 745 | 0.9713
-0.5      | 159 | 0.9938
-0.8      | 100 | 0.9999
-
+That is, if there is just one non-null effect we let it be from the largest study, if there are two non-null effects we let them be from the two largest studies, etc.
+Note, however that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
+The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ were $n_i=745; 159;$ and $100$ respectively.
 See TableXX in the Appendix for the full sample size vectors that were used to conduct the "best-case scenario" power simulations.
 
-As shown in TableXX, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
+As shown in the first row of TableXX, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
 When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340)
@@
 ##Appendix
+
+Assume each study $i$ is testing the presence of some treatment effect $\theta_i$ and has equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$).
+Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($j=1,...n_i$), and assume $Y_{ij}^t\sim N(\mu_i +\theta_i, \sigma^2)$ and $Y_{ij}^c\sim N(\mu_i, \sigma^2)$, where $\sigma^2$ is known.
+Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in this subset of 64 replicate studies are of these two types (89%).
+
+Note then, $\theta_i$ is the difference in means between the treatment and control groups, and its estimate $T_i=\overline{Y_{i.}^t}-\overline{Y_{i.}^c}$ has variance $v_i=\frac{\sigma^2}{n_{ci}} +\frac{\sigma^2}{n_{ti}}=\frac{2\sigma^2}{n_i}.$ For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma}$.
+Note then that $\delta_i$ is estimated by $d_i=\frac{T_i}{\sigma}=\frac{T_i}{\sqrt{\frac{v_i}{2/n_i}}}=\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}}.$ Since under $H_0$, $T_i\sim N(\theta_i,v_i) \Rightarrow \frac{T_i}{\sqrt{v_i}} \sim N(\frac{\theta_i}{\sqrt{v_i}},1)$ and $\sqrt{\frac{2}{n_i}}$ is a constant, then we have  $Var(d_i)=Var(\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}})=\frac{2}{n_i}$.
+Therefore, $d_i\sim N(\delta_i,\frac{2}{n_i})$.
+
+In our simulations, we will assume that there are $m$ false negatives among the 64 studies, $m=1,...,64$.
+Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
+That is, we draw a random variable $d_i$ from a $N(\delta_i,\frac{2}{n_i})$ distribution, where $\delta_i\neq0$ and compute its p-value.
+We continue drawing $d_i's$ until we obtain $m$ p-values greater than 0.05 (due to the OSC restriction of only considering replicate studies with non-significant results).
+We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{i=0}^{k} ln(\frac{p_i-0.05}{0.95})$.
+We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{l=0}^{N}I_{\{X^2_l>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.
 ```{r,echo=FALSE,warning=FALSE,message=FALSE}
@@
 replicate_power<-data.frame(n,power_delta_0.2=power_fun(0.2,2/n),power_delta_0.5=power_fun(0.5,2/n),power_delta_0.8=power_fun(0.8,2/n))
+
+replicate_power<-rbind(replicate_power[1:12,],c("...","...","...","..."))
 library(knitr)
@@
 large_n_used<-readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/large_n_used.RDS")
-kable(large_n_used,format="markdown",digits=1,col.names=c("d=0.2","d=0.5","d=0.8"))
+large_n_table<-rbind(large_n_used[1:13,],c("...","...","..."),large_n_used[64,])
+rownames(large_n_table)[14]<-c("...")
+kable(large_n_table,format="markdown",digits=1,col.names=c("d=0.2","d=0.5","d=0.8"))
 

### 2018-06-20 McNemar's Test/draft_mcnemars_ttest.Rmd
Add files via upload

---
title: "Assessing Replication via P-values"
author: "Sarah Peko-Spicer & Mena Whalen"
date: "June 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Two of the methods proposed by the Open Science Foundation (2015) for assessing replication via p-values are McNemar's test and the t-test for paired data. We discuss these tests in tandem as they are both extensions on the sign test and, as a result, have similar properties in the context of replication. For both tests, we consider $k$ pairs of studies, each containing an original study and a replication study. As described earlier, the original studies are characterized by an estimate $T_{1j} \sim N(\theta_{1j}, v_{1j})$ for $j = 1, \ldots, k$. Similarly, the replication studies are characterized by an estimate $T_{2j} \sim N(\theta_{2j}, v_{2j})$ for $j= 1, \ldots, k$. Further, we let $p_{1j}$ denote the \textit{p}-value of the j\textsuperscript{th} original study and $p_{2j}$ denote the \textit{p}-value of the j\textsuperscript{th} replication study.

McNemar's test is used to determine if there is a difference between the proportion of statistically significant results among original studies and the proportion of statistically significant results among replication studies. Each \textit{p}-value from an original or replication study can be categorized as either significant or non-significant, creating a dichotomous variable suitable for the use of this test. This information is generally summarized in a 2 $\times$ 2 contingency table as shown below where $n_{11}$ denotes the number of study pairs in which both the original and replication findings are signficant, $n_{10}$ denotes the number of study pairs in which the original finding is significant but the replication finding is not, $n_{01}$ denotes the number of study pairs in which the original finding is not significant but the replication finding is, and $n_{00}$ denotes the number of study pairs in which neither the original finding nor the replication finding is significant.

\begin{table}[h!]
\begin{center}
\begin{tabular}{l  c  c } \\
& \multicolumn{2}{c}{Replication Finding} \\
\cline{2-3}
Original Finding & Significant & Non-Significant \\
\hline
Significant & $n_{11}$ & $n_{10}$ \\
Non-Significant & $n_{01}$ & $n_{00}$ \\
\hline
\end{tabular}
\end{center}
\end{table}

The null hypothesis for McNemar's test states that the marginal probabilities for each outcome are the same. That is, $H_0: p_{11} + p_{10} = p_{11} + p_{01}$ where the $P$'s are the probabilities of occurrence in cells with the corresponding labels. The test statistic is given by $\chi^2 = \frac{(n_{10} - n_{01})^2}{n_{10} + n_{01}}$. Under the null hypothesis, the test statistic follow a chi-squared distribution with one degree of freedom. The null hypothesis is reject for large values of the test statistic. The results of the OSF article are summarized in the contingency table below. 97% of original findings were significant compared to 36% of replication findings. The authors found a test statistic of $X^2 = 59.06$, rejected the null hypothesis, and concluded that the findings from the original studies did not replicate.

\begin{table}[h!]
\begin{center}
\begin{tabular}{l  c  c } \\
& \multicolumn{2}{c}{Replication Finding} \\
\cline{2-3}
Original Finding & Significant & Non-Significant \\
\hline
Significant & $35$ & $62$ \\
Non-Significant & $1$ & $2$ \\
\hline
\end{tabular}
\end{center}
\end{table}

Unlike McNemar's test, the t-test assesses replication by directly comparing original and replication \textit{p}-values within a study pair. The null hypothesis for the t-test states that the original and replication studies have the same \textit{p}-value (i.e. $H_0: p_{1j} = p_{2j}$). The test statistic is given by $t = \frac{\bar{X}_{diff}}{SD_{diff}/\sqrt{k}}$ where $\bar{X}_{diff}$ is the mean of all differences between original and replication \textit{p}-values (i.e. $\bar{X}_{diff} = \frac{1}{k} \sum\limits_{j=1}^k (p_{1j} - p_{2j})$), $SD_{diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of study pairs. A paired t-test is used in this case since the original and replication studies in each study pair are conducting the same experiment, using the same methods, and drawing conclusions on the same phenomenon.

In the OSF article, the authors used the t-test on $k=99$ studies. They omitted one study pair where the exact \textit{p}-value was not available for the original or replication study. For the 99 available study pairs, the authors found a mean difference, $\bar{X}_{diff} = -0.274$, indicating that the replication \textit{p}-values were larger than the original \textit{p}-values. The test statistic was $-8.207$ $(p < 0.0001)$. Thus, the authors concluded that the original and replication \textit{p}-values are significantly different and the original and replication findings are significantly different.

It should be noted that both the t-test and McNemar's test are used to assess similarity between a \textit{group} of original studies and a \textit{group} of replication studies. We assume that an original and replication study within a study pair are conducting the same experiment, using the same methods, to draw conclusions about the same phenomenon. However, in the OSF data we observe some variation between study pairs on these factors. That is, the 100 original studies may be conducting different experiments, using different methods, to draw conclusions about different phenomena. On these grounds alone, the t-test and McNemar's test are not well suited to assess whether an original finding has been replicated unless all $k$ study pairs are drawing conclusions about the same phenomenon. In the analysis that follows, we assume that this is, in fact, the case.

## Analysis of McNemar's Test

Two properties you might use to assess the quality of a statistical test are its Type I and Type II error rates. To better understand these properties for McNemar's test, we rewrite the null hypothesis to reflect its relationship to the power of the original and replication studies. First, note that that the null hypothesis can be rewritten as

$$
H_0: P(p_{1j} \leq 0.05) = P(p_{2j} \leq 0.05)
$$

In other words, under the null hypothesis, if we were to draw a study pair at random, the probability of obtaining a significant result in the original study is equal to the probability of obtaining a significant result in the replication study. Taking this observation into consideration and nothing that $p_{ij} \leq 0.05 \equiv \frac{|T_{ij}|}{\sqrt{v_{ij}}} \geq 1.96$, we can reformulate the null hypothesis as follows.

$$
H_0: \sum\limits_{j=1}^k P \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 |i \Big) P(i) = \sum\limits_{j=1}^k P \Big ( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96 |i \Big) P(i)
$$

That is, we can think of McNemar's test as a test of difference in the average power rather than a difference in paired proportions. As such, a rejection of the null hypothesis is akin to concluding that there is a statistically significant difference between the average power of the original studies and the average power of the replication studies. However, as we will show, this does not necessarily imply that the original findings were not replicated in the replication studies, if we take replication to be $\theta_{1j} = \theta_{2j}$. Similarly, we will show that a failure to reject the null hypothesis for McNemar's test does not necessarily imply that the findings did, in fact, replicate.

Assuming the true null hypothesis for replication is $H_0: \theta_{1j} = \theta_{2j}$, a Type I error occurs when the original and replication studies differ in their power to detect their common effect size. In this case, McNemar's test will almost always reject its null hypothesis leading to conclusions of irreplicability despite having identical true effects. The power of each of the $2k$ original and replication studies is determined by the true effect size $(\theta_{ij})$, the sample size $(n_{ij})$, and the population variance $(\sigma^2)$. If the original and replication studies have the same true effect size and population variance, we can let their powers differ by increasing the sample size of the replication studies. As $n_{2j}$ grows large, $\frac{|T_{2j}|}{\sqrt{v_{2j}}}$ will also grow large. If we assume that original and replication studies are independent then,

$$
n_{10} = \sum\limits_{j=1}^k \mathbf{1}\Big\lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 \Big\rbrace \mathbf{1}\Big\lbrace \frac{|T_{2j}|}{\sqrt{v_{2j}}} < 1.96 \Big\rbrace \rightarrow 0
$$

$$
n_{01} = \sum\limits_{j=1}^k \mathbf{1} \Big \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96  \Big \rbrace \mathbf{1} \Big \lbrace \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96  \Big \rbrace \rightarrow \sum\limits_{j=1}^k \mathbf{1} \Big \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \Big \rbrace
$$
This implies that the McNemar's test statistic $X^2$  $\rightarrow$ $\sum\limits_{j=1}^k \mathbf{1} \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \rbrace$ as the replication sample sizes grow large. Let's assume that all $k$ original studies have 80% power to detect their true effect. Then, $X^2$ approaches $0.2k$ as the replication sample sizes increase. Recall that McNemar's test rejects if $X^2 > \chi^2_{[0.05,1]}$. In this instance, as long as $k>2$, McNemar's test will always reject despite the true effect sizes being identical for all $k$ study pairs.

A type II error occurs when the original and replication studies have different true effects (i.e. $\theta_{1j} \neq \theta_{2j}$) but have the same power to detect those effects (i.e. $(1-\beta_{1j}) = (1-\beta_{2j})$). The impact of the power of the original and replication studies on this test's Type II error rate is easiest understood through a consideration of the test's power function. The power function for McNemar's test is given by

$$
\Phi \Big \lbrace \frac{(p_{10} - p_{01})\sqrt{k} - z_{1 - \alpha/2}\sqrt{p_{10} + p_{01}}}{\sqrt{p_{10} + p_{01} - (p_{10} - p_{01})^2} } \Big \rbrace
$$
where $\Phi$ is the standard normal distribution function and $p_{10}$ and $p_{01}$ can be written as $\sum\limits_{j=1}^k (1-\beta_{1j})\beta_{2j}$ and $\sum\limits_{j=1^k} \beta_{1j}(1-\beta_{2j})$, respectively. When the power of the original study is equal to the power of the replication study for all $k$ studies, $p_{10} = p_{01}$. The power of McNemar's test can then be reduced to $\Phi \lbrace -z_{1-\alpha/2} \rbrace$. Assuming that $z=1.96$, we know the power is 0.05. Thus, when the true effects of the original and replication studies are different, McNemar's test will only reject the null hypothesis 5% of the time. In other words, the Type II error rate is roughly 95%.

## Analysis of the t-test

Like McNemar's test, the power of the t-test to detect if the \textit{p}-values of the original studies are different from those of the replication studies depends upon the power of both the original and replication studies. Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$. The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function.. Assuming that the variance is known for such a study and is $\frac{4}{n}$ where $n$ is the sample size then, the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$. Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as

$$
\frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
$$

 Given that this mean takes into account both the estimate and sample size of the original and replication studies, the power to detect a difference in \textit{p}-values depends on the power of both the original and replication studies.

Again, we consider the Type I and Type II error for the t-test. A Type I error occurs when the effect sizes from the original and replication studies are equal but the t-test concludes that their \textit{p}-values are significant different from one another. This is likely to occur when the power of the original and replication studies differ. In a simple thought experiment, we consider a single study pair. The t-test used to compare the original and replication findings in this pair would have a test statistic of $\frac{{p_1-p_2}}{\sqrt{V_{p_1-p_2}}}$. If we fix the parameters of the original study and let the sample size of the replication studies tend to $\infty$, the variance of the replication estimate will tend to 0 and, as a result, $p_2$ will also tend to 0. This would not be unusual in a real-world appication as researchers often increase replication sample sizes to get a more precise estimate of the original finding. In this scenario, the t-test stastic depends only on the \textit{p}-value of the original study $\Big( \frac{p_1}{\sqrt{V_{p_1}}}\Big)$ and the error rate tends to 1. In other words, the t-test will rarely conclude that the two \textit{p}-values are similar to one another. Another way to think about this error rate is to consider the case where the original and replication test statistics differ (i.e. $\frac{T_1}{\sqrt{V_1}} \neq \frac{T_2}{\sqrt{V_2}}$). If the original and replication studies have different power to detect the same effect, the numerators of these statistics will be similar, but the sample sizes and variances will differ. As a result, the \textit{p}-values from the original and replication studies will be quite different, and once again the Type I error rate will tend towards 0.80.

A Type II error occurs when $\theta_{1j} \neq \theta_{2j}$ but the powers are the same: $(1-\beta_{1j}) = (1-\beta_{2j})$. The power of the original and replication studies is determined by the effect size and the sample size. When the effect size is small, the sample size needs to be relatively large in order for the test to be well-powered. If two studies have different effect sizes but the same power, their test statistics can conceivably be identical (i.e. $\Big ( \frac{T_1}{\sqrt{V_1}} = \frac{T_2}{\sqrt{V_2}} \Big)$). For example, the study with the smaller effect size could have a larger sample size, reducing its variance while the study with the larger effect size could have a smaller sample size, resulting in a larger variance such that $cV_{small} = V_{big}$. Then, both test statistic would be similar enough to produce similar \textit{p}-values, even if their effect sizes are wildly different from one another. In that case, a t-test would fail to reject the null hypothesis despite the fact that the true effect sizes are different from one another. We posit that the power of this test will tend towards 0.05 (i.e. the Type II error rate tends to 0.95).

## Simulation Results

In data simulations to test our analysis of the Type I error rate, we fix the effect size of the original and replication studies at either a small, medium, or large effect. As is convention, we consider $\theta_{ij} = 0.2$ to be a small effect, $\theta_{ij} = 0.5$ to be a medium effect, and $\theta_{ij} = 0.8$ to be a large effect. For each of these effect size specifications, we then consider the case where (1) the original study is powered at 40% and the replication study is powered at 60%, (2) the original study is powered at 60% and the replication study is powered at 80%, and (3) the original study is powered at 40% and the replication study is powered at 80%. These power specifications are consistent with the range observed in the OSF data. For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the t-test or McNemar's test reject the null hypothesis in favor of the alternative. The results of these simulations are displayed in Table 1. As expected, the probability of rejecting a true null hypothesis ranges from 0.8 to 1.0 for the two tests.

\begin{table}[h!]
\centering
\begin{tabular}{|l | c | c | c |}
\hline\hline
& \textbf{Power = 40/60} & \textbf{Power = 60/80} & \textbf{Power = 40/80} \\
\hline
$\theta =$ \textbf{0.2} & & & \\
\hspace{3mm} McNemar's & 0.808 (0.01) & 0.873 (0.01) & 0.999 (0.0003) \\
\hspace{3mm} T test & 0.849 (0.01) & 0.825 (0.01) & 0.999 (0.0002)\\
\hline
$\theta =$ \textbf{0.5} & & & \\
\hspace{3mm} McNemar's & 0.808 (0.01) & 0.874 (0.01) & 0.999 (0.0002)  \\
\hspace{3mm} T test & 0.851 (0.01) & 0.826 (0.01) & 0.999 (0.0002) \\
\hline
$\theta =$ \textbf{0.8} & & & \\
\hspace{3mm} McNemar's & 0.804 (0.01) & 0.874 (0.01) & 0.999 (0.0001) \\
\hspace{3mm} T test & 0.851 (0.01) & 0.824 (0.01) & 1.000 (0.0000)\\
\hline\hline
\end{tabular}
\caption{Probability of rejecting a true null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in  parentheses and rounded to the first significant digit.}
\end{table}

In data simulations to test our analysis of power of the test, we fix the power of the original and replication studies at either 40%, 60%, or 80%. For each of these power specifications, we then consider the case where (1) the original study has a small effect size and the replication study has a medium effect size, (2) the original study has a medium effect size and the replication study has a large effect size, and (3) the original study has a small effect size and the replication study has a large effect size.  For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the \textit{t} test or McNemar's test reject the null hypothesis in favor of the alternative. The results of these simulations are displayed in Table 2. As expected, the probability of rejecting a false null hypothesis is approximately 0.05 for both the \textit{t} test and McNemar's test. That is, the Type II error rate for both tests is 95%.

\begin{table}[h!]
\centering
\begin{tabular}{|l | c | c | c |}
\hline\hline
& $\theta =$ \textbf{0.2/0.5} & $\theta=$ \textbf{0.5/0.8} & $\theta=$ \textbf{0.2/0.8} \\
\hline
\textbf{40\%\ Power} & & & \\
\hspace{3mm} McNemar's & 0.049 (0.007) & 0.050 (0.007) & 0.049 (0.007) \\
\hspace{3mm} T test & 0.050 (0.007) & 0.051 (0.007) & 0.050 (0.007)\\
\hline
\textbf{60\%\ Power} & & & \\
\hspace{3mm} McNemar's & 0.050 (0.007) & 0.050 (0.007) & 0.049 (0.007) \\
\hspace{3mm} T test & 0.050 (0.006) & 0.049(0.008) & 0.049 (0.006)\\
\hline
\textbf{80\%\ Power} & & & \\
\hspace{3mm} McNemar's & 0.051 (0.008) & 0.051 (0.007) & 0.051 (0.008) \\
\hspace{3mm} T test &  0.048 (0.008) & 0.048 (0.007) & 0.048 (0.007)\\
\hline\hline
\end{tabular}
\caption{Probability of rejecting a false null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in paraentheses and rounded to the first significant digit.}
\end{table}

### 2018-06-21 Fishers_method/Fishers_method_short.Rmd
Edits to shortened write-up, cleaned power_sims code

@@
 author: "Katie Fitzgerald"
-date: "06/18/2018"
+date: "06/20/2018"
 output: pdf_document
@@
 Among the 100 replication studies conducted by the Open Science Collaboration (OSC), 64 found null effects, defined as having a p-value greater than 0.05.
-The OSC authors applied Fisher's method to this set of non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
-Therefore, their hypothesis for the Fisher's test could be formalized as follows, where $\theta_i$ is the treatment effect for study $i$:
+The OSC applied Fisher's method to this set of non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
+Therefore, their hypothesis for Fisher's test could be formalized as follows, where $\theta_i$ is the treatment effect for study $i$:
 \begin{center}
@@
 
-Because the OSC authors are conducting Fisher's test only among the 64 replication studies that did not have significant results and thus $p_i \geq 0.05$ for all $i$, they use the following transformation of Fisher's test statistic  $$X^2=-2\sum_{i=1}^{k} ln(p_i^*)=-2\sum_{i=1}^{k} ln(\frac{p_i-0.05}{0.95}),$$ where under $H_0$, the $p_i^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$.
+Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results and thus $p_i \geq 0.05$ for all $i$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{i=1}^{k} ln(p_i^*)=-2\sum_{i=1}^{k} ln(\frac{p_i-0.05}{0.95}),$$ where under $H_0$, the $p_i^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies, in this case 64.
 Low p-values give a larger test statistic, leading to a rejection of $H_0$.
@@
 
-Even though the OSC was able to reject their null hypothesis ($X^2=155.83, p=0.048$), we think that in general Fisher's method is underpowered to answer the question at hand, so we want to explore this more rigorously.
-The intuition is that if there is just one $\theta_i\neq0$, and the remaining $k-1$ studies have p-values from a U[0,1] distribution, the one non-null effect will need to be very large to skew the distribution enough to look differently from U[0,1].
-Or alternatively, there would need to be several non-null effects to skew the distribution to a detectable degree.
-In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall $\Theta=0$, this type of conservative test may be appropriate.
+Even though the OSC was able to reject their null hypothesis ($X^2=155.83, p=0.048$), we think that in general Fisher's method is underpowered to answer the question at hand.
+We hypothesize that it requires many and possibly large non-null effects in order to skew the distribution of p-values enough to reject Fisher's null hypothesis.
+In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall treatment effect $\Theta=0$, this type of conservative test may be appropriate.
 In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_i$ has no bearing on the 63 other $\theta_i's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
-We will use power simulations to determine if the intuition holds that it would take many and possibly large non-null effects for Fisher's test to reject $H_0$.
+We turn to simulations to investigate the power of Fisher's method in the OSC scenario.
 
-For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma_i}$, where $\theta_i$ is the mean difference between the treatment and control groups in study $i$ and $\sigma_i$ is the constant variance among the treatment and control populations.
-Assuming equal sample sizes in the treatment and control groups within study $i$ (i.e.
-$n_i^t=n_i^c=n_i$), $\delta_i$ is estimated by $d_i\sim N(\delta_i,\frac{2}{n_i})$ (See Appendix for proof and discussion of simplifying assumptions).
-Under this framework, Fisher's method can be represented as testing the hypotheses
+For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma_i}$, where $\theta_i$ is the mean difference between the treatment and control groups in study $i$ and $\sigma_i$ is the known and equal variance among the treatment and control populations.
+Assuming equal sample sizes in the treatment and control groups within study $i$ (that is, let $n_i^t=n_i^c=n_i$), $\delta_i$ is estimated by $d_i\sim N(\delta_i,\frac{2}{n_i})$.[^1] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
+
+[^1]:See Appendix A.1 for proof and discussion of simplifying assumptions.
 
@@
 
-$H_a$: at least one $\theta_i\neq 0,\quad i=1,...,64$
-\end{center} and the 64 p-values can be calculated as $p_i=2(1-\Phi(\frac{|d_i|}{\sqrt{2/n_i}})).$
+$H_a$: at least one $\delta_i\neq 0,\quad i=1,...,64,$
+\end{center} 
+and the 64 p-values to be summed in Fisher's test statistic can be calculated as $p_i=2(1-\Phi(\frac{|d_i|}{\sqrt{2/n_i}})).$
 
 We first consider the power of Fisher's method when $n_i=76/2=38$ for all $i$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
-The results of the power simulations for median sample size are given in TableXX (See Appendix for code and details on how the power simulations were conducted).
-Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
-We find that there needs to be 10 large effects in order to achieve close to the standard 80% power (power=0.77426).
+The results of the power simulations for median sample size are given in Table 1.[^2] Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
+As shown in the last column of Table 1, we find that there needs to be 10 large effects in order to achieve close to the standard 80% power (power=0.7743).
 If there is only one large effect, Fisher's test only has approximately 8% power to reject.
 Even when all 64 studies have small effects, Fisher's method only has 70% power to reject.
+
+[^2]: See Appendix A.2 for code and details on how the power simulations were conducted.
+
+Table: Power of Fisher's method given median sample size $(n_i=38)$ for varying $\delta$ and true # of non-null effects
 ```{r,echo=FALSE}
@@
 In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
-That is, if there is just one non-null effect we let it be from the largest study, if there are two non-null effects we let them be from the two largest studies, etc.
-Note, however that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
-The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ were $n_i=745; 159;$ and $100$ respectively.
-See TableXX in the Appendix for the full sample size vectors that were used to conduct the "best-case scenario" power simulations.
+That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^3] As shown in the first row of Table 2, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
+When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340).
 
-As shown in the first row of TableXX, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
-When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340)
-
+Table: Power of Fisher's method given large sample sizes for varying $\delta$ and true # of non-null effects (i.e. "Best-case scenario")
 ```{r, echo=FALSE}
@@
 ```
+[^3]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
+The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ were $n_i=745; 159;$ and $100$ respectively.
+See Table 4 in Appendix A.2 for the full sample size vectors that were used to conduct these "best-case scenario" power simulations.
 
 ##Asymptotic power
-If all 64 have 0.2, need n=278 (total sample size) for 80% power, n=45 for d=0.5, n=18 for d=0.8.
 
-##Appendix
+#Appendices
 
-Assume each study $i$ is testing the presence of some treatment effect $\theta_i$ and has equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$).
-Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($j=1,...n_i$), and assume $Y_{ij}^t\sim N(\mu_i +\theta_i, \sigma^2)$ and $Y_{ij}^c\sim N(\mu_i, \sigma^2)$, where $\sigma^2$ is known.
-Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in this subset of 64 replicate studies are of these two types (89%).
+##Appendix A: Fisher's method power simulations
 
-Note then, $\theta_i$ is the difference in means between the treatment and control groups, and its estimate $T_i=\overline{Y_{i.}^t}-\overline{Y_{i.}^c}$ has variance $v_i=\frac{\sigma^2}{n_{ci}} +\frac{\sigma^2}{n_{ti}}=\frac{2\sigma^2}{n_i}.$ For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma}$.
+###A.1 Framework and assumptions
+
+Assume each study $i$ is testing the presence of some treatment effect $\theta_i$, where $i=1,...,k$ and $k$ is the number of studies.
+Assume equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$).
+Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($i=1,...,k$, and $j=1,...n_i$), and assume $Y_{ij}^t\sim N(\mu_i +\theta_i, \sigma_i^2)$ and $Y_{ij}^c\sim N(\mu_i, \sigma_i^2)$, where the $\sigma_i^2$'s are known.
+Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in the OSC subset of 64 replicate studies are of these two types (89%).
+
+Note then, $\theta_i$ is the difference in means between the treatment and control groups, and its estimate $T_i=\overline{Y_{i.}^t}-\overline{Y_{i.}^c}$ has variance $v_i=\frac{\sigma^2}{n_{ci}} +\frac{\sigma^2}{n_{ti}}=\frac{2\sigma^2}{n_i}.$ For simplicity of interpretation but without loss of generality, we work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma}$.
 Note then that $\delta_i$ is estimated by $d_i=\frac{T_i}{\sigma}=\frac{T_i}{\sqrt{\frac{v_i}{2/n_i}}}=\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}}.$ Since under $H_0$, $T_i\sim N(\theta_i,v_i) \Rightarrow \frac{T_i}{\sqrt{v_i}} \sim N(\frac{\theta_i}{\sqrt{v_i}},1)$ and $\sqrt{\frac{2}{n_i}}$ is a constant, then we have  $Var(d_i)=Var(\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}})=\frac{2}{n_i}$.
@@
 
-In our simulations, we will assume that there are $m$ false negatives among the 64 studies, $m=1,...,64$.
+###A.2 Power simulation logic and code
+Let there be $m$ false negatives among the 64 studies, $m=1,...,64$, that is $m$ true non-null effects.
 Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
@@
 We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{l=0}^{N}I_{\{X^2_l>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.
+We let N=100,000.
+
+Table: Power of OSC replicate studies to detect $\delta$ given $n$
 ```{r,echo=FALSE,warning=FALSE,message=FALSE}
 #test upper bounds of power (by largest sample size study having false negatives)
+
+#read in OSC data
 rpp_data<-read.csv("~/Desktop/NU/Replication/replication_metrics/data/rpp_data.csv")
+
+#subset only the 64 studies used in Fisher's method and extract sample sizes
 data_forFisher<- rpp_data[ !is.na(rpp_data$T_pval_USE..R.) & (rpp_data$T_pval_USE..R.>0.05)  , ]
@@
 
+#calculate power for OSC replicate studies to detect effects of 0.2, 0.5, and 0.8 given their exact sample size
 source("~/Desktop/NU/Replication/replication_metrics/src/metrics_funs.R")
-replicate_power<-data.frame(n,power_delta_0.2=power_fun(0.2,2/n),power_delta_0.5=power_fun(0.5,2/n),power_delta_0.8=power_fun(0.8,2/n))
+replicate_power<-round(data.frame(n,power_delta_0.2=power_fun(0.2,2/n),power_delta_0.5=power_fun(0.5,2/n),power_delta_0.8=power_fun(0.8,2/n)),4)
 
+#shorten # of rows to display in table
 replicate_power<-rbind(replicate_power[1:12,],c("...","...","...","..."))
 library(knitr)
-kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("n","d=0.2","d=0.5","d=0.8"),caption="Power to reject H0 with median sample size (n=38)")
-
+kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("n","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power of OSC replicate studies to detect $\\delta$ given $\n$")
+```
+Table: Sample size vectors used in "best case scenario" power simulations for given $\delta$'s
+```{r, echo=FALSE,warning=FALSE,message=FALSE}
 large_n_used<-readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/large_n_used.RDS")
@@
 rownames(large_n_table)[14]<-c("...")
-kable(large_n_table,format="markdown",digits=1,col.names=c("d=0.2","d=0.5","d=0.8"))
+kable(large_n_table,format="markdown",digits=1,col.names=c("$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"))
 
 ```
+
+```{r power_sims}
+power_sims<-function(N,M,delta,n){
+  #########################################################################################
+  # TAKES: N; number of simulations 
+  #        M; vector of number of non-null effects
+  #        delta; effect size under alternative hypothesis, on scale of cohen's d
+  #        n; vector of treatment/control sample size across studies (total sample size/2)
+  # RETURNS: power of Fisher's test to reject
+  # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
+  #########################################################################################
+
+  T<-c() #empty list to store Fisher's test statistic 
+  Power<-matrix() #empty matrix to store results
+
+  for(k in 1:length(M)){
+
+    for (i in 1:N){
+
+      #print(i) # can uncomment to show progress for lengthy simulations
+
+      p0<-runif(64 - M[k], 0.05, 1) #draws p-values for the true null effects
+
+      p1<-c() #create list to store p-values drawn for non-null effects
+
+      for (j in 1:M[k]){
+
+        p1[j]<-0
+
+        while (p1[j] <= 0.05) { #throw away p-values<=0.05
+          p1[j]<-2 * (1 - pnorm(abs(rnorm(1, delta, sqrt(2 / n[j]))) / sqrt(2 / n[j])))
+        }
+
+        #print(n[j]) # can uncomment to show progress for lengthy simulations 
+        }
+
+      #test statistic for Fisher's method, with transformation for truncating p-values
+      T[i]<--2 * sum(log((p0 - 0.05)/0.95)) - 2 * sum(log((p1 - 0.05) / 0.95))
+
+      }
+
+    Power[k]<-sum(T > 155.4047) / N
+
+    }
+
+    return(Power)
+}
+```

### 2018-06-21 Fishers_method/power_sims.R

@@ -7,26 +7,41 @@ power_sims<-function(N,M,delta,n){
   # RETURNS: power of Fisher's test to reject
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
   #########################################################################################
-  T<-c() #empty list for Fisher's test statistic 
+  
+  T<-c() #empty list to store Fisher's test statistic 
   Power<-matrix() #empty matrix to store results
+  
   for(k in 1:length(M)){
+    
     for (i in 1:N){
-      print(i) #uncomment to show progress for lengthy simulations
-      p0<-runif(64-M[k],0.05,1) #draws p-values for the true negatives
-      p1<-c() 
-      for (j in 1:M[k]){
+      
+      #print(i) # can uncomment to show progress for lengthy simulations
+    
+      p0<-runif(64 - M[k], 0.05, 1) #draws p-values for the true null effects
+      
+      p1<-c() #create list to store p-values drawn for non-null effects
+      
+      for (j in 1:M[k]){ 
+        
         p1[j]<-0
-        while (p1[j]<=0.05) { #throw away p-values<=0.05
-          p1[j]<-2*(1-pnorm(abs(rnorm(1,delta,sqrt(2/n[j])))/sqrt(2/n[j])))
+        
+        while (p1[j] <= 0.05) { #throw away p-values<=0.05
+          p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n[j])))/sqrt(2/n[j])))
         }
-        print(n[j]) #uncomment to show progress for lengthy simulations 
-      }
+        
+        #print(n[j]) # can uncomment to show progress for lengthy simulations 
+        }
+      
       #test statistic for Fisher's method, with transformation for truncating p-values
       T[i]<--2*sum(log((p0-0.05)/0.95))-2*sum(log((p1-0.05)/0.95)) 
-    }
+    
+      }
+    
     Power[k]<-sum(T>155.4047)/N
-  }
-  return(Power)
+  
+    }
+  
+    return(Power)
 }
 
 M<-c(1,2,3,4,5,10,32,64) #vector of number of non-null effects

### 2018-06-21 effect sizes/effect_sizes.Rmd
fixed bad merge
changes to ES analysis

@@
 They determined that studies replicated if the estimate from the initial experiment $T_1$ is contained in a 95\% confidence interval for the replication effect parameter $\theta_2$. 
-The logic behind this can be seen in two different ways. 
+The logic behind this can be seen in at least two different ways. 
 On the one hand, it can be conceived of as determining if the effect parameter in the replicate $\theta_2$ is different from the estimate of the initial study $T_1$. 
@@
 \[
-S = |T_1 - T_2|/\sqrt{v_2} \leq 1.96
+|S| = |T_1 - T_2|/\sqrt{v_2} \leq 1.96
 \]
@@
 If study 2 is twice the size of study 1, so that $v_1 = v_2/2$, then the error rate is 51.4\%. 
+In other words, even if studies replicate exactly, this test may be very likely to determine that they do not.
 
-[INSERT FIGURE? error rate vs. v1/v2]
+# Weighted Average
 
-If, however, $\theta_1 \neq \theta_2$ we would want the test to determine that the studies do not replicate, so that $S$ is at least as large as 1.96 in magnitude.
+Another consideration of the OSF analyses was not whether study results were the same, but rather if taken in concert they provided convincing evidence of a nonzero effect. 
+The authors determined that an effect was nonzero if the precision weighted average---a fixed effects meta-analysis---of the two estimates was statistically significant.
+That is, they computed 
+\[
+\bar{T}_{\cdot} = \frac{T_1/v_1 + T_2/v_2}{1/v_1 + 1/v_2}
+\]
+Under the model, $\bar{T}_{\cdot}$ is normally distributed with mean 
+\[
+\bar{\theta}_{\cdot} = \frac{\theta_1/v_1 + \theta_2/v_2}{1/v_1 + 1/v_2}
+\]
+and variance 
+\[
+\frac{1}{1/v_1 + 1/v_2}
+\]
+
+When the effect parameters are approximately the same, so that $\theta_1 \approx \theta_2$, then this is a sensible approach, and it provides a valid test of wehther the effect is nonzero.
+However, if $\theta_1 \neq \theta_2$, then inferences will pertain to their average, $\bar{\theta}_{\cdot}$.
+It can be unclear what this average means scientifically in the context of only two studies, which is part of the reason why in the meta-analysis literature, it is recommended to present study results individually instead of an average.
+
+It is worth noting that this approach will tend to align with some of the determinations made about whether a finding was replicated.
+If the studies correspond in sign and statistical significance, then their meta-analytic average will necessarily be significant.
+However, it is possible for both estimates to be nonsignificant but still obtain a statistically significant average.

### 2018-06-25 draft_mcnemars_ttest.Rmd
appendix added with some light explainations of the tables.

--- 2018-06-20 McNemar's Test/draft_mcnemars_ttest.Rmd
@@
 The null hypothesis for McNemar's test states that the marginal probabilities for each outcome are the same.
-That is, $H_0: p_{11} + p_{10} = p_{11} + p_{01}$ where the $P$'s are the probabilities of occurrence in cells with the corresponding labels.
+That is, $H_0: p_{11} + p_{10} = p_{11} + p_{01}$ where the $p$'s are the probabilities of occurrence in cells with the corresponding labels.
 The test statistic is given by $\chi^2 = \frac{(n_{10} - n_{01})^2}{n_{10} + n_{01}}$.
@@
 The power of each of the $2k$ original and replication studies is determined by the true effect size $(\theta_{ij})$, the sample size $(n_{ij})$, and the population variance $(\sigma^2)$.
-If the original and replication studies have the same true effect size and population variance, we can let their powers differ by increasing the sample size of the replication studies.
+If the original and replication studies ahve the same true effect size and population variance, we can let their powers differ by increasing the sample size of the replication studies.
 As $n_{2j}$ grows large, $\frac{|T_{2j}|}{\sqrt{v_{2j}}}$ will also grow large.
@@
 Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$.
-The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function..
+The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$.
 Assuming that the variance is known for such a study and is $\frac{4}{n}$ where $n$ is the sample size then, the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$.
@@
 
- Given that this mean takes into account both the estimate and sample size of the original and replication studies, the power to detect a difference in \textit{p}-values depends on the power of both the original and replication studies.
+where $\Phi^{-1}$ is the standard normal quantile function.
+Given that this mean takes into account both the estimate and sample size of the original and replication studies, the power to detect a different in \textit{p}-values depends on the power of both the original and replication studies.
 
@@
 In a simple thought experiment, we consider a single study pair.
-The t-test used to compare the original and replication findings in this pair would have a test statistic of $\frac{{p_1-p_2}}{\sqrt{V_{p_1-p_2}}}$.
+The t-test to compare the original and replication findings in this pair would have a test statistic of $\frac{{p_1-p_2}}{\sqrt{V_{p_1-p_2}}}$.
 If we fix the parameters of the original study and let the sample size of the replication studies tend to $\infty$, the variance of the replication estimate will tend to 0 and, as a result, $p_2$ will also tend to 0.
@@
 These power specifications are consistent with the range observed in the OSF data.
-For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the t-test or McNemar's test reject the null hypothesis in favor of the alternative.
+For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the t-test or McNemar's test reject the null hypothesis in favor of the alternatie.
 The results of these simulations are displayed in Table 1.
@@
 \end{tabular}
-\caption{Probability of rejecting a true null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in  parentheses and rounded to the first significant digit.}
+\caption{Probability of rejecting a true null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in  parentheses and rounded to one significant digit.}
 \end{table}
@@
 \hspace{3mm} McNemar's & 0.051 (0.008) & 0.051 (0.007) & 0.051 (0.008) \\
-\hspace{3mm} T test &  0.048 (0.008) & 0.048 (0.007) & 0.048 (0.007)\\
+\hspace{3mm} T test &  0.048 (0.008) & 0.048(.007) & 0.048 (0.007)\\
 \hline\hline

### 2018-06-25 Fishers_method/Fishers_method_short.Rmd
edits to include sentences about asymptotic results and clean up appendices

@@
 knitr::opts_chunk$set(echo = TRUE)
-#setwd("~/Desktop/NU/Replication/replication_metrics")
 ```
@@
 
-Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results and thus $p_i \geq 0.05$ for all $i$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{i=1}^{k} ln(p_i^*)=-2\sum_{i=1}^{k} ln(\frac{p_i-0.05}{0.95}),$$ where under $H_0$, the $p_i^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies, in this case 64.
+Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results, and thus $p_i \geq 0.05$ for all $i$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{i=1}^{k} ln(p_i^*)=-2\sum_{i=1}^{k} ln(\frac{p_i-0.05}{0.95}),$$ where under $H_0$, the $p_i^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies (in this case $k=64$).
 Low p-values give a larger test statistic, leading to a rejection of $H_0$.
@@
 In the event that you do reject $H_0$, this result only tells you that at least one study was a false negative, but it does not till you *which* study did in fact have a true effect.
-Even still, it provides no information about the size or direction of that true effect and whether or not it replicates the original finding.
+Even still, it provides no information about the size or direction of that true non-zero effect and whether or not it replicates the original finding.
 Presumably, a finding of "no false negatives" would be most informative in assessing replication in this scenario, but this can never be validly concluded from Fisher's method since that would require concluding the null hypothesis.
@@
 In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_i$ has no bearing on the 63 other $\theta_i's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
-We turn to simulations to investigate the power of Fisher's method in the OSC scenario.
 
-For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma_i}$, where $\theta_i$ is the mean difference between the treatment and control groups in study $i$ and $\sigma_i$ is the known and equal variance among the treatment and control populations.
-Assuming equal sample sizes in the treatment and control groups within study $i$ (that is, let $n_i^t=n_i^c=n_i$), $\delta_i$ is estimated by $d_i\sim N(\delta_i,\frac{2}{n_i})$.[^1] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
+The true distribution of Fisher's test statistic under the alternative hypothesis is unknown, and therefore the power cannot be calculated exactly.
+The asymptotic distribution of $X^2$ can be shown to be approximately normal, but this approximation is not valid for the sample sizes in the OSC data.[^1] We therefore turn to simulations to investigate the power of Fisher's method in the OSC scenario.
 
-[^1]:See Appendix A.1 for proof and discussion of simplifying assumptions.
+[^1]: See Appendix B
+
+For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma_i}$, where $\theta_i$ is the mean difference between the treatment and control groups in study $i$, and $\sigma_i$ is the known and equal variance among the treatment and control populations.
+Assuming equal sample sizes in the treatment and control groups within study $i$ (that is, let $n_i^t=n_i^c=n_i$), $\delta_i$ is estimated by $d_i\sim N(\delta_i,\frac{2}{n_i})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
+
+[^2]:See Appendix A.1 for proof and discussion of simplifying assumptions.
 
@@
 We first consider the power of Fisher's method when $n_i=76/2=38$ for all $i$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
-The results of the power simulations for median sample size are given in Table 1.[^2] Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
+The results of the power simulations for median sample size are given in Table 1.[^3] Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
 As shown in the last column of Table 1, we find that there needs to be 10 large effects in order to achieve close to the standard 80% power (power=0.7743).
@@
 
-[^2]: See Appendix A.2 for code and details on how the power simulations were conducted.
+[^3]: See Appendix A.2 for code and details on how the power simulations were conducted.
 
@@
 library(knitr)
-table <- readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/power sims table with changing n.RDS")
+table <- readRDS("./power sims table with changing n.RDS")
 table_median<-table[,1:4]
@@
 In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
-That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^3] As shown in the first row of Table 2, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
+That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^4] As shown in the first row of Table 2, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
 When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340).
@@
 ```{r, echo=FALSE}
-table <- readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/power_large_n.RDS")
+table <- readRDS("./power_large_n.RDS")
 kable(table,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power to reject H0 with large n")
 ```
-[^3]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
+[^4]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
 The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ were $n_i=745; 159;$ and $100$ respectively.
-See Table 4 in Appendix A.2 for the full sample size vectors that were used to conduct these "best-case scenario" power simulations.
-
-##Asymptotic power
+See Appendix A.3 for further discussion.
 
@@
 
-Assume each study $i$ is testing the presence of some treatment effect $\theta_i$, where $i=1,...,k$ and $k$ is the number of studies.
+Assume each study $i$ is testing the presence of some treatment effect $\theta_i$, where $i=1,...,k$, and $k$ is the number of studies.
 Assume equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$).
@@
 ###A.2 Power simulation logic and code
-Let there be $m$ false negatives among the 64 studies, $m=1,...,64$, that is $m$ true non-null effects.
+Let there be $m$ false negatives (i.e.
+$m$ true non-zero effects) among the 64 studies, $m=1,...,64$.
 Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
@@
 We let N=100,000.
-
-Table: Power of OSC replicate studies to detect $\delta$ given $n$
-```{r,echo=FALSE,warning=FALSE,message=FALSE}
-#test upper bounds of power (by largest sample size study having false negatives)
-
-#read in OSC data
-rpp_data<-read.csv("~/Desktop/NU/Replication/replication_metrics/data/rpp_data.csv")
-
-#subset only the 64 studies used in Fisher's method and extract sample sizes
-data_forFisher<- rpp_data[ !is.na(rpp_data$T_pval_USE..R.) & (rpp_data$T_pval_USE..R.>0.05)  , ]
-n<-sort(data_forFisher$N..R.,decreasing=TRUE)/2
-
-#calculate power for OSC replicate studies to detect effects of 0.2, 0.5, and 0.8 given their exact sample size
-source("~/Desktop/NU/Replication/replication_metrics/src/metrics_funs.R")
-replicate_power<-round(data.frame(n,power_delta_0.2=power_fun(0.2,2/n),power_delta_0.5=power_fun(0.5,2/n),power_delta_0.8=power_fun(0.8,2/n)),4)
-
-#shorten # of rows to display in table
-replicate_power<-rbind(replicate_power[1:12,],c("...","...","...","..."))
-library(knitr)
-kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("n","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power of OSC replicate studies to detect $\\delta$ given $\n$")
-```
-Table: Sample size vectors used in "best case scenario" power simulations for given $\delta$'s
-```{r, echo=FALSE,warning=FALSE,message=FALSE}
-large_n_used<-readRDS("~/Desktop/NU/Replication/replication_metrics/Fishers_method/large_n_used.RDS")
-large_n_table<-rbind(large_n_used[1:13,],c("...","...","..."),large_n_used[64,])
-rownames(large_n_table)[14]<-c("...")
-kable(large_n_table,format="markdown",digits=1,col.names=c("$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"))
-
-```
+The code is given below.
 
@@
 power_sims<-function(N,M,delta,n){
-  #########################################################################################
+  ########################################################################################
   # TAKES: N; number of simulations 
-  #        M; vector of number of non-null effects
+  #        M; vector of number of non-null effects (e.g. M=c(1, 2, 3, 4, 5, 32, 64))
   #        delta; effect size under alternative hypothesis, on scale of cohen's d
@@
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
-  #########################################################################################
+  ########################################################################################
 
@@
 ```
+
+##A.3 "Best-case scenario"" power simulations
+
+As noted in footnote 4 of the text, the largest sample sizes were dropped out of necessity in the "best-case scenario" power simulations presented in Table 2.
+For example, the largest study had $n_i=384351.5$, which is powered at 100% to detect even a small $\delta=0.2$.
+Therefore, it would have been impossible for this study to result in a p-value greater than 0.05 if there was a true non-null effect.
+Table 3 presents the power of the OSC replicate studies to detect a given $\delta$ using the true sample sizes for the 64 studies.
+For each $\delta$, we began the power simulations with the largest $n$ for which the power was at most 0.9999.
+The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ therefore were $n_i=745; 159;$ and $100$ respectively.
+Table 4 presents the sample size vectors used for the "best-case scenario" power simulations.
+
+Table: Power of OSC replicate studies to detect $\delta$ given $n$
+```{r,echo=FALSE,warning=FALSE,message=FALSE}
+#test upper bounds of power (by largest sample size study having false negatives)
+
+#read in OSC data
+rpp_data<-read.csv("../data/rpp_data.csv")
+
+#subset only the 64 studies used in Fisher's method and extract sample sizes
+data_forFisher<- rpp_data[ !is.na(rpp_data$T_pval_USE..R.) & (rpp_data$T_pval_USE..R.>0.05)  , ]
+n<-sort(data_forFisher$N..R.,decreasing=TRUE)/2
+
+#calculate power for OSC replicate studies to detect effects of 0.2, 0.5, and 0.8 given their exact sample size
+source("../src/metrics_funs.R")
+replicate_power<-round(data.frame(n,power_delta_0.2=power_fun(0.2,2/n),power_delta_0.5=power_fun(0.5,2/n),power_delta_0.8=power_fun(0.8,2/n)),4)
+
+#shorten # of rows to display in table
+replicate_power<-rbind(replicate_power[1:12,],c("...","...","...","..."))
+library(knitr)
+kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("n","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power of OSC replicate studies to detect $\\delta$ given $\n$")
+```
+Table: Sample size vectors used in "best case scenario" power simulations for given $\delta$'s
+```{r, echo=FALSE,warning=FALSE,message=FALSE}
+large_n_used<-readRDS("./large_n_used.RDS")
+large_n_table<-rbind(large_n_used[1:13,],c("...","...","..."),large_n_used[64,])
+rownames(large_n_table)[14]<-c("...")
+kable(large_n_table,format="markdown",digits=1,col.names=c("$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"))
+
+```
+
+##Appendix B: Fisher's method asymptotic results

### 2018-06-25 McNemar's Test/draft_mcnemars_ttest.Rmd
Update - McNemar's & t

@@
+<<<<<<< HEAD
+---
+title: "Assessing Replication via P-values"
+author: "Sarah Peko-Spicer & Mena Whalen"
+date: "June 19, 2018"
+output: pdf_document
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+Two of the methods proposed by the Open Science Collaboration (2015) for assessing replication via p-values are McNemar's test and the t-test for paired data.
+We discuss these tests in tandem as they are both extensions on the sign test and, as a result, have similar properties and agreement of significant results in the context of replication.
+For both tests, we consider $k$ pairs of studies, each containing an original study and a replication study.
+As described earlier, the original studies are characterized by an estimate $T_{1j} \sim N(\theta_{1j}, v_{1j})$ for $j = 1, \ldots, k$.
+Similarly, the replication studies are characterized by an estimate $T_{2j} \sim N(\theta_{2j}, v_{2j})$ for $j= 1, \ldots, k$.
+Further, we let $p_{1j}$ denote the \textit{p}-value of the j\textsuperscript{th} original study and $p_{2j}$ denote the \textit{p}-value of the j\textsuperscript{th} replication study.
+
+McNemar's test is used to determine if there is a difference between the proportion of statistically significant results among original studies and the proportion of statistically significant results among replication studies.
+Each \textit{p}-value from an original or replication study can be categorized as either significant or non-significant, creating a dichotomous variable suitable for the use of this test.
+This information is generally summarized in a 2 $\times$ 2 contingency table as shown below where $n_{11}$ denotes the number of study pairs in which both the original and replication findings are signficant, $n_{10}$ denotes the number of study pairs in which the original finding is significant but the replication finding is not, $n_{01}$ denotes the number of study pairs in which the original finding is not significant but the replication finding is, and $n_{00}$ denotes the number of study pairs in which neither the original finding nor the replication finding is significant.
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{l  c  c } \\
+& \multicolumn{2}{c}{Replication Finding} \\
+\cline{2-3}
+Original Finding & Significant & Non-Significant \\
+\hline
+Significant & $n_{11}$ & $n_{10}$ \\
+Non-Significant & $n_{01}$ & $n_{00}$ \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+The null hypothesis for McNemar's test states that the marginal probabilities for each outcome are the same.
+That is, $H_0: p_{11} + p_{10} = p_{11} + p_{01}$ where the $p$'s are the probabilities of occurrence in cells with the corresponding labels.
+The test statistic is given by $\chi^2 = \frac{(n_{10} - n_{01})^2}{n_{10} + n_{01}}$.
+Under the null hypothesis, the test statistic follow a chi-squared distribution with one degree of freedom.
+The null hypothesis is reject for large values of the test statistic.
+The results of the OSC article are summarized in the contingency table below.
+97% of original findings were significant compared to 36% of replication findings.
+The authors found a test statistic of $X^2 = 59.06$, rejected the null hypothesis, and concluded that the findings from the original studies did not replicate.
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{l  c  c } \\
+& \multicolumn{2}{c}{Replication Finding} \\
+\cline{2-3}
+Original Finding & Significant & Non-Significant \\
+\hline
+Significant & $35$ & $62$ \\
+Non-Significant & $1$ & $2$ \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Unlike McNemar's test, the t-test assesses replication by directly comparing original and replication \textit{p}-values within a study pair.
+The null hypothesis for the t-test states that the original and replication studies have the same \textit{p}-value (i.e.
+$H_0: p_{1j} = p_{2j}$).
+The test statistic is given by $t = \frac{\bar{X}_{diff}}{SD_{diff}/\sqrt{k}}$ where $\bar{X}_{diff}$ is the mean of all differences between original and replication \textit{p}-values (i.e.
+$\bar{X}_{diff} = \frac{1}{k} \sum\limits_{j=1}^k (p_{1j} - p_{2j})$), $SD_{diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of study pairs.
+A paired t-test is used in this case since the original and replication studies in each study pair are conducting the same experiment, using the same methods, and drawing conclusions on the same phenomenon.
+
+In the OSC article, the authors used the t-test on $k=99$ studies.
+They omitted one study pair where the exact \textit{p}-value was not available for the original or replication study.
+For the 99 available study pairs, the authors found a mean difference, $\bar{X}_{diff} = -0.274$, indicating that the replication \textit{p}-values were larger than the original \textit{p}-values.
+The test statistic was $-8.207$ $(p < 0.0001)$.
+Thus, the authors concluded that the original and replication \textit{p}-values are significantly different and the original and replication findings are significantly different.
+
+It should be noted that both the t-test and McNemar's test are used to assess similarity between a \textit{group} of original studies and a \textit{group} of replication studies.
+We assume that an original and replication study within a study pair are conducting the same experiment, using the same methods, to draw conclusions about the same phenomenon.
+However, in the OSC data we observe some variation between study pairs on these factors.
+That is, the 100 original studies may be conducting different experiments, using different methods, to draw conclusions about different phenomena.
+On these grounds alone, the t-test and McNemar's test are not well suited to assess whether an original finding has been replicated unless all $k$ study pairs are drawing conclusions about the same phenomenon.
+In the analysis that follows, we assume that this is, in fact, the case.
+
+## Analysis of McNemar's Test
+
+Two properties you might use to assess the quality of a statistical test for the purpose of replication are its \textit{false invalidation} and \textit{false validation} rates.
+These rates are similar in nature to Type I and Type II error rates, but use a different null hypothesis as a point of reference.
+Assuming that the null hypothesis for replication is given by $H_0: \theta_{1j} = \theta_{2j}$ and that the null hypotheses for the tests used in the OSC article are considered an "approximation" of this null, a false invalidation occurs when $\theta_{1j} = \theta_{2j}$ but the chosen statistical test rejects its null hypothesis.
+Similarly, a false validation occurs when $\theta_{1j} \neq \theta_{2j}$ but the chosen statistical test fails to reject its null hypothesis.
+To better understand these properties for McNemar's test, we rewrite the null hypothesis to reflect its relationship to the power of the original and replication studies.
+First, note that that the null hypothesis can be rewritten as
+
+$$
+H_0: P(p_{1j} \leq 0.05) = P(p_{2j} \leq 0.05)
+$$
+
+In other words, under the null hypothesis, if we were to draw a study pair at random, the probability of obtaining a significant result in the original study is equal to the probability of obtaining a significant result in the replication study.
+Taking this observation into consideration and noting that $p_{ij} \leq 0.05 \equiv \frac{|T_{ij}|}{\sqrt{v_{ij}}} \geq 1.96$, we can reformulate the null hypothesis as follows.
+
+$$
+H_0: \sum\limits_{j=1}^k P \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 |i \Big) P(i) = \sum\limits_{j=1}^k P \Big ( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96 |i \Big) P(i)
+$$
+
+That is, we can think of McNemar's test as a test of difference in the average power rather than a difference in paired proportions.
+As such, a rejection of the null hypothesis is akin to concluding that there is a statistically significant difference between the average power of the original studies and the average power of the replication studies.
+However, this does not necessarily imply that the original findings were not replicated in the replication studies, if we take replication to be $\theta_{1j} = \theta_{2j}$.
+Similarly, a failure to reject the null hypothesis for McNemar's test does not necessarily imply that the findings did, in fact, replicate.
+
+Assuming the true null hypothesis for replication is $H_0: \theta_{1j} = \theta_{2j}$, a false invalidation occurs when the original and replication studies differ in their power to detect their common effect size.
+In this case, McNemar's test will almost always reject its null hypothesis leading to conclusions of irreplicability despite having identical true effects.
+The power of each of the $2k$ original and replication studies is determined by the true effect size $(\theta_{ij})$, the sample size $(n_{ij})$, and the population variance $(\sigma^2)$.
+If the original and replication studies have the same true effect size and population variance, we can let their powers differ by increasing the sample size of the replication studies.
+As $n_{2j}$ grows large, $\frac{|T_{2j}|}{\sqrt{v_{2j}}}$ will also grow large.
+If we assume that original and replication studies are independent then,
+
+$$
+n_{10} = \sum\limits_{j=1}^k \mathbf{1}\Big\lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 \Big\rbrace \mathbf{1}\Big\lbrace \frac{|T_{2j}|}{\sqrt{v_{2j}}} < 1.96 \Big\rbrace \rightarrow 0
+$$
+
+$$
+n_{01} = \sum\limits_{j=1}^k \mathbf{1} \Big \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96  \Big \rbrace \mathbf{1} \Big \lbrace \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96  \Big \rbrace \rightarrow \sum\limits_{j=1}^k \mathbf{1} \Big \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \Big \rbrace
+$$
+This implies that the McNemar's test statistic $X^2$  $\rightarrow$ $\sum\limits_{j=1}^k \mathbf{1} \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \rbrace$ as the replication sample sizes grow large.
+Let's assume that all $k$ original studies have 80% power to detect their true effect.
+Then, $X^2$ approaches $0.2k$ as the replication sample sizes increase.
+Recall that McNemar's test rejects if $X^2 > \chi^2_{[0.05,1]}$.
+In this instance, as long as $k>2$, McNemar's test will always reject despite the true effect sizes being identical for all $k$ study pairs.
+
+A false validation occurs when the original and replication studies have different true effects (i.e.
+$\theta_{1j} \neq \theta_{2j}$) but have the same power to detect those effects (i.e.
+$(1-\beta_{1j}) = (1-\beta_{2j})$).
+The impact of the power of the original and replication studies on this test's false validation rate is easiest understood through a consideration of the test's power function.
+The power function for McNemar's test is given by
+
+$$
+\Phi \Big \lbrace \frac{(p_{10} - p_{01})\sqrt{k} - z_{1 - \alpha/2}\sqrt{p_{10} + p_{01}}}{\sqrt{p_{10} + p_{01} - (p_{10} - p_{01})^2} } \Big \rbrace
+$$
+where $\Phi$ is the standard normal distribution function and $p_{10}$ and $p_{01}$ can be written as $\sum\limits_{j=1}^k (1-\beta_{1j})\beta_{2j}$ and $\sum\limits_{j=1^k} \beta_{1j}(1-\beta_{2j})$, respectively.
+When the power of the original study is equal to the power of the replication study for all $k$ studies, $p_{10} = p_{01}$.
+The power of McNemar's test can then be reduced to $\Phi \lbrace -z_{1-\alpha/2} \rbrace$.
+Assuming that $z=1.96$, we know the power is 0.05.
+Thus, when the true effects of the original and replication studies are different, McNemar's test will only reject the null hypothesis 5% of the time.
+In other words, the false validation rate is roughly 95%.
+
+## Analysis of the t-test
+
+Like McNemar's test, the power of the t-test to detect if the \textit{p}-values of the original studies are different from those of the replication studies depends upon the power of both the original and replication studies.
+Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$.
+The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function..
+Assuming that the variance is known for such a study and is $\frac{4}{n}$ where $n$ is the sample size then, the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$.
+Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as
+
+$$
+\frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
+$$
+
+ Given that this mean takes into account both the estimate and sample size of the original and replication studies, the power to detect a difference in \textit{p}-values depends on the power of both the original and replication studies.
+
+Again, we consider the false invalidation and false validation for the t-test.
+A false invalidation error occurs when the effect sizes from the original and replication studies are equal but the t-test concludes that their \textit{p}-values are significant different from one another.
+This is likely to occur when the power of the original and replication studies differ.
+In a simple thought experiment, we consider a single study pair.
+The t-test used to compare the original and replication findings in this pair would have a test statistic of $\frac{{p_1-p_2}}{\sqrt{V_{p_1-p_2}}}$.
+If we fix the parameters of the original study and let the sample size of the replication studies tend to $\infty$, the variance of the replication estimate will tend to 0 and, as a result, $p_2$ will also tend to 0.
+This would not be unusual in a real-world appication as researchers often increase replication sample sizes to get a more precise estimate of the original finding.
+In this scenario, the t-test stastic depends only on the \textit{p}-value of the original study $\Big( \frac{p_1}{\sqrt{V_{p_1}}}\Big)$ and the error rate tends to 1.
+In other words, the t-test will rarely conclude that the two \textit{p}-values are similar to one another.
+Another way to think about this error rate is to consider the case where the original and replication test statistics differ (i.e.
+$\frac{T_1}{\sqrt{V_1}} \neq \frac{T_2}{\sqrt{V_2}}$).
+If the original and replication studies have different power to detect the same effect, the numerators of these statistics will be similar, but the sample sizes and variances will differ.
+As a result, the \textit{p}-values from the original and replication studies will be quite different, and once again the false invalidation rate will tend towards 0.80.
+
+A false validation occurs when $\theta_{1j} \neq \theta_{2j}$ but the powers are the same: $(1-\beta_{1j}) = (1-\beta_{2j})$.
+The power of the original and replication studies is determined by the effect size and the sample size.
+When the effect size is small, the sample size needs to be relatively large in order for the test to be well-powered.
+If two studies have different effect sizes but the same power, their test statistics can conceivably be identical (i.e.
+$\Big ( \frac{T_1}{\sqrt{V_1}} = \frac{T_2}{\sqrt{V_2}} \Big)$).
+For example, the study with the smaller effect size could have a larger sample size, reducing its variance while the study with the larger effect size could have a smaller sample size, resulting in a larger variance such that $cV_{small} = V_{big}$.
+Then, both test statistic would be similar enough to produce similar \textit{p}-values, even if their effect sizes are wildly different from one another.
+In that case, a t-test would fail to reject the null hypothesis despite the fact that the true effect sizes are different from one another.
+We posit that the power of this test will tend towards 0.05 (i.e.
+the false validation rate tends to 0.95).
+
+## Simulation Results
+
+In data simulations to test our analysis of the false invalidation rate, we fix the effect size of the original and replication studies at either a small, medium, or large effect.
+As is convention, we consider $\theta_{ij} = 0.2$ to be a small effect, $\theta_{ij} = 0.5$ to be a medium effect, and $\theta_{ij} = 0.8$ to be a large effect.
+For each of these effect size specifications, we then consider the case where (1) the original study is powered at 40% and the replication study is powered at 60%, (2) the original study is powered at 60% and the replication study is powered at 80%, and (3) the original study is powered at 40% and the replication study is powered at 80%.
+These power specifications are consistent with the range observed in the OSC data.
+For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the t-test or McNemar's test reject the null hypothesis in favor of the alternative.
+The results of these simulations are displayed in Table 1.
+As expected, the probability of rejecting a true null hypothesis ranges from 0.8 to 1.0 for the two tests.
+This probability does not vary for different values of $\theta$ but does vary depending on the difference in power.
+The false invalidation rate and the difference in powers are not linearly related, as evidenced by the different results between the first two columns.
+That is, we observe different false invalidation rates in the first two columns of Table 1 despite the difference in power between original and replication studies being identical across these two columns.
+That being said, as the disparity in power between the original and replication studies grows larger, the false invalidation rate tends towards 1.
+This result is consistent with the thought experiment presented in the analysis of the \textit{t} test.
+
+\newpage
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l | c | c | c |}
+\hline\hline
+& \textbf{Power = 40/60} & \textbf{Power = 60/80} & \textbf{Power = 40/80} \\
+\hline
+$\theta =$ \textbf{0.2} & & & \\
+\hspace{3mm} McNemar's & 0.808 (0.01) & 0.873 (0.01) & 0.999 (0.0003) \\
+\hspace{3mm} T test & 0.849 (0.01) & 0.825 (0.01) & 0.999 (0.0002)\\
+\hline
+$\theta =$ \textbf{0.5} & & & \\
+\hspace{3mm} McNemar's & 0.808 (0.01) & 0.874 (0.01) & 0.999 (0.0002)  \\
+\hspace{3mm} T test & 0.851 (0.01) & 0.826 (0.01) & 0.999 (0.0002) \\
+\hline
+$\theta =$ \textbf{0.8} & & & \\
+\hspace{3mm} McNemar's & 0.804 (0.01) & 0.874 (0.01) & 0.999 (0.0001) \\
+\hspace{3mm} T test & 0.851 (0.01) & 0.824 (0.01) & 1.000 (0.0000)\\
+\hline\hline
+\end{tabular}
+\caption{Probability of rejecting a true null hypothesis ($\theta_{1j} = \theta_{2j}$) for the \textit{t} test and McNemar's test.
+Standard errors are reported in  parentheses and rounded to the first significant digit.}
+\end{table}
+
+In data simulations to test our analysis of power of the test, we fix the power of the original and replication studies at either 40%, 60%, or 80%.
+For each of these power specifications, we then consider the case where (1) the original study has a small effect size and the replication study has a medium effect size, (2) the original study has a medium effect size and the replication study has a large effect size, and (3) the original study has a small effect size and the replication study has a large effect size.
+For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the \textit{t} test or McNemar's test reject the null hypothesis in favor of the alternative.
+The results of these simulations are displayed in Table 2.
+As expected, the probability of rejecting a false null hypothesis is approximately 0.05 for both the \textit{t} test and McNemar's test.
+That is, the false validation rate for both tests is 95%.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l | c | c | c |}
+\hline\hline
+& $\theta =$ \textbf{0.2/0.5} & $\theta=$ \textbf{0.5/0.8} & $\theta=$ \textbf{0.2/0.8} \\
+\hline
+\textbf{40\%\ Power} & & & \\
+\hspace{3mm} McNemar's & 0.049 (0.007) & 0.050 (0.007) & 0.049 (0.007) \\
+\hspace{3mm} T test & 0.050 (0.007) & 0.051 (0.007) & 0.050 (0.007)\\
+\hline
+\textbf{60\%\ Power} & & & \\
+\hspace{3mm} McNemar's & 0.050 (0.007) & 0.050 (0.007) & 0.049 (0.007) \\
+\hspace{3mm} T test & 0.050 (0.006) & 0.049(0.008) & 0.049 (0.006)\\
+\hline
+\textbf{80\%\ Power} & & & \\
+\hspace{3mm} McNemar's & 0.051 (0.008) & 0.051 (0.007) & 0.051 (0.008) \\
+\hspace{3mm} T test &  0.048 (0.008) & 0.048 (0.007) & 0.048 (0.007)\\
+\hline\hline
+\end{tabular}
+\caption{Probability of rejecting a false null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in paraentheses and rounded to the first significant digit.}
+\end{table}
+
+\newpage 
+## Appendix
+
+Given that McNemar's test and the \textit{t} test have similar properties, it is worth considering the rate at which the two tests agree on conclusions about replication.
+To determine whether McNemar's test and the \textit{t} test always reach the same conclusions when $\theta_{1j} = \theta{2j}$, we ran 1,000 trials on the same simulated data and calculated the proportion of trials in which both tests rejected their null hypothesis.
+The results of these simulations are found in Table 3.
+Similarly, to determine whether the two tests always reach the same conclusions when $\theta_{1j} \neq \theta_{2j}$, we ran 1,000 trials on the same simulated data and calucated the porpotion of trials in which both tests failed to reject their null hypothesis.
+The results of these simulations are found in Table 4.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l | c | c | c |}
+\hline\hline
+& \textbf{Power = 40/60} & \textbf{Power = 60/80} & \textbf{Power = 40/80} \\
+\hline
+$\theta =$ \textbf{0.2}  & 0.980 & 0.966 & 1.000 \\
+\hline
+$\theta =$ \textbf{0.5} & 0.994 & 0.980 & 1.000  \\
+\hline
+$\theta =$ \textbf{0.8} & 0.966 & 0.973 & 1.000 \\
+\hline\hline
+\end{tabular}
+\caption{Rate of agreement between McNemar's test and the \textit{t} test when $\theta_{1j} = \theta_{2j}$}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l | c | c | c |}
+\hline\hline
+& $\theta =$ \textbf{0.2/0.5} & $\theta=$ \textbf{0.5/0.8} & $\theta=$ \textbf{0.2/0.8} \\
+\hline
+\textbf{40\%\ Power} & 0.993 & 0.987 & 0.997 \\
+\hline
+\textbf{60\%\ Power} &  0.980 & 0.995 & 0.999 \\
+\hline
+\textbf{80\%\ Power} & 0.986 & 0.993 & 0.980 \\
+\hline\hline
+\end{tabular}
+\caption{Rate of agreement between McNemar's test and the \textit{t} test when $\theta_{1j} \neq \theta_{2j}$}
+\end{table}
+
+=======
 ---
@@
 \end{table}
+
+>>>>>>> 796555b595c5e5de3516f9efbf8d6204c04cdeaa

### 2018-06-25 McNemar's Test/mcnemar_sims.R

@@ -79,8 +79,9 @@ mc_case <- function(N, theta1, theta2, n1, n2, k){
 # will rarely reject (~7%) the null hypothesis in favor of the alternative despite the difference in
 # true effect size.
 
-# All studies have 80% power. Original studies have a true effect of 0.2 and replication studies
-# have a true effect of 0.8.
+
+# All studies have 80% power. Original studiesand replication studies differ in the size of 
+# their true effects.
 
 sl80 <- replicate(100,mc_case(1000, 0.2, 0.8, 392.4541, 24.52838, 100))
 sm80 <- replicate(100,mc_case(1000, 0.2, 0.5, 392.4541, 62.79265, 100))
@@ -90,8 +91,8 @@ mean(sl80); sd(sl80)
 mean(sm80); sd(sm80)
 mean(ml80); sd(ml80)
 
-# All studies have 60% power. Original studies have a true effect of 0.5 and replication studies 
-# have true effect of 0.8.
+# All studies have 60% power. Original studies and replication studies differ in the size of 
+# their true effects. 
 
 sl60 <- replicate(100, mc_case(1000, 0.2, 0.8, 244.9453, 15.309079, 100))
 sm60 <- replicate(100, mc_case(1000, 0.2, 0.5, 244.9453, 39.19124, 100))
@@ -101,8 +102,8 @@ mean(sl60); sd(sl60)
 mean(sm60); sd(sm60)
 mean(ml60); sd(ml60)
 
-# All studies have 40% power. Original studies have a true effect of 0.2 and replication studies
-# have a true effect of 0.8.
+# All studies have 40% power. Original studies and replication studies differ in the size of 
+# their true effects.
 
 sl40 <- replicate(100, mc_case(1000, 0.2, 0.8, 145.6332, 9.102075, 100))
 sm40 <- replicate(100, mc_case(1000, 0.2, 0.5, 145.6332, 23.30131, 100))

### 2018-06-25 McNemar's Test/mdes_script.R

@@ -1,0 +1,45 @@
+
+setwd(choose.dir())
+
+# Load the RPP data.
+
+rpp_dat <- read.csv("rpp_data.csv")
+
+# Take the meta-analytic subset of the data.
+
+meta_sub <- rpp_dat[is.na(rpp_dat$Meta.analytic.estimate..Fz.) == F,]
+
+# Convert the correlation coefficients into Cohen's d. 
+
+d.o <- (2*meta_sub$T_r..O.)/sqrt(1 - meta_sub$T_r..O.^2)
+d.r <- (2*meta_sub$T_r..R.)/sqrt(1 - meta_sub$T_r..R.^2)
+
+# Calculate the minimum detectable effect size given a certain power and the 
+# sample size. (Do we use the observed sample size? Or the sample size needed for
+# x% power?)
+
+mdes <- function(power, n){
+  (1.96 - qnorm(1-power))/sqrt(n)
+}
+
+# What is the MDES for replication studies @ 80% power?
+
+mdes(0.8, meta_sub$T_N..R.) # observed sample size
+mdes(0.8, as.numeric(meta_sub$X80..power)) # sample needed for 80% power
+
+# What is the MDES for original studies @ 60% power? @ 40% power?
+
+mdes(0.8, meta_sub$T_N..O.) # using the observed sample size
+mdes(0.6, meta_sub$T_N..O.) # using the observed sample size
+mdes(0.4, meta_sub$T_N..O.) # using the observed sample size
+
+# So, I suppose we want to know the difference between mdes for original and
+# replication studies. 
+
+mdes(0.8, meta_sub$T_N..O.) - mdes(0.8, meta_sub$T_N..R.)
+
+
+summary(meta_sub$Power..R.)
+
+
+

### 2018-06-25 McNemar's Test/draft_mcnemars_ttest.Rmd
updates

@@
-<<<<<<< HEAD
 ---
@@
 \end{table}
-
-=======
----
-title: "Assessing Replication via P-values"
-author: "Sarah Peko-Spicer & Mena Whalen"
-date: "June 19, 2018"
-output: pdf_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-Two of the methods proposed by the Open Science Foundation (2015) for assessing replication via p-values are McNemar's test and the t-test for paired data.
-We discuss these tests in tandem as they are both extensions on the sign test and, as a result, have similar properties in the context of replication.
-For both tests, we consider $k$ pairs of studies, each containing an original study and a replication study.
-As described earlier, the original studies are characterized by an estimate $T_{1j} \sim N(\theta_{1j}, v_{1j})$ for $j = 1, \ldots, k$.
-Similarly, the replication studies are characterized by an estimate $T_{2j} \sim N(\theta_{2j}, v_{2j})$ for $j= 1, \ldots, k$.
-Further, we let $p_{1j}$ denote the \textit{p}-value of the j\textsuperscript{th} original study and $p_{2j}$ denote the \textit{p}-value of the j\textsuperscript{th} replication study.
-
-McNemar's test is used to determine if there is a difference between the proportion of statistically significant results among original studies and the proportion of statistically significant results among replication studies.
-Each \textit{p}-value from an original or replication study can be categorized as either significant or non-significant, creating a dichotomous variable suitable for the use of this test.
-This information is generally summarized in a 2 $\times$ 2 contingency table as shown below where $n_{11}$ denotes the number of study pairs in which both the original and replication findings are signficant, $n_{10}$ denotes the number of study pairs in which the original finding is significant but the replication finding is not, $n_{01}$ denotes the number of study pairs in which the original finding is not significant but the replication finding is, and $n_{00}$ denotes the number of study pairs in which neither the original finding nor the replication finding is significant.
-
-\begin{table}[h!]
-\begin{center}
-\begin{tabular}{l  c  c } \\
-& \multicolumn{2}{c}{Replication Finding} \\
-\cline{2-3}
-Original Finding & Significant & Non-Significant \\
-\hline
-Significant & $n_{11}$ & $n_{10}$ \\
-Non-Significant & $n_{01}$ & $n_{00}$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The null hypothesis for McNemar's test states that the marginal probabilities for each outcome are the same.
-That is, $H_0: p_{11} + p_{10} = p_{11} + p_{01}$ where the $P$'s are the probabilities of occurrence in cells with the corresponding labels.
-The test statistic is given by $\chi^2 = \frac{(n_{10} - n_{01})^2}{n_{10} + n_{01}}$.
-Under the null hypothesis, the test statistic follow a chi-squared distribution with one degree of freedom.
-The null hypothesis is reject for large values of the test statistic.
-The results of the OSF article are summarized in the contingency table below.
-97% of original findings were significant compared to 36% of replication findings.
-The authors found a test statistic of $X^2 = 59.06$, rejected the null hypothesis, and concluded that the findings from the original studies did not replicate.
-
-\begin{table}[h!]
-\begin{center}
-\begin{tabular}{l  c  c } \\
-& \multicolumn{2}{c}{Replication Finding} \\
-\cline{2-3}
-Original Finding & Significant & Non-Significant \\
-\hline
-Significant & $35$ & $62$ \\
-Non-Significant & $1$ & $2$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Unlike McNemar's test, the t-test assesses replication by directly comparing original and replication \textit{p}-values within a study pair.
-The null hypothesis for the t-test states that the original and replication studies have the same \textit{p}-value (i.e.
-$H_0: p_{1j} = p_{2j}$).
-The test statistic is given by $t = \frac{\bar{X}_{diff}}{SD_{diff}/\sqrt{k}}$ where $\bar{X}_{diff}$ is the mean of all differences between original and replication \textit{p}-values (i.e.
-$\bar{X}_{diff} = \frac{1}{k} \sum\limits_{j=1}^k (p_{1j} - p_{2j})$), $SD_{diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of study pairs.
-A paired t-test is used in this case since the original and replication studies in each study pair are conducting the same experiment, using the same methods, and drawing conclusions on the same phenomenon.
-
-In the OSF article, the authors used the t-test on $k=99$ studies.
-They omitted one study pair where the exact \textit{p}-value was not available for the original or replication study.
-For the 99 available study pairs, the authors found a mean difference, $\bar{X}_{diff} = -0.274$, indicating that the replication \textit{p}-values were larger than the original \textit{p}-values.
-The test statistic was $-8.207$ $(p < 0.0001)$.
-Thus, the authors concluded that the original and replication \textit{p}-values are significantly different and the original and replication findings are significantly different.
-
-It should be noted that both the t-test and McNemar's test are used to assess similarity between a \textit{group} of original studies and a \textit{group} of replication studies.
-We assume that an original and replication study within a study pair are conducting the same experiment, using the same methods, to draw conclusions about the same phenomenon.
-However, in the OSF data we observe some variation between study pairs on these factors.
-That is, the 100 original studies may be conducting different experiments, using different methods, to draw conclusions about different phenomena.
-On these grounds alone, the t-test and McNemar's test are not well suited to assess whether an original finding has been replicated unless all $k$ study pairs are drawing conclusions about the same phenomenon.
-In the analysis that follows, we assume that this is, in fact, the case.
-
-## Analysis of McNemar's Test
-
-Two properties you might use to assess the quality of a statistical test are its Type I and Type II error rates.
-To better understand these properties for McNemar's test, we rewrite the null hypothesis to reflect its relationship to the power of the original and replication studies.
-First, note that that the null hypothesis can be rewritten as
-
-$$
-H_0: P(p_{1j} \leq 0.05) = P(p_{2j} \leq 0.05)
-$$
-
-In other words, under the null hypothesis, if we were to draw a study pair at random, the probability of obtaining a significant result in the original study is equal to the probability of obtaining a significant result in the replication study.
-Taking this observation into consideration and nothing that $p_{ij} \leq 0.05 \equiv \frac{|T_{ij}|}{\sqrt{v_{ij}}} \geq 1.96$, we can reformulate the null hypothesis as follows.
-
-$$
-H_0: \sum\limits_{j=1}^k P \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 |i \Big) P(i) = \sum\limits_{j=1}^k P \Big ( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96 |i \Big) P(i)
-$$
-
-That is, we can think of McNemar's test as a test of difference in the average power rather than a difference in paired proportions.
-As such, a rejection of the null hypothesis is akin to concluding that there is a statistically significant difference between the average power of the original studies and the average power of the replication studies.
-However, as we will show, this does not necessarily imply that the original findings were not replicated in the replication studies, if we take replication to be $\theta_{1j} = \theta_{2j}$.
-Similarly, we will show that a failure to reject the null hypothesis for McNemar's test does not necessarily imply that the findings did, in fact, replicate.
-
-Assuming the true null hypothesis for replication is $H_0: \theta_{1j} = \theta_{2j}$, a Type I error occurs when the original and replication studies differ in their power to detect their common effect size.
-In this case, McNemar's test will almost always reject its null hypothesis leading to conclusions of irreplicability despite having identical true effects.
-The power of each of the $2k$ original and replication studies is determined by the true effect size $(\theta_{ij})$, the sample size $(n_{ij})$, and the population variance $(\sigma^2)$.
-If the original and replication studies have the same true effect size and population variance, we can let their powers differ by increasing the sample size of the replication studies.
-As $n_{2j}$ grows large, $\frac{|T_{2j}|}{\sqrt{v_{2j}}}$ will also grow large.
-If we assume that original and replication studies are independent then,
-
-$$
-n_{10} = \sum\limits_{j=1}^k \mathbf{1}\Big\lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 \Big\rbrace \mathbf{1}\Big\lbrace \frac{|T_{2j}|}{\sqrt{v_{2j}}} < 1.96 \Big\rbrace \rightarrow 0
-$$
-
-$$
-n_{01} = \sum\limits_{j=1}^k \mathbf{1} \Big \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96  \Big \rbrace \mathbf{1} \Big \lbrace \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96  \Big \rbrace \rightarrow \sum\limits_{j=1}^k \mathbf{1} \Big \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \Big \rbrace
-$$
-This implies that the McNemar's test statistic $X^2$  $\rightarrow$ $\sum\limits_{j=1}^k \mathbf{1} \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \rbrace$ as the replication sample sizes grow large.
-Let's assume that all $k$ original studies have 80% power to detect their true effect.
-Then, $X^2$ approaches $0.2k$ as the replication sample sizes increase.
-Recall that McNemar's test rejects if $X^2 > \chi^2_{[0.05,1]}$.
-In this instance, as long as $k>2$, McNemar's test will always reject despite the true effect sizes being identical for all $k$ study pairs.
-
-A type II error occurs when the original and replication studies have different true effects (i.e.
-$\theta_{1j} \neq \theta_{2j}$) but have the same power to detect those effects (i.e.
-$(1-\beta_{1j}) = (1-\beta_{2j})$).
-The impact of the power of the original and replication studies on this test's Type II error rate is easiest understood through a consideration of the test's power function.
-The power function for McNemar's test is given by
-
-$$
-\Phi \Big \lbrace \frac{(p_{10} - p_{01})\sqrt{k} - z_{1 - \alpha/2}\sqrt{p_{10} + p_{01}}}{\sqrt{p_{10} + p_{01} - (p_{10} - p_{01})^2} } \Big \rbrace
-$$
-where $\Phi$ is the standard normal distribution function and $p_{10}$ and $p_{01}$ can be written as $\sum\limits_{j=1}^k (1-\beta_{1j})\beta_{2j}$ and $\sum\limits_{j=1^k} \beta_{1j}(1-\beta_{2j})$, respectively.
-When the power of the original study is equal to the power of the replication study for all $k$ studies, $p_{10} = p_{01}$.
-The power of McNemar's test can then be reduced to $\Phi \lbrace -z_{1-\alpha/2} \rbrace$.
-Assuming that $z=1.96$, we know the power is 0.05.
-Thus, when the true effects of the original and replication studies are different, McNemar's test will only reject the null hypothesis 5% of the time.
-In other words, the Type II error rate is roughly 95%.
-
-## Analysis of the t-test
-
-Like McNemar's test, the power of the t-test to detect if the \textit{p}-values of the original studies are different from those of the replication studies depends upon the power of both the original and replication studies.
-Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$.
-The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function..
-Assuming that the variance is known for such a study and is $\frac{4}{n}$ where $n$ is the sample size then, the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$.
-Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as
-
-$$
-\frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
-$$
-
- Given that this mean takes into account both the estimate and sample size of the original and replication studies, the power to detect a difference in \textit{p}-values depends on the power of both the original and replication studies.
-
-Again, we consider the Type I and Type II error for the t-test.
-A Type I error occurs when the effect sizes from the original and replication studies are equal but the t-test concludes that their \textit{p}-values are significant different from one another.
-This is likely to occur when the power of the original and replication studies differ.
-In a simple thought experiment, we consider a single study pair.
-The t-test used to compare the original and replication findings in this pair would have a test statistic of $\frac{{p_1-p_2}}{\sqrt{V_{p_1-p_2}}}$.
-If we fix the parameters of the original study and let the sample size of the replication studies tend to $\infty$, the variance of the replication estimate will tend to 0 and, as a result, $p_2$ will also tend to 0.
-This would not be unusual in a real-world appication as researchers often increase replication sample sizes to get a more precise estimate of the original finding.
-In this scenario, the t-test stastic depends only on the \textit{p}-value of the original study $\Big( \frac{p_1}{\sqrt{V_{p_1}}}\Big)$ and the error rate tends to 1.
-In other words, the t-test will rarely conclude that the two \textit{p}-values are similar to one another.
-Another way to think about this error rate is to consider the case where the original and replication test statistics differ (i.e.
-$\frac{T_1}{\sqrt{V_1}} \neq \frac{T_2}{\sqrt{V_2}}$).
-If the original and replication studies have different power to detect the same effect, the numerators of these statistics will be similar, but the sample sizes and variances will differ.
-As a result, the \textit{p}-values from the original and replication studies will be quite different, and once again the Type I error rate will tend towards 0.80.
-
-A Type II error occurs when $\theta_{1j} \neq \theta_{2j}$ but the powers are the same: $(1-\beta_{1j}) = (1-\beta_{2j})$.
-The power of the original and replication studies is determined by the effect size and the sample size.
-When the effect size is small, the sample size needs to be relatively large in order for the test to be well-powered.
-If two studies have different effect sizes but the same power, their test statistics can conceivably be identical (i.e.
-$\Big ( \frac{T_1}{\sqrt{V_1}} = \frac{T_2}{\sqrt{V_2}} \Big)$).
-For example, the study with the smaller effect size could have a larger sample size, reducing its variance while the study with the larger effect size could have a smaller sample size, resulting in a larger variance such that $cV_{small} = V_{big}$.
-Then, both test statistic would be similar enough to produce similar \textit{p}-values, even if their effect sizes are wildly different from one another.
-In that case, a t-test would fail to reject the null hypothesis despite the fact that the true effect sizes are different from one another.
-We posit that the power of this test will tend towards 0.05 (i.e.
-the Type II error rate tends to 0.95).
-
-## Simulation Results
-
-In data simulations to test our analysis of the Type I error rate, we fix the effect size of the original and replication studies at either a small, medium, or large effect.
-As is convention, we consider $\theta_{ij} = 0.2$ to be a small effect, $\theta_{ij} = 0.5$ to be a medium effect, and $\theta_{ij} = 0.8$ to be a large effect.
-For each of these effect size specifications, we then consider the case where (1) the original study is powered at 40% and the replication study is powered at 60%, (2) the original study is powered at 60% and the replication study is powered at 80%, and (3) the original study is powered at 40% and the replication study is powered at 80%.
-These power specifications are consistent with the range observed in the OSF data.
-For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the t-test or McNemar's test reject the null hypothesis in favor of the alternative.
-The results of these simulations are displayed in Table 1.
-As expected, the probability of rejecting a true null hypothesis ranges from 0.8 to 1.0 for the two tests.
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|l | c | c | c |}
-\hline\hline
-& \textbf{Power = 40/60} & \textbf{Power = 60/80} & \textbf{Power = 40/80} \\
-\hline
-$\theta =$ \textbf{0.2} & & & \\
-\hspace{3mm} McNemar's & 0.808 (0.01) & 0.873 (0.01) & 0.999 (0.0003) \\
-\hspace{3mm} T test & 0.849 (0.01) & 0.825 (0.01) & 0.999 (0.0002)\\
-\hline
-$\theta =$ \textbf{0.5} & & & \\
-\hspace{3mm} McNemar's & 0.808 (0.01) & 0.874 (0.01) & 0.999 (0.0002)  \\
-\hspace{3mm} T test & 0.851 (0.01) & 0.826 (0.01) & 0.999 (0.0002) \\
-\hline
-$\theta =$ \textbf{0.8} & & & \\
-\hspace{3mm} McNemar's & 0.804 (0.01) & 0.874 (0.01) & 0.999 (0.0001) \\
-\hspace{3mm} T test & 0.851 (0.01) & 0.824 (0.01) & 1.000 (0.0000)\\
-\hline\hline
-\end{tabular}
-\caption{Probability of rejecting a true null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in  parentheses and rounded to the first significant digit.}
-\end{table}
-
-In data simulations to test our analysis of power of the test, we fix the power of the original and replication studies at either 40%, 60%, or 80%.
-For each of these power specifications, we then consider the case where (1) the original study has a small effect size and the replication study has a medium effect size, (2) the original study has a medium effect size and the replication study has a large effect size, and (3) the original study has a small effect size and the replication study has a large effect size.
-For each of these data specifications, we run 10,000 simulations and calculate the proportion in which the \textit{t} test or McNemar's test reject the null hypothesis in favor of the alternative.
-The results of these simulations are displayed in Table 2.
-As expected, the probability of rejecting a false null hypothesis is approximately 0.05 for both the \textit{t} test and McNemar's test.
-That is, the Type II error rate for both tests is 95%.
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|l | c | c | c |}
-\hline\hline
-& $\theta =$ \textbf{0.2/0.5} & $\theta=$ \textbf{0.5/0.8} & $\theta=$ \textbf{0.2/0.8} \\
-\hline
-\textbf{40\%\ Power} & & & \\
-\hspace{3mm} McNemar's & 0.049 (0.007) & 0.050 (0.007) & 0.049 (0.007) \\
-\hspace{3mm} T test & 0.050 (0.007) & 0.051 (0.007) & 0.050 (0.007)\\
-\hline
-\textbf{60\%\ Power} & & & \\
-\hspace{3mm} McNemar's & 0.050 (0.007) & 0.050 (0.007) & 0.049 (0.007) \\
-\hspace{3mm} T test & 0.050 (0.006) & 0.049(0.008) & 0.049 (0.006)\\
-\hline
-\textbf{80\%\ Power} & & & \\
-\hspace{3mm} McNemar's & 0.051 (0.008) & 0.051 (0.007) & 0.051 (0.008) \\
-\hspace{3mm} T test &  0.048 (0.008) & 0.048 (0.007) & 0.048 (0.007)\\
-\hline\hline
-\end{tabular}
-\caption{Probability of rejecting a false null hypothesis for the \textit{t} test and McNemar's test. Standard errors are reported in paraentheses and rounded to the first significant digit.}
-\end{table}
-
->>>>>>> 796555b595c5e5de3516f9efbf8d6204c04cdeaa

### 2018-06-26 Fishers_method/Fishers_method_short.Rmd
added asymptotic results to appendix B

@@
 ##Appendix B: Fisher's method asymptotic results
+
+Under the Neyman and Pearson hypothesis testing framework, the significance level follows a uniform distribution on $\lbrack0,1\rbrack$ when the null hypothesis holds, however, the exact distribution under the alternative hypothesis is unknown.
+As a consequence of this, we rely on long known results from asymptotic theory.
+Lambert and Hall have shown that, given the test statistic is asymptotically normal, the one sided P-value follows a lognormal distribution with mean $-nc(\theta)$ and variance $n\tau^2(\theta)$ (1982).
+The parameter $c(\theta)$ is defined as half the Bahadur slope, given by $-\frac{1}{n}lim_{n\to\infty}logP_n=c(\theta)$, and is the exponential rate at which the significance level converges to zero under the alternative hypothesis.
+In addition, observe that the variance of the standardized P-value is $\frac {\tau^2(\theta)}{n}$.
+We can approximate the two sided P-value by doubling the one sided one, which implies that $P_n\sim AlogN\lbrack-\frac{1}{2}nc(\theta), \frac {1}{4}n\tau^2(\theta)\rbrack$, and thus $logP_n\sim AN\lbrack-\frac{1}{2} nc(\theta), \frac{1}{4}n\tau^2(\theta)\rbrack$.
+Multiplying $logP_n$ by $-2$, we obtain $-2logP_n\sim AN\lbrack nc(\theta), n\tau^2(\theta)\rbrack$.
+Note that $-2logP_{n_j}$, $j=1,...,k$, are asymptotically independent random variables following identical distributions, therefore under $H_a$ $$X^2_n=-2\sum_{j=1}^{k} logP_{n_j} \sim AN(\sum_{j=1}^{k}n_j c_j(\theta), \sum_{j=1}^{k}n_j \tau^2_j(\theta))$$.
+
+Examining the behavior of $X^2_n$ using simulated data, we find that the normal approximation is valid if the within study sample size $n_j$ is at least 1500, 250, and 100 for $\theta$ equal to 0.2, 0.5, and 0.8, respectively.
+Note that $\theta$ in this case can be interpreted as Cohen's d because $\sigma$ is assumed to be 1.
+When the within study sample sizes are smaller than indicated previously, the distribution of the test statistic is skewed to the right, and therefore the normal approximation is not valid.
+
+```{r,echo=FALSE, message=FALSE, fig.show='hold', fig.align='center'}
+
+p1<-c()
+dist_of_X2<-function(delta,n){
+  for (j in 1:100000){ 
+    p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n)))/sqrt(2/n)))
+  }
+  X2<--2*log(p1)
+}
+
+library(latex2exp)
+par(mfrow=c(3,2))
+
+hist(dist_of_X2(0.2,700), main=TeX('n=700, $\\theta$=0.2'), xlab=expression('X'^2))
+hist(dist_of_X2(0.2,1500), main=TeX('n=1500, $\\theta$=0.2'), xlab=expression('X'^2))
+
+hist(dist_of_X2(0.5,100), main=TeX('n=100, $\\theta$=0.5'),  xlab=expression('X'^2))
+hist(dist_of_X2(0.5,250), main=TeX('n=250, $\\theta$=0.5'),  xlab=expression('X'^2))
+
+hist(dist_of_X2(0.8,50), main=TeX('n=50, $\\theta$=0.8'),  xlab=expression('X'^2))
+hist(dist_of_X2(0.8,100), main=TeX('n=100, $\\theta$=0.8'),  xlab=expression('X'^2))
+```
+
+Now consider the two sample shift problem as an example.
+Let $Y_{11},Y_{21},...,Y_{n1}$ denote a sample of i.i.d.
+observations from a normal distribution with mean $\mu$ and standard deviation 1.
+Let $Y_{12},Y_{22},...,Y_{n2}$ denote a second sample, independent of the first, with i.i.d.
+observations from a normal $(\mu+\theta, 1)$.
+For simplicity, suppose that $n_1$ and $n_2$ are equal.
+We test $H_0:\theta=0$ versus $H_a:\theta\neq0$ by using $\frac{\sqrt{n}(\overline{Y_2}-\overline{Y_1})}{\sqrt{2}}$.
+According to Lambert and Hall, $c(\theta)=\frac{1}{2}\lambda\overline{\lambda}\theta^2$ and $\tau^2(\theta)=\lambda\overline{\lambda}\theta^2$, where $\lambda$ denotes the fractional sample size, and $\overline{\lambda}=1-\lambda$.
+Since we are assuming $n_1$ is equal to $n_2$, both $\lambda$ and $\overline{\lambda}$ are $\frac{1}{2}$.
+Then, if all k studies are testing the same hypothesis, it follows that $X^2_n$ is asymptotically normal with grand mean $\sum_{j=1}^{k}\frac{1}{2}n_j \lambda_j \overline{\lambda_j}\theta_j^2$ and grand variance $\sum_{j=1}^{k}n_j \lambda_j \overline{\lambda_j}\theta_j^2$.
+Consequently, the asymptotic power is:
+
+$$\begin{aligned}
+Power&=Pr(reject \ H_0 \ | \ H_a \ is \ true)\\
+&=1-Pr(fail \ to \ reject \ H_0 \ | \ H_a \ is \ true)\\
+&=1-Pr(|X^2_n| \ < \ \chi^2_{2k} \ | \ H_a \ is \ true)\\
+&=1-\Phi\Big(\frac{\chi^2_{2k}-\sum_{j=1}^{k} \frac{1}{2} n_j\lambda_j \overline{\lambda}_j\theta^2_j}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}\Big)+\Phi\Big(\frac{-\chi^2_{2k}-\sum_{j=1}^{k} \frac{1}{2} n_j\lambda_j \overline{\lambda}_j\theta^2_j}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}\Big)\\
+&=1-\Phi\Big(\frac{\chi^2_{2k}}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}-\frac{1}{2}\Big)+\Phi\Big(\frac{-\chi^2_{2k}}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}-\frac{1}{2}\Big)\\
+&=1
+\end{aligned}.$$
+
+The following plots of the power function reveal that asymptotic power approaches 1 when the total within study sample size is greater than 600, 100, and 50 for a fixed $\theta$ of 0.2, 0.5, and 0.8, respectively.
+
+```{r prt 1, echo=FALSE, message=FALSE, fig.width=3, fig.height=5,fig.show='hold',fig.align='center'}
+power<-function(n,delta){
+  1-pnorm(155.4047, mean=64*(1/2)*n*(1/4)*(delta^2), sqrt(64*n*(1/4)*(delta^2)))+pnorm(-155.4047, mean=64*(1/2)*n*(1/4)*(delta^2), sqrt(64*n*(1/4)*(delta^2)))
+}
+require(latex2exp)
+par(mfrow=c(3,1))
+plot(power(1:700,0.2), type='l', xlab='n', ylab='Power', main=TeX('$\\theta$=0.2'))
+plot(power(1:700,0.5), type='l', xlab='n', ylab='Power', main=TeX('$\\theta$=0.5'))
+plot(power(1:700,0.8), type='l', xlab='n', ylab='Power', main=TeX('$\\theta$=0.8'))
+```

### 2018-06-28 significance/significance.Rmd

@@
 
-Programs of research that examine the replicability of research findings have become more common in the social sciences. 
-On its face, assessing replication seems trivial: just repeat an experiment and check that you get the same results.
-However analyses of these programs reveal that we might mean several different things when we refer to "the same results."
-Indeed, the Open Science Collaboration (OSC) (2015) argued, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.
+<!-- Programs of research that examine the replicability of research findings have become more common in the social sciences.  -->
+<!-- On its face, assessing replication seems trivial: just repeat an experiment and check that you get the same results. -->
+<!-- However analyses of these programs reveal that we might mean several different things when we refer to "the same results." -->
+<!-- Indeed, the Open Science Collaboration (RPP) (2015) argued, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways. -->
 
-Despite immediate criticism of the analyses (see Gilbert, Pettigrew, and King, 2015), the core set of OSC analyses has become something of a standard in replication research.
-Camerer, et al. (2016) conducted an almost identical research design with the same analytic methods on findings in experimental economics. 
-Schweinsberg, et al. (2016), while conducting several---i.e., more than one---replicates of the same findings, aggregated replicate results so that they could use the same analyses as the OSC.
-Thus, the idea that replication is impossible to assess has persisted across multiple high-profile replication research efforts.
+<!-- Despite immediate criticism of the analyses (see Gilbert, Pettigrew, and King, 2015), the core set of RPP analyses has become something of a standard in replication research. -->
+<!-- Camerer, et al. (2016) conducted an almost identical research design with the same analytic methods on findings in experimental economics.  -->
+<!-- Schweinsberg, et al.
+(2016), while conducting several---i.e., more than one---replicates of the same findings, aggregated replicate results so that they could use the same analyses as the RPP.
+-->
+<!-- Thus, the idea that replication is impossible to assess has persisted across multiple high-profile replication research efforts. -->
 
-Given their prevalence in the replication literature, the analyses conducted by the OSC have a few important aspects worth exploring.
-The first involves the definitions of replication that they imply. 
-For instance, comparing $p$-values versus effect estimates may suggest very different ways we may think about studies' "results" and whether they are "the same."
-The second are the properties of the methods used to assess definitions of replication.
-For most replication studies, authors seek to make determinations about whether a finding was successfully replicated.
-Doing so amid uncertainty---such as in effect estimates---runs into similar issues of null hypothesis tests. 
-Thus, unambiguous interpretation of the results of replication research, can depend on the properties of the methods authors use to make these determinations.
+<!-- Given their prevalence in the replication literature, the analyses conducted by the RPP have a few important aspects worth exploring. -->
+<!-- The first involves the definitions of replication that they imply.  -->
+<!-- For instance, comparing $p$-values versus effect estimates may suggest very different ways we may think about studies' "results" and whether they are "the same." -->
+<!-- The second are the properties of the methods used to assess definitions of replication. -->
+<!-- For most replication studies, authors seek to make determinations about whether a finding was successfully replicated. -->
+<!-- Doing so amid uncertainty---such as in effect estimates---runs into similar issues of null hypothesis tests.  -->
+<!-- Thus, unambiguous interpretation of the results of replication research, can depend on the properties of the methods authors use to make these determinations. -->
 
-In this paper, we explore methods for assessing replication as proposed by the OSC.
-We argue that definitions of replication ought to involve effect parameters, rather than $p$-values. 
-Using this logic, we then frame the set of analyses common in replication research as null hypothesis tests about replication. 
-The properties of these tests are computed under some simplifying assumptions, and shown that there are conditions where they may be suboptimal.
+<!-- In this paper, we explore methods for assessing replication as proposed by the RPP. -->
+<!-- We argue that definitions of replication ought to involve effect parameters, rather than $p$-values.  -->
+<!-- Using this logic, we then frame the set of analyses common in replication research as null hypothesis tests about replication.  -->
+<!-- The properties of these tests are computed under some simplifying assumptions, and shown that there are conditions where they may be suboptimal. -->
 
@@
 In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
-For programs such as that of the OSC, this involves $\theta_1$ for the original study and $\theta_2$ for the replicate.
+For programs such as that of the Replication Project: Psychology (RPP) (Open Science Collaboration, 2015), this involves $\theta_1$ for the original study and $\theta_2$ for the replicate.
 However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
@@
 Thus, it may make sense to define replication in terms of differences in experiments rather than statistical noise.
-Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2018).
+Thus, we offer $\theta_1 = \theta_2$ as a definition of replication worth testing.
+<!-- Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2018). -->
 
-In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication.
-However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
-Thus, a more practical definition might posit that studies replicate if $\theta_1 \approx \theta_2$.
-For further discussion on this distinction see Hedges and Schauer (2018).
-The key distinction here, is that the \textit{definition} of replication involves the parameters $\theta_i$, while \textit{analyses} involve the estimates $T_i$ and their variances $v_i$
+<!-- In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication. -->
+<!-- However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results. -->
+<!-- Thus, a more practical definition might posit that studies replicate if $\theta_1 \approx \theta_2$. -->
+<!-- For further discussion on this distinction see Hedges and Schauer (2018). -->
+<!-- The key distinction here, is that the \textit{definition} of replication involves the parameters $\theta_i$, while \textit{analyses} involve the estimates $T_i$ and their variances $v_i$ -->
 
-For this paper, we interpret the OSC analyses in the tradition of statistical hypothesis testing.
-Each analysis can be seen of a test of the null hypothesis $H_0: \theta_1 = \theta_2$, the metric used can be seen as a test statistic, and determinations that a finding failed to replicate correspond to a rejection of this null hypothesis.
-Note that this is not the only structure we might infer here, and is not even the only way to operationalize these methods as hypothesis tests.
-However, it is in the tradition of statistical hypothesis tests about heterogeneity across effects.
+<!-- For this paper, we interpret the RPP analyses in the tradition of statistical hypothesis testing. -->
+<!-- Each analysis can be seen of a test of the null hypothesis $H_0: \theta_1 = \theta_2$, the metric used can be seen as a test statistic, and determinations that a finding failed to replicate correspond to a rejection of this null hypothesis.
+-->
+<!-- Note that this is not the only structure we might infer here, and is not even the only way to operationalize these methods as hypothesis tests. -->
+<!-- However, it is in the tradition of statistical hypothesis tests about heterogeneity across effects. -->
 
@@
 
-One metric used by the OSC to determine if studies successfully replicated to to check that both the original and replicate study corresponded in sign and statistical significance.
-That is, findings were deemed to have replicated if both studies found significant effects in the same direction---i.e., both positive---or if both studies found null effects.
-This type of metric would seem to answer an important question about replication, particularly of published research which more often involves statistically significant findings.
-Indeed, Fisher (1935) noted that "we may say that a phenomenon is experimentally demonstrable when we know how to conduct an experiment which will rarely fail to give us statistically significant results."
+One metric used by the RPP to determine if studies successfully replicated was if the original and replicate study corresponded in sign and statistical significance.
+Findings were deemed to have replicated if both studies found significant effects in the same direction---i.e., both positive---or if both studies found null effects.
+Heuristically, this would seem logical, and indeed, Fisher (1935) noted that "we may say that a phenomenon is experimentally demonstrable when we know how to conduct an experiment which will rarely fail to give us statistically significant results."
 
@@
 The null hypothesis would be rejected if $T_1$ is significant and $T_2$ is either not significant or significant and has an opposite direction (or visa versa). 
-Thus, the type I error rate is given by
+If both $T_1$ and $T_2$ are subject to two-tailed $\alpha = 0.05$ level tests, the type I error rate is given by
 <!-- \begin{align} -->
@@
 
-In Figure \ref{fig:sig:falsepos}, we see that the false positive rates are nearly all above the nominal 5\%. 
+In Figure \ref{fig:sig:falsepos}, we see that the error rate is largest when the power of both studies is moderate. 
+If both studies have high power, then it becomes more likely that both estimates will be significant; and if both studies have low power, then it becomes more likely that both estimates will be nonsignificant.
+However, the false positive rates in Figure \ref{fig:sig:falsepos} are nearly all above the nominal 5\%. 
 This is not necessarily an artifact of the scale of the graph. 
@@
 Thus, unless study 1 has at least 95\% power, determining if studies replicated according to correspondence of sign and statistical significance can result in type I error rates well above 5\%.
+
+For reference, suppose all 100 of the RPP studies involved studies that replicated exactly, with effects that are nonzero ($\theta_1 = \theta_2 \neq 0$).
+If all studies were powered at 80\%, then we would expect this procedure to falsely determine that 41 findings failed to replicate. 
+If the power of the studies was lower, say 60%, then we would expect this procedure to falsely conclude that 48 of findings failed to replicate.
 
@@
 
-Conversely, if we assume $\theta_1 \neq \theta_2$, then we would want the test to reject the null hypothesis frequently---that is, we would want a low type II error rate.
+Conversely, if we assume $\theta_1 \neq \theta_2$, then we would want the procedure to determine that the findings do not replicate frequently---that is, we would want a low type II error rate.
 To conclude that the studies do replicate would require both findings to be significant and in the same direction (e.g., positive), or to both be nonsignificant.
@@
 Thus, if study 1 has low power, so that $\beta_1$ is large, then this test would be likely to conclude that the studies successfully replicated. 
-For instance, if study 1 has power 80\%, then this test would conclude that the studies do not replicate 21\% of the time. 
+For instance, if study 1 has power 80\%, then this test would conclude that the studies do  replicate 21\% of the time. 
 However, if the power were lower, say 60\%, then the test would make this error 39.5\% of the time.
 
-If both of the results are nonzero but the same direction (e.g., both positive), then the error rate will be large whenever both studies have high (or low) power. 
+If both of the parameters are nonzero but the same direction (e.g., both positive), then the error rate will be large whenever both studies have high (or low) power. 
 For instance, if $\theta_1$ is much larger than $\theta_2$, but both have power 80\%, then this test will conclude that the studies replicate almost 68\% of the time! 

### 2018-06-29 bootsims.R
updated n's to that they would have exact power

@@ -142,9 +142,9 @@ case2tab<-function(N, theta, powern1, powern2, powern3, rowNames, colNames){
 
 N<-1000
 #theta<-c(0.2,0.5,0.8)
-#power40<-c(147,24,10)
-#power60<-c(245,40,16)
-#power80<-c(393,63,25)
+power40<-c(145.6332,23.30131,9.102075)
+power60<-c(244.9453,39.19124,15.309079)
+power80<-c(392.4541,62.79265,24.528380)
 row<-c("Theta 0.2", "Theta 0.5", "Theta 0.8")
 col<-c("Power 40/60", "Power 40/80", "Power 60/80")
 c2table<-case2tab(N,theta, power80,power60, power40, row, col)

### 2018-06-29 McNemar's Test/draft_mcnemars_ttest.Rmd
edits based off critics for t-test.

@@
 Unlike McNemar's test, the t-test assesses replication by directly comparing original and replication \textit{p}-values within a study pair.
-The null hypothesis for the t-test states that the original and replication studies have the same \textit{p}-value (i.e.
-$H_0: p_{1j} = p_{2j}$).
+The null hypothesis for the t-test states that the original and replication studies' mean difference is zero (i.e.
+$H_0: \mu_{p_{1j}- p_{2j}}=0$).
 The test statistic is given by $t = \frac{\bar{X}_{diff}}{SD_{diff}/\sqrt{k}}$ where $\bar{X}_{diff}$ is the mean of all differences between original and replication \textit{p}-values (i.e.
@@
 The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function..
-Assuming that the variance is known for such a study and is $\frac{4}{n}$ where $n$ is the sample size then, the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$.
+Assuming that the variance is known for such a study which is a function of the sample size $n_i$, then the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n_i$.
 Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as
@@
 $$
-\frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
+\mu_{p_{1j}- p_{2j}} = \frac{1}{k}\sum\limits_{j=1}^k \Big [ \Phi^{-1} \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \Big) - \Phi^{-1} \Big( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \Big) \Big]
 $$

### 2018-06-29 McNemar's Test/draft_mcnemars_ttest.Rmd
deleted extra period.

@@
 Consider a single study (either original or replication), with the null hypothesis $H_0:\theta_i = 0$.
-The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function..
+The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function.
 Assuming that the variance is known for such a study which is a function of the sample size $n_i$, then the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n_i$.

### 2018-07-12 McNemar's Test/draft_mcnemars_ttest.Rmd
Sentence change about variance of 4/n

@@
 The finding from this study is significant if $\frac{|T_{i}|}{\sqrt{v_i}} \geq 1.96$ which is equivalent to $p_i \leq 0.05$ since $p_i = \Phi^{-1} \Big ( \frac{|T_i|}{\sqrt{v_i}} \Big)$ where $\Phi^{-1}$ is the standard normal quantile function.
-Assuming that the variance is known for such a study which is a function of the sample size $n_i$, then the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n_i$.
+Assuming that for each group $\sigma_i^2=1$, then since they are coming from paired data of equal size the variance for each $v_i$ is $\frac{4\sigma^2}{n}$ thus making $v_i = \frac{4}{n}$.
+Then the power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n_i$.
 Extending this to the test statistic for the t-test, we note that the mean of the paired difference of \textit{p}-values can be written as

### 2018-07-12 McNemar's Test/draft_mcnemars_ttest.Rmd
Made some edits

@@
 $$
-H_0: \sum\limits_{j=1}^k P \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 |i \Big) P(i) = \sum\limits_{j=1}^k P \Big ( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96 |i \Big) P(i)
+H_0: \sum\limits_{j=1}^k P \Big ( \frac{|T_{1j}|}{\sqrt{v_{1j}}} \geq 1.96 |j \Big) P(j) = \sum\limits_{j=1}^k P \Big ( \frac{|T_{2j}|}{\sqrt{v_{2j}}} \geq 1.96 |j \Big) P(j)
 $$
@@
 $$
-This implies that the McNemar's test statistic $X^2$  $\rightarrow$ $\sum\limits_{j=1}^k \mathbf{1} \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \rbrace$ as the replication sample sizes grow large.
+This implies that the McNemar's test statistic $X^2 = \frac{(n_{10} - n_{01})^2}{n_{10} + n_{01}}$  $\rightarrow$ $\sum\limits_{j=1}^k \mathbf{1} \lbrace \frac{|T_{1j}|}{\sqrt{v_{1j}}} < 1.96 \rbrace$ as the replication sample sizes grow large.
 Let's assume that all $k$ original studies have 80% power to detect their true effect.
@@
 Recall that McNemar's test rejects if $X^2 > \chi^2_{[0.05,1]}$.
-In this instance, as long as $k>2$, McNemar's test will always reject despite the true effect sizes being identical for all $k$ study pairs.
+In this instance, as long as $k \geq 20$, McNemar's test will always reject despite the true effect sizes being identical for all $k$ study pairs.
 
@@
 \newpage
-## Appendix
+
+## Appendix A: Reframing McNemar's Null Hypothesis
+
+Recall that the null hypothesis for McNemar's test states that the marginal probabilities for each outcome are the same.
+That is, $H_0: p_{11} + p_{10} = p_{11} + p_{01}$ where $p_{11}$ is the proportion of $k$ study pairs in which both findings were significant, $p_{10}$ is the proportion of $k$ study pairs in which the original finding was significant and the replication finding was not, and $p_{01}$ is the proportion of $k$ study pairs in which the original finding was not sigificant and the replication finding was.
+Note that this is equivalent to the null hypothesis $p_{1+} = p_{+1}$ where $p_{1+}$ is the proportion of $k$ study pairs in which the original finding was significant and $p_{+1}$ is the proportion of $k$ study pairs in which the replication finding was significant.
+We can reformulate the null hypothesis in terms of the p-values of the original and replication studies as follows.
+
+$$
+\begin{aligned}
+H_0&:  p_{1+} = p_{+1} \\
+H_0&: \frac{1}{k}\sum\limits_{j=1}^k p_{1j} \leq 0.05 = \frac{1}{k}\sum\limits_{j=1}^k p_{2j} \leq 0.05
+
+\end{aligned}
+$$
+
+## Appendix B: Agreement between McNemar's test and the \textit{t} test
 

### 2018-07-19 Fishers_method/Fishers_method_short.Rmd
added Rrita's name :)
changed subscripts

@@
 title: "Fisher’s Method"
-author: "Katie Fitzgerald"
+author: "Katie Fitzgerald and Rrita Zejnullahi"
 date: "06/20/2018"
@@
 The OSC applied Fisher's method to this set of non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
-Therefore, their hypothesis for Fisher's test could be formalized as follows, where $\theta_i$ is the treatment effect for study $i$:
+Therefore, their hypothesis for Fisher's test could be formalized as follows, where $\theta_j$ is the treatment effect for study $j$:
 \begin{center}
@@
 
-$H_a$: at least one $\theta_i\neq 0,\quad i=1,...,64$
+$H_a$: at least one $\theta_j\neq 0,\quad j=1,...,64$
 \end{center}
 
-Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results, and thus $p_i \geq 0.05$ for all $i$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{i=1}^{k} ln(p_i^*)=-2\sum_{i=1}^{k} ln(\frac{p_i-0.05}{0.95}),$$ where under $H_0$, the $p_i^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies (in this case $k=64$).
+Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results, and thus $p_j \geq 0.05$ for all $j$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{j=1}^{k} ln(p_j^*)=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95}),$$ where under $H_0$, the $p_j^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies (in this case $k=64$).
 Low p-values give a larger test statistic, leading to a rejection of $H_0$.
@@
 In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall treatment effect $\Theta=0$, this type of conservative test may be appropriate.
-In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_i$ has no bearing on the 63 other $\theta_i's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
+In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_j$ has no bearing on the 63 other $\theta_j's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
 
@@
 
-For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma_i}$, where $\theta_i$ is the mean difference between the treatment and control groups in study $i$, and $\sigma_i$ is the known and equal variance among the treatment and control populations.
-Assuming equal sample sizes in the treatment and control groups within study $i$ (that is, let $n_i^t=n_i^c=n_i$), $\delta_i$ is estimated by $d_i\sim N(\delta_i,\frac{2}{n_i})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
+For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the known and equal variance among the treatment and control populations.
+Assuming equal sample sizes in the treatment and control groups within study $j$ (that is, let $n_j^t=n_j^c=n_j$), $\delta_j$ is estimated by $d_j\sim N(\delta_j,\frac{2}{n_j})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
 
@@
 
-$H_a$: at least one $\delta_i\neq 0,\quad i=1,...,64,$
+$H_a$: at least one $\delta_j\neq 0,\quad j=1,...,64,$
 \end{center} 
-and the 64 p-values to be summed in Fisher's test statistic can be calculated as $p_i=2(1-\Phi(\frac{|d_i|}{\sqrt{2/n_i}})).$
+and the 64 p-values to be summed in Fisher's test statistic can be calculated as $p_j=2(1-\Phi(\frac{|d_j|}{\sqrt{2/n_j}})).$
 
-We first consider the power of Fisher's method when $n_i=76/2=38$ for all $i$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
+We first consider the power of Fisher's method when $n_j=76/2=38$ for all $j$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
 The results of the power simulations for median sample size are given in Table 1.[^3] Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
@@
 
-Table: Power of Fisher's method given median sample size $(n_i=38)$ for varying $\delta$ and true # of non-null effects
+Table: Power of Fisher's method given median sample size $(n_j=38)$ for varying $\delta$ and true # of non-null effects
 ```{r,echo=FALSE}
@@
 ```
-[^4]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_i$ among the studies for which the power is at most 99.99% to detect the given $\delta_i$.
-The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ were $n_i=745; 159;$ and $100$ respectively.
+[^4]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
+The largest sample sizes used for $\delta_j=0.2; 0.5$; and $0.8$ were $n_j=745; 159;$ and $100$ respectively.
 See Appendix A.3 for further discussion.
@@
 
-Assume each study $i$ is testing the presence of some treatment effect $\theta_i$, where $i=1,...,k$, and $k$ is the number of studies.
-Assume equal sample sizes for the treatment and control groups ($n_{ti}=n_{ci}=n_i$).
-Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($i=1,...,k$, and $j=1,...n_i$), and assume $Y_{ij}^t\sim N(\mu_i +\theta_i, \sigma_i^2)$ and $Y_{ij}^c\sim N(\mu_i, \sigma_i^2)$, where the $\sigma_i^2$'s are known.
+Assume each study $j$ is testing the presence of some treatment effect $\theta_j$, where $j=1,...,k$, and $k$ is the number of studies.
+Assume equal sample sizes for the treatment and control groups ($n_{tj}=n_{cj}=n_j$).
+Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($i=1,...,k$, and $j=1,...n_i$), and assume $Y_{jl}^t\sim N(\mu_j +\theta_j, \sigma_j^2)$ and $Y_{jl}^c\sim N(\mu_j, \sigma_j^2)$, where the $\sigma_j^2$'s are known.
 Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in the OSC subset of 64 replicate studies are of these two types (89%).
 
-Note then, $\theta_i$ is the difference in means between the treatment and control groups, and its estimate $T_i=\overline{Y_{i.}^t}-\overline{Y_{i.}^c}$ has variance $v_i=\frac{\sigma^2}{n_{ci}} +\frac{\sigma^2}{n_{ti}}=\frac{2\sigma^2}{n_i}.$ For simplicity of interpretation but without loss of generality, we work with the standardized scale of Cohen's d, defined as $\delta_i=\frac{\theta_i}{\sigma}$.
-Note then that $\delta_i$ is estimated by $d_i=\frac{T_i}{\sigma}=\frac{T_i}{\sqrt{\frac{v_i}{2/n_i}}}=\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}}.$ Since under $H_0$, $T_i\sim N(\theta_i,v_i) \Rightarrow \frac{T_i}{\sqrt{v_i}} \sim N(\frac{\theta_i}{\sqrt{v_i}},1)$ and $\sqrt{\frac{2}{n_i}}$ is a constant, then we have  $Var(d_i)=Var(\frac{T_i}{\sqrt{v_i}}\sqrt{{\frac{2}{n_i}}})=\frac{2}{n_i}$.
-Therefore, $d_i\sim N(\delta_i,\frac{2}{n_i})$.
+Note then, $\theta_j$ is the difference in means between the treatment and control groups, and its estimate $T_j=\overline{Y_{j.}^t}-\overline{Y_{j.}^c}$ has variance $v_j=\frac{\sigma^2}{n_{cj}} +\frac{\sigma^2}{n_{tj}}=\frac{2\sigma^2}{n_j}.$ For simplicity of interpretation but without loss of generality, we work with the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma}$.
+Note then that $\delta_j$ is estimated by $d_j=\frac{T_j}{\sigma}=\frac{T_j}{\sqrt{\frac{v_j}{2/n_j}}}=\frac{T_j}{\sqrt{v_j}}\sqrt{{\frac{2}{n_j}}}.$ Since under $H_0$, $T_j\sim N(\theta_j,v_j) \Rightarrow \frac{T_j}{\sqrt{v_j}} \sim N(\frac{\theta_j}{\sqrt{v_j}},1)$ and $\sqrt{\frac{2}{n_j}}$ is a constant, then we have  $Var(d_j)=Var(\frac{T_j}{\sqrt{v_j}}\sqrt{{\frac{2}{n_j}}})=\frac{2}{n_j}$.
+Therefore, $d_j\sim N(\delta_j,\frac{2}{n_j})$.
 
@@
 Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
-That is, we draw a random variable $d_i$ from a $N(\delta_i,\frac{2}{n_i})$ distribution, where $\delta_i\neq0$ and compute its p-value.
-We continue drawing $d_i's$ until we obtain $m$ p-values greater than 0.05 (due to the OSC restriction of only considering replicate studies with non-significant results).
-We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{i=0}^{k} ln(\frac{p_i-0.05}{0.95})$.
-We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{l=0}^{N}I_{\{X^2_l>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.
+That is, we draw a random variable $d_j$ from a $N(\delta_j,\frac{2}{n_j})$ distribution, where $\delta_j\neq0$ and compute its p-value.
+We continue drawing $d_j's$ until we obtain $m$ p-values greater than 0.05 (due to the OSC restriction of only considering replicate studies with non-significant results).
+We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95})$.
+We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{q=1}^{N}I_{\{X^2_q>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.
 We let N=100,000.
@@
 As noted in footnote 4 of the text, the largest sample sizes were dropped out of necessity in the "best-case scenario" power simulations presented in Table 2.
-For example, the largest study had $n_i=384351.5$, which is powered at 100% to detect even a small $\delta=0.2$.
+For example, the largest study had $n_j=384351.5$, which is powered at 100% to detect even a small $\delta=0.2$.
 Therefore, it would have been impossible for this study to result in a p-value greater than 0.05 if there was a true non-null effect.
@@
 For each $\delta$, we began the power simulations with the largest $n$ for which the power was at most 0.9999.
-The largest sample sizes used for $\delta_i=0.2; 0.5$; and $0.8$ therefore were $n_i=745; 159;$ and $100$ respectively.
+The largest sample sizes used for $\delta_j=0.2; 0.5$; and $0.8$ therefore were $n_j=745; 159;$ and $100$ respectively.
 Table 4 presents the sample size vectors used for the "best-case scenario" power simulations.

### 2018-07-20 Fishers_method/Fishers_method_short.Rmd
minor edits

@@
 We believe that Fisher's test is not well suited to assess replication.
-In the event that you do reject $H_0$, this result only tells you that at least one study was a false negative, but it does not till you *which* study did in fact have a true effect.
-Even still, it provides no information about the size or direction of that true non-zero effect and whether or not it replicates the original finding.
+In the event that you do reject $H_0$, this result only tells you that at least one study was a false negative, but it does not tell you *which* study did in fact have a true non-zero effect.
+Even still, it provides no information about the size or direction of that non-zero effect and whether or not it replicates the original finding.
 Presumably, a finding of "no false negatives" would be most informative in assessing replication in this scenario, but this can never be validly concluded from Fisher's method since that would require concluding the null hypothesis.
@@
 In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall treatment effect $\Theta=0$, this type of conservative test may be appropriate.
-In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_j$ has no bearing on the 63 other $\theta_j's$, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
+In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_j$ has no bearing on the 63 other $\theta_j$'s, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
 
@@
 
-For simplicity of interpretation but without loss of generality, we will work with the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the known and equal variance among the treatment and control populations.
+For simplicity of interpretation but without loss of generality, we will work with treatment effects on the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the known and equal variance among the treatment and control populations.
 Assuming equal sample sizes in the treatment and control groups within study $j$ (that is, let $n_j^t=n_j^c=n_j$), $\delta_j$ is estimated by $d_j\sim N(\delta_j,\frac{2}{n_j})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
@@
 The results of the power simulations for median sample size are given in Table 1.[^3] Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
-As shown in the last column of Table 1, we find that there needs to be 10 large effects in order to achieve close to the standard 80% power (power=0.7743).
+As shown in the last column of Table 1, we find that there need to be 10 large effects in order to achieve close to the standard 80% power (power=0.7743).
 If there is only one large effect, Fisher's test only has approximately 8% power to reject.
-Even when all 64 studies have small effects, Fisher's method only has 70% power to reject.
+Even when all 64 studies have small effects, Fisher's method has just 70% power to reject.
 
@@
 In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
-That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^4] As shown in the first row of Table 2, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
+That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^4][^5] As shown in the first row of Table 2, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
 When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340).
@@
 ```
-[^4]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
+[^4]:This provides a "best-case scenario" because large $n_j$'s lead to smaller $p_j$'s, which in turn result in a larger test statistic $X^2$ and greater likelihood of rejecting $H_0$ (i.e.
+higher power).
+
+[^5]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
 The largest sample sizes used for $\delta_j=0.2; 0.5$; and $0.8$ were $n_j=745; 159;$ and $100$ respectively.

### 2018-07-24 Fishers_method/power_sims.R
adapt simulation code to be for general k created .RDS files for k=10, n=50 & n=100 sims

@@ -1,9 +1,10 @@
-power_sims<-function(N,M,delta,n){
+power_sims<-function(N,M,delta,n,K){
   #########################################################################################
   # TAKES: N; number of simulations 
   #        M; vector of number of non-null effects
   #        delta; effect size under alternative hypothesis, on scale of cohen's d
   #        n; vector of treatment/control sample size across studies (total sample size/2)
+  #        K; number of studies
   # RETURNS: power of Fisher's test to reject
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
   #########################################################################################
@@ -15,9 +16,9 @@ power_sims<-function(N,M,delta,n){
     
     for (i in 1:N){
       
-      #print(i) # can uncomment to show progress for lengthy simulations
+      print(i) # can uncomment to show progress for lengthy simulations
     
-      p0<-runif(64 - M[k], 0.05, 1) #draws p-values for the true null effects
+      p0<-runif(K - M[k], 0.05, 1) #draws p-values for the true null effects
       
       p1<-c() #create list to store p-values drawn for non-null effects
       
@@ -26,10 +27,10 @@ power_sims<-function(N,M,delta,n){
         p1[j]<-0
         
         while (p1[j] <= 0.05) { #throw away p-values<=0.05
-          p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n[j])))/sqrt(2/n[j])))
+          p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n[j] + delta^2/(4*n[j]))))/sqrt(2/n[j] + delta^2/(4*n[j]))))
         }
         
-        #print(n[j]) # can uncomment to show progress for lengthy simulations 
+        print(n[j]) # can uncomment to show progress for lengthy simulations 
         }
       
       #test statistic for Fisher's method, with transformation for truncating p-values
@@ -37,7 +38,7 @@ power_sims<-function(N,M,delta,n){
     
       }
     
-    Power[k]<-sum(T>155.4047)/N
+    Power[k]<-sum(T > qchisq(0.95, 2*K) )/N
   
[…]

### 2018-07-24 Fishers_method/Fishers_method_short.Rmd
run sims and add tables for k=10, n=25 and n=75

@@
 
-For simplicity of interpretation but without loss of generality, we will work with treatment effects on the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the known and equal variance among the treatment and control populations.
+For simplicity of interpretation but without loss of generality, we will work with treatment effects on the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the equal variance among the treatment and control populations.
 Assuming equal sample sizes in the treatment and control groups within study $j$ (that is, let $n_j^t=n_j^c=n_j$), $\delta_j$ is estimated by $d_j\sim N(\delta_j,\frac{2}{n_j})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses

### 2018-07-24 Fishers_method/power_sims.R

@@ -75,14 +75,34 @@ saveRDS(power_large_n,"power_large_n.RDS")
 saveRDS(large_n_used,"large_n_used.RDS")
 
 ###NEW TABLES###
+
+##k=10##
 M<-c(1,2,3,4,5,10)
+
+#n=25
+n<-rep(25,10)
+power_delta_0.2_k10_n25 <-power_sims(100000,M,0.2,n,10)
+power_delta_0.5_k10_n25<-power_sims(100000,M,0.5,n,10)
+power_delta_0.8_k10_n25<-power_sims(100000,M,0.8,n,10)
+table_k10_n25<-data.frame(M,power_delta_0.2_k10_n25,power_delta_0.5_k10_n25,power_delta_0.8_k10_n25)
+saveRDS(table_k10_n25,"table_k10_n25.RDS")
+#n=50
 n<-rep(50,10)
 power_delta_0.2_k10 <-power_sims(100000,M,0.2,n,10)
 power_delta_0.5_k10<-power_sims(100000,M,0.5,n,10)
 power_delta_0.8_k10<-power_sims(100000,M,0.8,n,10)
 table_k10_n50<-data.frame(M,power_delta_0.2_k10,power_delta_0.5_k10,power_delta_0.8_k10)
 saveRDS(table_k10_n50,"table_k10_n50.RDS")
-
+#n=75
+n<-rep(75,10)
+power_delta_0.2_k10_n75 <-power_sims(100000,M,0.2,n,10)
+power_delta_0.5_k10_n75<-power_sims(100000,M,0.5,n,10)
+power_delta_0.8_k10_n75<-power_sims(100000,M,0.8,n,10)
+table_k10_n75<-data.frame(M,power_delta_0.2_k10_n75,power_delta_0.5_k10_n75,power_delta_0.8_k10_n75)
+saveRDS(table_k10_n75,"table_k10_n75.RDS")
+saveRDS(power_delta_0.2_k10_n75,"power_delta_0.2_k10_n75.RDS")
+saveRDS(power_delta_0.5_k10_n75,"power_delta_0.5_k10_n75.RDS")
+#n=100
 n<-rep(100,10)
 power_d.2_k10_n100 <-power_sims(100000,M,0.2,n,10)
 power_d.5_k10_n100<-power_sims(100000,M,0.5,n,10)

### 2018-07-24 Fishers_method/power_sims.R
simulation results for k=100

@@ -109,3 +109,28 @@ power_d.5_k10_n100<-power_sims(100000,M,0.5,n,10)
 power_d.8_k10_n100<-power_sims(100000,M,0.8,n,10)
 table_k10_n100<-data.frame(M,power_d.2_k10_n100,power_d.5_k10_n100,power_d.8_k10_n100)
 saveRDS(table_k10_n100,"table_k10_n100.RDS")
+
+##k=100##
+M<-c(1,2,3,4,5,10,50,100)
+
+#n=25
+n<-rep(25,100)
+power_delta_0.2_k100_n25 <-power_sims(100000,M,0.2,n,100)
+power_delta_0.5_k100_n25<-power_sims(100000,M,0.5,n,100)
+power_delta_0.8_k100_n25<-power_sims(100000,M,0.8,n,100)
+table_k100_n25<-data.frame(M,power_delta_0.2_k100_n25,power_delta_0.5_k100_n25,power_delta_0.8_k100_n25)
+saveRDS(power_delta_0.2_k100_n25,"power_delta_0.2_k100_n25.RDS")
+saveRDS(power_delta_0.5_k100_n25,"power_delta_0.5_k100_n25.RDS")
+saveRDS(power_delta_0.8_k100_n25,"power_delta_0.8_k100_n25.RDS")
+saveRDS(table_k100_n25,"table_k100_n25.RDS")
+
+#n=50
+n<-rep(50,100)
+power_delta_0.2_k100_n50 <-power_sims(100000,M,0.2,n,100)
+power_delta_0.5_k100_n50<-power_sims(100000,M,0.5,n,100)
+power_delta_0.8_k100_n50<-power_sims(100000,M,0.8,n,100)
+table_k100_n25<-data.frame(M,power_delta_0.2_k100_n25,power_delta_0.5_k100_n25,power_delta_0.8_k100_n25)
+saveRDS(power_delta_0.2_k100_n25,"power_delta_0.2_k100_n25.RDS")
+saveRDS(power_delta_0.5_k100_n25,"power_delta_0.5_k100_n25.RDS")
+saveRDS(power_delta_0.8_k100_n25,"power_delta_0.8_k100_n25.RDS")
+saveRDS(table_k100_n25,"table_k100_n25.RDS")

### 2018-07-25 Fishers_method/power_sims.R
reformatting data for plot, create plot

@@ -127,10 +127,137 @@ saveRDS(table_k100_n25,"table_k100_n25.RDS")
 #n=50
 n<-rep(50,100)
 power_delta_0.2_k100_n50 <-power_sims(100000,M,0.2,n,100)
+saveRDS(power_delta_0.2_k100_n50,"power_delta_0.2_k100_n50.RDS")
 power_delta_0.5_k100_n50<-power_sims(100000,M,0.5,n,100)
+saveRDS(power_delta_0.5_k100_n50,"power_delta_0.5_k100_n50.RDS")
 power_delta_0.8_k100_n50<-power_sims(100000,M,0.8,n,100)
-table_k100_n25<-data.frame(M,power_delta_0.2_k100_n25,power_delta_0.5_k100_n25,power_delta_0.8_k100_n25)
-saveRDS(power_delta_0.2_k100_n25,"power_delta_0.2_k100_n25.RDS")
-saveRDS(power_delta_0.5_k100_n25,"power_delta_0.5_k100_n25.RDS")
-saveRDS(power_delta_0.8_k100_n25,"power_delta_0.8_k100_n25.RDS")
-saveRDS(table_k100_n25,"table_k100_n25.RDS")
+saveRDS(power_delta_0.8_k100_n50,"power_delta_0.8_k100_n50.RDS")
+table_k100_n50<-data.frame(M,power_delta_0.2_k100_n50,power_delta_0.5_k100_n50,power_delta_0.8_k100_n50)
+saveRDS(table_k100_n50,"table_k100_n50.RDS")
+
+#n=75
+n<-rep(75,100)
+power_delta_0.2_k100_n75 <-power_sims(100000,M,0.2,n,100)
+saveRDS(power_delta_0.2_k100_n75,"power_delta_0.2_k100_n75.RDS")
+power_delta_0.5_k100_n75<-power_sims(100000,M,0.5,n,100)
+saveRDS(power_delta_0.5_k100_n75,"power_delta_0.5_k100_n75.RDS")
+power_delta_0.8_k100_n75<-power_sims(100000,M,0.8,n,100)
+saveRDS(power_delta_0.8_k100_n75,"power_delta_0.8_k100_n75.RDS")
+table_k100_n75<-data.frame(M,power_delta_0.2_k100_n75,power_delta_0.5_k100_n75,power_delta_0.8_k100_n75)
[…]

### 2018-07-25 Fishers_method/power_sims.R
change graph colors

@@ -253,6 +253,7 @@ labels <- c('25' = "n=25", '50' = "n=50", '75' = "n=75", '100'= "n=100")
 ggplot(k10, aes(x = M, y = power, color = factor(delta))) +
   theme(panel.grid.minor = element_blank(),
         axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) +
+  scale_color_brewer(palette="Dark2")+
   scale_x_continuous(name = "# of false negatives", limits = c(1,10), breaks = seq(1,10,1)) +
   scale_y_continuous(name = "Power", limits = c(0,1), breaks = seq(0,1,0.1))+
   geom_point(size = 3) +

### 2018-08-20 Fishers_method/Fishers_method_short.Rmd
cleaned code for generating data for power tables and creating grid plots updated code for asymptotic plots

@@
   for (j in 1:100000){ 
-    p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n)))/sqrt(2/n)))
+    p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n + delta^2 / (4*n))))/sqrt(2/n + delta^2 / (4*n))))
   }
@@
 
+d1 <- as.data.frame(dist_of_X2(0.2,700))
+colnames(d1)[1]<-"V1"
+ggplot(d1, aes(x = V1)) +
+  geom_density()
+
 library(latex2exp)
@@
 
-hist(dist_of_X2(0.2,700), main=TeX('n=700, $\\theta$=0.2'), xlab=expression('X'^2))
-hist(dist_of_X2(0.2,1500), main=TeX('n=1500, $\\theta$=0.2'), xlab=expression('X'^2))
+hist(dist_of_X2(0.2,700), main=TeX('n=700, $\\theta$=0.2'), xlab=expression(-2*log(p_i)))
+x<-seq(0,60,0.1)
+curve(dnorm(x, mean = 0.5*700*2*1/4*0.2^2, sd = sqrt(2*700*1/4*0.2^2)), add = TRUE)
+hist(dist_of_X2(0.2,1500), main=TeX('n=1500, $\\theta$=0.2'), xlab=expression(-2*log(p_i)))
 
-hist(dist_of_X2(0.5,100), main=TeX('n=100, $\\theta$=0.5'),  xlab=expression('X'^2))
-hist(dist_of_X2(0.5,250), main=TeX('n=250, $\\theta$=0.5'),  xlab=expression('X'^2))
+hist(dist_of_X2(0.5,100), main=TeX('n=100, $\\theta$=0.5'),  xlab=expression(-2*log(p_i)))
+hist(dist_of_X2(0.5,250), main=TeX('n=250, $\\theta$=0.5'),  xlab=expression(-2*log(p_i)))
 
-hist(dist_of_X2(0.8,50), main=TeX('n=50, $\\theta$=0.8'),  xlab=expression('X'^2))
-hist(dist_of_X2(0.8,100), main=TeX('n=100, $\\theta$=0.8'),  xlab=expression('X'^2))
+hist(dist_of_X2(0.8,50), main=TeX('n=50, $\\theta$=0.8'),  xlab=expression(-2*log(p_i)))
+hist(dist_of_X2(0.8,100), main=TeX('n=100, $\\theta$=0.8'),  xlab=expression(-2*log(p_i)))
 ```

### 2018-08-20 Fishers_method/k10 table.R

@@ -1,0 +1,128 @@
+##k=10##
+M<-c(1,2,3,4,5,10)
+
+#n=25
+n<-rep(25,10)
+power_delta_0.2_k10_n25 <-power_sims(100000,M,0.2,n,10)
+power_delta_0.5_k10_n25<-power_sims(100000,M,0.5,n,10)
+power_delta_0.8_k10_n25<-power_sims(100000,M,0.8,n,10)
+table_k10_n25<-data.frame(M,power_delta_0.2_k10_n25,power_delta_0.5_k10_n25,power_delta_0.8_k10_n25)
+saveRDS(table_k10_n25,"table_k10_n25.RDS")
+#n=50
+n<-rep(50,10)
+power_delta_0.2_k10 <-power_sims(100000,M,0.2,n,10)
+power_delta_0.5_k10<-power_sims(100000,M,0.5,n,10)
+power_delta_0.8_k10<-power_sims(100000,M,0.8,n,10)
+table_k10_n50<-data.frame(M,power_delta_0.2_k10,power_delta_0.5_k10,power_delta_0.8_k10)
+saveRDS(table_k10_n50,"table_k10_n50.RDS")
+#n=75
+n<-rep(75,10)
+power_delta_0.2_k10_n75 <-power_sims(100000,M,0.2,n,10)
+power_delta_0.5_k10_n75<-power_sims(100000,M,0.5,n,10)
+power_delta_0.8_k10_n75<-power_sims(100000,M,0.8,n,10)
+table_k10_n75<-data.frame(M,power_delta_0.2_k10_n75,power_delta_0.5_k10_n75,power_delta_0.8_k10_n75)
+saveRDS(table_k10_n75,"table_k10_n75.RDS")
+#n=100
+n<-rep(100,10)
+power_d.2_k10_n100 <-power_sims(100000,M,0.2,n,10)
+power_d.5_k10_n100<-power_sims(100000,M,0.5,n,10)
+power_d.8_k10_n100<-power_sims(100000,M,0.8,n,10)
+table_k10_n100<-data.frame(M,power_d.2_k10_n100,power_d.5_k10_n100,power_d.8_k10_n100)
+saveRDS(table_k10_n100,"table_k10_n100.RDS")
+
+##get data in correct format
[…]

### 2018-08-20 Fishers_method/power_sims.R

@@ -74,213 +74,77 @@ power_large_n<-data.frame(M,delta_0.2_large_n,delta_0.5_large_n,delta_0.8_large_
 saveRDS(power_large_n,"power_large_n.RDS")
 saveRDS(large_n_used,"large_n_used.RDS")
 
-###NEW TABLES###
-
-##k=10##
-M<-c(1,2,3,4,5,10)
-
-#n=25
-n<-rep(25,10)
-power_delta_0.2_k10_n25 <-power_sims(100000,M,0.2,n,10)
-power_delta_0.5_k10_n25<-power_sims(100000,M,0.5,n,10)
-power_delta_0.8_k10_n25<-power_sims(100000,M,0.8,n,10)
-table_k10_n25<-data.frame(M,power_delta_0.2_k10_n25,power_delta_0.5_k10_n25,power_delta_0.8_k10_n25)
-saveRDS(table_k10_n25,"table_k10_n25.RDS")
-#n=50
-n<-rep(50,10)
-power_delta_0.2_k10 <-power_sims(100000,M,0.2,n,10)
-power_delta_0.5_k10<-power_sims(100000,M,0.5,n,10)
-power_delta_0.8_k10<-power_sims(100000,M,0.8,n,10)
-table_k10_n50<-data.frame(M,power_delta_0.2_k10,power_delta_0.5_k10,power_delta_0.8_k10)
-saveRDS(table_k10_n50,"table_k10_n50.RDS")
-#n=75
-n<-rep(75,10)
-power_delta_0.2_k10_n75 <-power_sims(100000,M,0.2,n,10)
-power_delta_0.5_k10_n75<-power_sims(100000,M,0.5,n,10)
-power_delta_0.8_k10_n75<-power_sims(100000,M,0.8,n,10)
-table_k10_n75<-data.frame(M,power_delta_0.2_k10_n75,power_delta_0.5_k10_n75,power_delta_0.8_k10_n75)
-saveRDS(table_k10_n75,"table_k10_n75.RDS")
-saveRDS(power_delta_0.2_k10_n75,"power_delta_0.2_k10_n75.RDS")
-saveRDS(power_delta_0.5_k10_n75,"power_delta_0.5_k10_n75.RDS")
-#n=100
-n<-rep(100,10)
-power_d.2_k10_n100 <-power_sims(100000,M,0.2,n,10)
-power_d.5_k10_n100<-power_sims(100000,M,0.5,n,10)
[…]

### 2018-08-23 Fishers_method/Fishers_method_082118.Rmd
edits to write-up: took out median table and put in new simulation plot and narrative

--- 2018-08-20 Fishers_method/Fishers_method_short.Rmd
@@
 Among the 100 replication studies conducted by the Open Science Collaboration (OSC), 64 found null effects, defined as having a p-value greater than 0.05.
-The OSC applied Fisher's method to this set of non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
-Therefore, their hypothesis for Fisher's test could be formalized as follows, where $\theta_j$ is the treatment effect for study $j$:
+Because only 3 of the original studies found null effects, when using the "sign and significance" metric to assess replication, the OSC took this to indicate that there was widespread failure in replication.
+One possible explanation for this "failure to replicate," however, would be that the replication studies were simply underpowered to detect the true non-zero effects that the original studies found.
+In general, a finding of a null effect indicates that either the true effect is in fact zero or that it is non-zero but the test was underpowered to detect it.
+Hoping to determine which of these two scenarios is true of their replication studies, the OSC applied Fisher's method to the set of 64 non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
+Their hypothesis for Fisher's test could be formalized as follows, where $\theta_j$ is the treatment effect for study $j$:
 \begin{center}
@@
 
-Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results, and thus $p_j \geq 0.05$ for all $j$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{j=1}^{k} ln(p_j^*)=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95}),$$ where under $H_0$, the $p_j^*$'s follow a uniform distribution on [0,1] and thus $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies (in this case $k=64$).
+Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results, and thus $p_j \geq 0.05$ for all $j$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{j=1}^{k} ln(p_j^*)=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95}),$$ where under $H_0$, $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies (in this case $k=64$).
 Low p-values give a larger test statistic, leading to a rejection of $H_0$.
@@
 The true distribution of Fisher's test statistic under the alternative hypothesis is unknown, and therefore the power cannot be calculated exactly.
-The asymptotic distribution of $X^2$ can be shown to be approximately normal, but this approximation is not valid for the sample sizes in the OSC data.[^1] We therefore turn to simulations to investigate the power of Fisher's method in the OSC scenario.
+The asymptotic distribution of $X^2$ can be shown to be approximately normal as within study sample sizes go to infinity, but this approximation is only valid for impractical sample sizes and detectable effect sizes.[^1] We therefore turn to simulations to investigate the power of Fisher's method in the OSC scenario.
 
@@
 
-For simplicity of interpretation but without loss of generality, we will work with treatment effects on the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the equal variance among the treatment and control populations.
-Assuming equal sample sizes in the treatment and control groups within study $j$ (that is, let $n_j^t=n_j^c=n_j$), $\delta_j$ is estimated by $d_j\sim N(\delta_j,\frac{2}{n_j})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
+For simplicity of interpretation but without loss of generality, we will work with treatment effects on the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma_j}$, where $\theta_j$ is the mean difference between the treatment and control groups in study $j$, and $\sigma_j$ is the equal variance among the treatment and control populations in study $j$.
+Assuming equal sample sizes in the treatment and control groups within study $j$ (that is, let $n_j^t=n_j^c=n_j$), $\delta_j$ is estimated by $d_j\sim N(\delta_j,\frac{2}{n_j} + \frac{\delta_j^2}{4n_j})$.[^2] Under this framework, Fisher's method can be represented in Cohen's d as testing the hypotheses
 
-[^2]:See Appendix A.1 for proof and discussion of simplifying assumptions.
+[^2]: Hedges & Olkin (1985).
 
@@
 \end{center} 
-and the 64 p-values to be summed in Fisher's test statistic can be calculated as $p_j=2(1-\Phi(\frac{|d_j|}{\sqrt{2/n_j}})).$
+and the 64 p-values to be summed in Fisher's test statistic can be calculated as $p_j=2(1-\Phi(\frac{|\delta_j|}{\sqrt{\frac{2}{n_j} + \frac{\delta_j^2}{4n_j}}})).$
 
-We first consider the power of Fisher's method when $n_j=76/2=38$ for all $j$, because 76 is the median sample size in the OSC dataset, and we are assuming equal sample sizes in the treatment and control groups.
-The results of the power simulations for median sample size are given in Table 1.[^3] Note that 0.2, 0.5, and 0.8 correspond to small, medium, and large effects sizes on the scale of Cohen's d.
-As shown in the last column of Table 1, we find that there need to be 10 large effects in order to achieve close to the standard 80% power (power=0.7743).
-If there is only one large effect, Fisher's test only has approximately 8% power to reject.
-Even when all 64 studies have small effects, Fisher's method has just 70% power to reject.
-
-[^3]: See Appendix A.2 for code and details on how the power simulations were conducted.
-
-Table: Power of Fisher's method given median sample size $(n_j=38)$ for varying $\delta$ and true # of non-null effects
-```{r,echo=FALSE}
-library(knitr)
-table <- readRDS("./power sims table with changing n.RDS")
-table_median<-table[,1:4]
-#colnames(table_median)[2]<
-kable(table_median,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"), caption="Power to reject H0 with median sample size (n=38)")
-
-```
+[^3]: See Appendix A.1 for code and details on how the power simulations were conducted.
 
 In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
-That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^4][^5] As shown in the first row of Table 2, even when a study with a very large sample size has a large effect, Fisher's method has less than 10% power to detect it.
+That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^3] This will provide the highest possible power of Fisher's method given these data.[^4] As shown in the first row of Table 1, even when the study with the largest sample size has a large effect, Fisher's method has less than 10% power to detect it.
 When the non-null effects come from the studies with the largest sample sizes, about half of the studies need to have $\delta=0.2$ in order to achieve approximately 80% power (power=0.8340).
@@
 ```{r, echo=FALSE}
+library(knitr)
 table <- readRDS("./power_large_n.RDS")
@@
 ```
-[^4]:This provides a "best-case scenario" because large $n_j$'s lead to smaller $p_j$'s, which in turn result in a larger test statistic $X^2$ and greater likelihood of rejecting $H_0$ (i.e.
+[^3]:This provides a "best-case scenario" because large $n_j$'s lead to smaller $p_j$'s, which in turn result in a larger test statistic $X^2$ and greater likelihood of rejecting $H_0$ (i.e.
 higher power).
 
-[^5]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
+[^4]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
 The largest sample sizes used for $\delta_j=0.2; 0.5$; and $0.8$ were $n_j=745; 159;$ and $100$ respectively.
-See Appendix A.3 for further discussion.
+See Appendix A.2 for further discussion.
+
+We now consider the power of Fisher's method under more general circumstances.
+Figure 1 plots the power of Fisher's method against the proportion of studies with false negatives, for varying numbers of total studies ($k$), effect sizes ($\delta$), and within-group within-study sample sizes ($n$).
+
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
+###CREATE PLOT####
+k10 <- readRDS("./k10_data.RDS")
+k50 <- readRDS("./k50_data.RDS")
+k100 <- readRDS("./k100_data.RDS")
+
+library(ggplot2)
+n_labels <- c('25' = "n=25", '50' = "n=50", '75' = "n=75", '100'= "n=100")
+k_labels <- c( '10' = "k=10", '50' = "k=50", '100' = "k=100")
+
+k10$prop <- k10$M/k10$k
+k50$prop <- k50$M/k50$k
+k100$prop <- k100$M/k100$k
+
+all <- rbind(k10,k50,k100)
+plot_all <- ggplot(all, aes(x = prop, y = power, color = factor(delta))) +
+  facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
+  theme(panel.grid.minor = element_blank(),
+        axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) +
+  scale_color_brewer(palette="Dark2")+
+  scale_x_continuous(name = "Proportion of false negatives", limits = c(0,1), breaks = seq(0,1,.20)) +
+  scale_y_continuous(name = "Power", limits = c(0,1), breaks = seq(0,1,0.1))+
+  #geom_point(size = 3) +
+  #facet_grid(.~n, labeller = labeller(n = labels)) +
+  labs(title = "Power of Fisher's method", color = expression(delta))+
+  #geom_smooth(method = "loess", se = FALSE, size = 0.5)
+  geom_line()
+
+plot_all
+```
+
+For example, in the first pane of the grid, the green curve indicates that when there are 10 studies each having a within-group sample sizes of 25, even when all 10 studies have true small effects, the power of Fisher's method is only around 15%.
+Note the general pattern: the power of Fisher's method increases as within-group sample size increases (left to right in each row), as the number of studies increases (top to bottom in each column), and as the magnitude of the true effect size increases (indicated by the color of the curve).
+
+As demonstrated by Figure 1, Fisher's method has a large Type II error rate under many common scenarios in social science research, and we do not recommend it as a valid metric for assessing replication.
+Failing to reject $H_0$ is likely an artifact of the poor properties of Fisher's method and is not valid evidence to conclude there are no false negatives.
+As stated previously, even in the event that $H_0$ is rejected, it tells the researcher nothing about *which* study has a non-zero effect or if the size or direction of that effect is consistent with the original study.
 
@@
 
-###A.1 Framework and assumptions
-
-Assume each study $j$ is testing the presence of some treatment effect $\theta_j$, where $j=1,...,k$, and $k$ is the number of studies.
-Assume equal sample sizes for the treatment and control groups ($n_{tj}=n_{cj}=n_j$).
-Let $Y_{ij}^t$ and $Y_{ij}^c$ be the observations from the treatment and control groups respectively ($i=1,...,k$, and $j=1,...n_i$), and assume $Y_{jl}^t\sim N(\mu_j +\theta_j, \sigma_j^2)$ and $Y_{jl}^c\sim N(\mu_j, \sigma_j^2)$, where the $\sigma_j^2$'s are known.
-Note this is not an unreasonbale assumption since this framework is required for both the t-test and the ANOVA test, and the majority of the tests in the OSC subset of 64 replicate studies are of these two types (89%).
-
-Note then, $\theta_j$ is the difference in means between the treatment and control groups, and its estimate $T_j=\overline{Y_{j.}^t}-\overline{Y_{j.}^c}$ has variance $v_j=\frac{\sigma^2}{n_{cj}} +\frac{\sigma^2}{n_{tj}}=\frac{2\sigma^2}{n_j}.$ For simplicity of interpretation but without loss of generality, we work with the standardized scale of Cohen's d, defined as $\delta_j=\frac{\theta_j}{\sigma}$.
-Note then that $\delta_j$ is estimated by $d_j=\frac{T_j}{\sigma}=\frac{T_j}{\sqrt{\frac{v_j}{2/n_j}}}=\frac{T_j}{\sqrt{v_j}}\sqrt{{\frac{2}{n_j}}}.$ Since under $H_0$, $T_j\sim N(\theta_j,v_j) \Rightarrow \frac{T_j}{\sqrt{v_j}} \sim N(\frac{\theta_j}{\sqrt{v_j}},1)$ and $\sqrt{\frac{2}{n_j}}$ is a constant, then we have  $Var(d_j)=Var(\frac{T_j}{\sqrt{v_j}}\sqrt{{\frac{2}{n_j}}})=\frac{2}{n_j}$.
-Therefore, $d_j\sim N(\delta_j,\frac{2}{n_j})$.
-
-###A.2 Power simulation logic and code
+###A.1 Power simulation logic and code
 Let there be $m$ false negatives (i.e.
@@
 Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
-That is, we draw a random variable $d_j$ from a $N(\delta_j,\frac{2}{n_j})$ distribution, where $\delta_j\neq0$ and compute its p-value.
+That is, we draw a random variable $d_j$ from a $N(\delta_j,\frac{2}{n_j} + \frac{\delta_j^2}{4n_j})$ distribution, where $\delta_j\neq0$ and compute its p-value.
 We continue drawing $d_j's$ until we obtain $m$ p-values greater than 0.05 (due to the OSC restriction of only considering replicate studies with non-significant results).
 We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95})$.
-We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{q=1}^{N}I_{\{X^2_q>155.4047\}}/N,$ where $I$ is the indicator function and $\chi^2_{128}=155.4047$ is the critical value for Fisher's test with $k=64$ studies.
+We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{q=1}^{N}I_{\{X^2_q > \chi^2_{2k,\alpha}\}}/N,$ where $I$ is the indicator function and $\chi^2_{2k,\alpha}$ is the critical value for Fisher's test with $k$ studies and level $\alpha$.
 We let N=100,000.
@@
 ```{r power_sims}
-power_sims<-function(N,M,delta,n){
-  ########################################################################################
+power_sims<-function(N,M,delta,n,k){
+  #########################################################################################
   # TAKES: N; number of simulations 
-  #        M; vector of number of non-null effects (e.g. M=c(1, 2, 3, 4, 5, 32, 64))
+  #        M; vector of number of non-null effects
   #        delta; effect size under alternative hypothesis, on scale of cohen's d
   #        n; vector of treatment/control sample size across studies (total sample size/2)
+  #        k; number of studies
   # RETURNS: power of Fisher's test to reject
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
-  ########################################################################################
+  #########################################################################################
 
@@
 
-  for(k in 1:length(M)){
+  for(l in 1:length(M)){
 
@@
 
-      p0<-runif(64 - M[k], 0.05, 1) #draws p-values for the true null effects
+      p0<-runif(k - M[l], 0.05, 1) #draws p-values for the true null effects
 
@@
 
-      for (j in 1:M[k]){
+      for (j in 1:M[l]){
 
@@
         while (p1[j] <= 0.05) { #throw away p-values<=0.05
-          p1[j]<-2 * (1 - pnorm(abs(rnorm(1, delta, sqrt(2 / n[j]))) / sqrt(2 / n[j])))
+          z[j] <- rnorm(1,delta, sqrt(2/n[j] + delta^2/(4*n[j])))
+          p1[j] <- 2*(1-pnorm(abs(z[j])/sqrt(2/n[j] + delta^2/(4*n[j]))))
         }
@@
       #test statistic for Fisher's method, with transformation for truncating p-values
-      T[i]<--2 * sum(log((p0 - 0.05)/0.95)) - 2 * sum(log((p1 - 0.05) / 0.95))
+      T[i]<--2*sum(log((p0-0.05)/0.95))-2*sum(log((p1-0.05)/0.95))
 
@@
 
-    Power[k]<-sum(T > 155.4047) / N
+    Power[l]<-sum(T > qchisq(0.95, 2*k) )/N
 
@@
 
-##A.3 "Best-case scenario"" power simulations
+##A.2 "Best-case scenario"" power simulations
 
-As noted in footnote 4 of the text, the largest sample sizes were dropped out of necessity in the "best-case scenario" power simulations presented in Table 2.
+As noted in footnote 4 of the text, the largest sample sizes were dropped out of necessity in the "best-case scenario" power simulations presented in Table 1.
 For example, the largest study had $n_j=384351.5$, which is powered at 100% to detect even a small $\delta=0.2$.
-Therefore, it would have been impossible for this study to result in a p-value greater than 0.05 if there was a true non-null effect.
-Table 3 presents the power of the OSC replicate studies to detect a given $\delta$ using the true sample sizes for the 64 studies.
+Therefore, it would have been impossible for this study to result in a p-value greater than 0.05 if there was a true small effect.
+Table 3 presents the power of the 12 largest OSC replicate studies (among the subset of 64) to detect a given $\delta$.
+For example, the twelfth largest OSC study had a within-group sample size of 100 and had 29% power to detect a small effect, 94% power to detect a medium effect, and >99% power to detect a large effect.
 For each $\delta$, we began the power simulations with the largest $n$ for which the power was at most 0.9999.
 The largest sample sizes used for $\delta_j=0.2; 0.5$; and $0.8$ therefore were $n_j=745; 159;$ and $100$ respectively.
-Table 4 presents the sample size vectors used for the "best-case scenario" power simulations.
+Table 3 presents the resulting sample size vectors used for the "best-case scenario" power simulations.
 
-Table: Power of OSC replicate studies to detect $\delta$ given $n$
+Table: Power of 12 largest OSC replicate studies to detect $\delta$ given $n$
 ```{r,echo=FALSE,warning=FALSE,message=FALSE}
@@
 #shorten # of rows to display in table
-replicate_power<-rbind(replicate_power[1:12,],c("...","...","...","..."))
+replicate_power<-replicate_power[1:12,]
+replicate_power$Study <- seq(1,12,1)
+replicate_power <- replicate_power[,c(5,1,2,3,4)]
 library(knitr)
-kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("n","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power of OSC replicate studies to detect $\\delta$ given $\n$")
+kable(replicate_power,format="markdown",digits=4,align=c('r','r','r','r'),col.names=c("Study","n","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power of OSC replicate studies to detect $\\delta$ given $\n$")
 ```

### 2018-08-23 Fishers_method/k100 table.R

@@ -1,0 +1,220 @@
+##k=100##
+M<-c(1,2,3,4,5,10,50,100)
+
+#n=25
+n<-rep(25,100)
+power_delta_0.2_k100_n25 <-power_sims(100000,M,0.2,n,100)
+power_delta_0.5_k100_n25<-power_sims(100000,M,0.5,n,100)
+power_delta_0.8_k100_n25<-power_sims(100000,M,0.8,n,100)
+table_k100_n25<-data.frame(M,power_delta_0.2_k100_n25,power_delta_0.5_k100_n25,power_delta_0.8_k100_n25)
+saveRDS(table_k100_n25,"table_k100_n25.RDS")
+
+#n=50
+n<-rep(50,100)
+power_delta_0.2_k100_n50 <-power_sims(100000,M,0.2,n,100)
+power_delta_0.5_k100_n50<-power_sims(100000,M,0.5,n,100)
+power_delta_0.8_k100_n50<-power_sims(100000,M,0.8,n,100)
+table_k100_n50<-data.frame(M,power_delta_0.2_k100_n50,power_delta_0.5_k100_n50,power_delta_0.8_k100_n50)
+saveRDS(table_k100_n50,"table_k100_n50.RDS")
+
+#n=75
+n<-rep(75,100)
+power_delta_0.2_k100_n75 <-power_sims(100000,M,0.2,n,100)
+power_delta_0.5_k100_n75<-power_sims(100000,M,0.5,n,100)
+power_delta_0.8_k100_n75<-power_sims(100000,M,0.8,n,100)
+table_k100_n75<-data.frame(M,power_delta_0.2_k100_n75,power_delta_0.5_k100_n75,power_delta_0.8_k100_n75)
+saveRDS(table_k100_n75,"table_k100_n75.RDS")
+
+#n=100
+n<-rep(100,100)
+power_delta_0.2_k100_n100 <-power_sims(100000,M,0.2,n,100)
+power_delta_0.5_k100_n100<-power_sims(100000,M,0.5,n,100)
+power_delta_0.8_k100_n100<-power_sims(100000,M,0.8,n,100)
+table_k100_n100<-data.frame(M,power_delta_0.2_k100_n100,power_delta_0.5_k100_n100,power_delta_0.8_k100_n100)
+saveRDS(table_k100_n100,"table_k100_n100.RDS")
+
+
+
+
+
+##get data in correct format
+table_k100_n100 <- readRDS("~table_k100_n100.RDS")
[…]

### 2018-08-23 Fishers_method/k50 table.R

@@ -1,0 +1,110 @@
+##get data in correct format
+table_k50_n100 <- readRDS("~table_k50_n100.RDS")
+table_k50_n25 <- readRDS("~table_k50_n25.RDS")
+table_k50_n50 <- readRDS("~table_k50_n50.RDS")
+table_k50_n75 <- readRDS("~table_k50_n75.RDS")
+
+k50_n25_d.2 <- as.data.frame(table_k50_n25[,2])
+colnames(k50_n25_d.2)[1] <- "power"
+k50_n25_d.2$n <- rep(25,length(k50_n25_d.2))
+k50_n25_d.2$k <- rep(50,length(k50_n25_d.2))
+k50_n25_d.2$M <- c(1,2,3,4,5,10,25,50)
+k50_n25_d.2$delta <- rep(0.2, dim(k50_n25_d.2)[1])
+
+k50_n25_d.5 <- as.data.frame(table_k50_n25[,3])
+colnames(k50_n25_d.5)[1] <- "power"
+k50_n25_d.5$n <- rep(25,length(k50_n25_d.5))
+k50_n25_d.5$k <- rep(50,length(k50_n25_d.5))
+k50_n25_d.5$M <-c(1,2,3,4,5,10,25,50)
+k50_n25_d.5$delta <- rep(0.5, dim(k50_n25_d.5)[1])
+
+k50_n25_d.8 <- as.data.frame(table_k50_n25[,4])
+colnames(k50_n25_d.8)[1] <- "power"
+k50_n25_d.8$n <- rep(25,length(k50_n25_d.8))
+k50_n25_d.8$k <- rep(50,length(k50_n25_d.8))
+k50_n25_d.8$M <- c(1,2,3,4,5,10,25,50)
+k50_n25_d.8$delta <- rep(0.8, dim(k50_n25_d.8)[1])
+
+
+##########
+k50_n50_d.2 <- as.data.frame(table_k50_n50[,2])
+colnames(k50_n50_d.2)[1] <- "power"
+k50_n50_d.2$n <- rep(50,length(k50_n50_d.2))
+k50_n50_d.2$k <- rep(50,length(k50_n50_d.2))
+k50_n50_d.2$M <- c(1,2,3,4,5,10,25,50)
+k50_n50_d.2$delta <- rep(0.2, dim(k50_n50_d.2)[1])
+
+k50_n50_d.5 <- as.data.frame(table_k50_n50[,3])
+colnames(k50_n50_d.5)[1] <- "power"
[…]

### 2018-08-23 Fishers_method/power_sims.R

@@ -110,7 +110,8 @@ diane<-function(delta,n){
     #p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n + delta^2 / (4*n))))/sqrt(2/n + delta^2 / (4*n))))
     #p1[j]<-2*(1 - pnorm(abs(rnorm(1, 1/2*2*n*1/4*delta^2, sqrt(2*n*1/4*delta^2)))/sqrt(2*n*1/4*delta^2)))
     z[j] <- rnorm(1,delta, sqrt(2/n + delta^2/(4*n)))
-    p1[j] <- (1-pnorm(z[j]/sqrt(2/n + delta^2/(4*n))))
+    #p1[j] <- (1-pnorm(z[j]/sqrt(2/n + delta^2/(4*n))))
+    p1[j] <- 2*(1-pnorm(abs(z[j])/sqrt(2/n + delta^2/(4*n)))) #two-sided
   }
   D<--2*log(p1)
 }
@@ -122,8 +123,9 @@ plot <- function(delta, n){
   ggplot(d1, aes(x = V1)) +
   scale_x_continuous(name = "-2log(p_i)", limits = c(-20,80), breaks = seq(-20,100,10)) +
   geom_histogram(aes(y = ..density..)) +
-  stat_function(fun = dnorm, args = list(mean = 2*n*1/4*delta^2, sd = sqrt(2*n*delta^2))) +
-  labs(title = c(n,delta))
+  #stat_function(fun = dnorm, args = list(mean = 2*n*1/4*delta^2, sd = sqrt(2*n*delta^2))) +
+  stat_function(fun = dnorm, args = list(mean = 1/2*2*n*1/4*delta^2, sd = sqrt(2*n*1/4*delta^2))) + #two-sided
+  labs(title = n)
 }
 
 p1 <- plot(0.2,700)

### 2018-08-31 Fishers_method/Fishers_method_082118.Rmd
extra simulations to fill in gaps in Figure 1 updated narrative to incorprate new plots in body and appendices

@@
 author: "Katie Fitzgerald and Rrita Zejnullahi"
-date: "06/20/2018"
+date: "08/31/2018"
 output: pdf_document
@@
 Because only 3 of the original studies found null effects, when using the "sign and significance" metric to assess replication, the OSC took this to indicate that there was widespread failure in replication.
-One possible explanation for this "failure to replicate," however, would be that the replication studies were simply underpowered to detect the true non-zero effects that the original studies found.
 In general, a finding of a null effect indicates that either the true effect is in fact zero or that it is non-zero but the test was underpowered to detect it.
 Hoping to determine which of these two scenarios is true of their replication studies, the OSC applied Fisher's method to the set of 64 non-significant p-values to test the null hypothesis that a true zero effect held for each study and that there were no false negatives among them.
+That is, they wish to rule out the possibility that being under-powered was the only reason for failure to find significant results in the replication studies.
 Their hypothesis for Fisher's test could be formalized as follows, where $\theta_j$ is the treatment effect for study $j$:
@@
 
-Because the OSC conducted Fisher's test only among the 64 replication studies that did not have significant results, and thus $p_j \geq 0.05$ for all $j$, they used the following transformation of Fisher's test statistic  $$X^2=-2\sum_{j=1}^{k} ln(p_j^*)=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95}),$$ where under $H_0$, $X^2 \sim \chi^2_{2k}$, where $k$ is the number of studies (in this case $k=64$).
+Note that since $p_j \geq 0.05$ for all $j$ in this scenario, the following transformation of Fisher's test statistic was used $$X^2=-2\sum_{j=1}^{k} ln(p_j^*)=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95}),$$ where $X^2 \sim \chi^2_{2k}$ under $H_0$ and $k$ is the number of studies (in this case $k=64$).
 Low p-values give a larger test statistic, leading to a rejection of $H_0$.
 
-We believe that Fisher's test is not well suited to assess replication.
-In the event that you do reject $H_0$, this result only tells you that at least one study was a false negative, but it does not tell you *which* study did in fact have a true non-zero effect.
-Even still, it provides no information about the size or direction of that non-zero effect and whether or not it replicates the original finding.
+We believe that Fisher's test is not well suited to distinguish between low power and the existence of false negatives.
 Presumably, a finding of "no false negatives" would be most informative in assessing replication in this scenario, but this can never be validly concluded from Fisher's method since that would require concluding the null hypothesis.
 While it is never advised to conduct a test in order to conclude the null hypothesis, this switched framework is especially problematic when the test is underpowered to reject $H_0$ because the Type II error rate will be large.
-
 Even though the OSC was able to reject their null hypothesis ($X^2=155.83, p=0.048$), we think that in general Fisher's method is underpowered to answer the question at hand.
 We hypothesize that it requires many and possibly large non-null effects in order to skew the distribution of p-values enough to reject Fisher's null hypothesis.
-In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall treatment effect $\Theta=0$, this type of conservative test may be appropriate.
-In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_j$ has no bearing on the 63 other $\theta_j$'s, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
 
 The true distribution of Fisher's test statistic under the alternative hypothesis is unknown, and therefore the power cannot be calculated exactly.
-The asymptotic distribution of $X^2$ can be shown to be approximately normal as within study sample sizes go to infinity, but this approximation is only valid for impractical sample sizes and detectable effect sizes.[^1] We therefore turn to simulations to investigate the power of Fisher's method in the OSC scenario.
+The asymptotic distribution of $X^2$ can be shown to be approximately normal as within study sample sizes go to infinity, but this approximation is only valid for impractical sample sizes and detectable effect sizes.[^1] We therefore turn to simulations to investigate the power of Fisher's method.
 
@@
 
-In order to consider the power of Fisher's test under a "best-case scenario" in this dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
+In order to consider the power of Fisher's test under a "best-case scenario" in the OSC dataset, we sort the 64 sample sizes and let the non-null effects be from the studies with the largest sample sizes first.
 That is, if there is just one non-null effect we let it be from the largest study; if there are two non-null effects we let them be from the two largest studies, etc.[^3] This will provide the highest possible power of Fisher's method given these data.[^4] As shown in the first row of Table 1, even when the study with the largest sample size has a large effect, Fisher's method has less than 10% power to detect it.
@@
 ```
-[^3]:This provides a "best-case scenario" because large $n_j$'s lead to smaller $p_j$'s, which in turn result in a larger test statistic $X^2$ and greater likelihood of rejecting $H_0$ (i.e.
-higher power).
-
-[^4]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
+[^3]:Note that because we are working with a set of replicate studies which found a p-value greater than 0.05, pairing large effects with very large sample sizes is not realistic, and therefore we begin the simulations with the largest $n_j$ among the studies for which the power is at most 99.99% to detect the given $\delta_j$.
 The largest sample sizes used for $\delta_j=0.2; 0.5$; and $0.8$ were $n_j=745; 159;$ and $100$ respectively.
 See Appendix A.2 for further discussion.
+
+[^4]:This provides a "best-case scenario" because large $n_j$'s lead to smaller $p_j$'s, which in turn result in a larger test statistic $X^2$ and greater likelihood of rejecting $H_0$ (i.e.
+higher power).
 
@@
   facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
+  #facet_grid(~n, labeller = labeller(n = n_labels)) +
   theme(panel.grid.minor = element_blank(),
@@
 
-For example, in the first pane of the grid, the green curve indicates that when there are 10 studies each having a within-group sample sizes of 25, even when all 10 studies have true small effects, the power of Fisher's method is only around 15%.
+For example, in the first pane of the grid, the green curve indicates that when there are 10 studies each having within-group sample sizes of 25, even when all 10 studies have true small effects, the power of Fisher's method is only around 15%.
 Note the general pattern: the power of Fisher's method increases as within-group sample size increases (left to right in each row), as the number of studies increases (top to bottom in each column), and as the magnitude of the true effect size increases (indicated by the color of the curve).
 
-As demonstrated by Figure 1, Fisher's method has a large Type II error rate under many common scenarios in social science research, and we do not recommend it as a valid metric for assessing replication.
-Failing to reject $H_0$ is likely an artifact of the poor properties of Fisher's method and is not valid evidence to conclude there are no false negatives.
-As stated previously, even in the event that $H_0$ is rejected, it tells the researcher nothing about *which* study has a non-zero effect or if the size or direction of that effect is consistent with the original study.
+As demonstrated by Figure 1, Fisher's method has low power and thus a large Type II error rate under many common scenarios in social science research, and we do not recommend it as a valid metric for assessing replication.
+Failing to reject $H_0$ is often likely an artifact of the poor properties of Fisher's method and is not valid evidence to conclude there are no false negatives.
+The test is only adequately powered when there are many effects, potentially large in magnitude and coming from large studies.
+In the scenario where a researcher is combining the results of $k$ studies of the same treatment to test if an overall treatment effect $\Theta=0$, this type of conservative test may be appropriate.
+In the OSC scenario, however, since the 64 studies are not testing the same treatment effect, and one $\theta_j$ has no bearing on the 63 other $\theta_j$'s, the presumed goal would be to detect if there are *any* false negatives among the replicate studies.
+Furthermore, in the event that $H_0$ is rejected, it tells the researcher nothing about *which* study has a non-zero effect or if the size or direction of that effect is consistent with the original study.
 
-#Appendices
+****
+#Appendix A: Fisher's method power simulations
 
-##Appendix A: Fisher's method power simulations
-
-###A.1 Power simulation logic and code
+##A.1 Power simulation logic and code
 Let there be $m$ false negatives (i.e.
-$m$ true non-zero effects) among the 64 studies, $m=1,...,64$.
+$m$ true non-zero effects) among the $k$ studies, $m=1,...,k$.
 Therefore we must draw $m$ p-values from a distribution consistent with the alternative hypothesis.
@@
 We continue drawing $d_j's$ until we obtain $m$ p-values greater than 0.05 (due to the OSC restriction of only considering replicate studies with non-significant results).
-We will draw the remaining $64-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95})$.
+We will draw the remaining $k-m$ p-values from a U[0.05,1] distribution and then calculate Fisher's test statistic $X^2=-2\sum_{j=1}^{k} ln(\frac{p_j-0.05}{0.95})$.
 We run this procedure N times and calculate the simulated power of Fisher's method under these conditions to be $\sum_{q=1}^{N}I_{\{X^2_q > \chi^2_{2k,\alpha}\}}/N,$ where $I$ is the indicator function and $\chi^2_{2k,\alpha}$ is the critical value for Fisher's test with $k$ studies and level $\alpha$.
@@
   #        k; number of studies
-  # RETURNS: power of Fisher's test to reject
+  # RETURNS: 5 column dataframe of results [Power N k M delta]
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
@@
   T<-c() #empty list to store Fisher's test statistic 
-  Power<-matrix() #empty matrix to store results
+  power<-matrix() #empty matrix to store results
 
   for(l in 1:length(M)){
-
     for (i in 1:N){
-
-      #print(i) # can uncomment to show progress for lengthy simulations
-
       p0<-runif(k - M[l], 0.05, 1) #draws p-values for the true null effects
-
       p1<-c() #create list to store p-values drawn for non-null effects
-
       for (j in 1:M[l]){ 
-
         p1[j]<-0
-
         while (p1[j] <= 0.05) { #throw away p-values<=0.05
@@
         }
-
-        #print(n[j]) # can uncomment to show progress for lengthy simulations 
+      }
+      #test statistic for Fisher's method, with transformation for truncating p-values
+      T[i]<--2*sum(log((p0-0.05)/0.95))-2*sum(log((p1-0.05)/0.95)) 
     }
 
-      #test statistic for Fisher's method, with transformation for truncating p-values
-      T[i]<--2*sum(log((p0-0.05)/0.95))-2*sum(log((p1-0.05)/0.95))
-
+    power[l]<-sum(T > qchisq(0.95, 2*k) )/N
+    print(power[l])
+    if(power[l]>0.99) break
   }
-
-    Power[l]<-sum(T > qchisq(0.95, 2*k) )/N
-
-    }
-
-    return(Power)
+  data <- as.data.frame(cbind(power = power, 
+                              n = n[1:length(power)], 
+                              k = rep(k,length(power)), 
+                              M = M[1:length(power)], 
+                              delta = rep(delta,length(power))))
+  return(data)
 }
@@
 
-##Appendix B: Fisher's method asymptotic results
+#Appendix B: Fisher's method asymptotic results
 
@@
 As a consequence of this, we rely on long known results from asymptotic theory.
-Lambert and Hall have shown that, given the test statistic is asymptotically normal, the one sided P-value follows a lognormal distribution with mean $-nc(\theta)$ and variance $n\tau^2(\theta)$ (1982).
+Lambert and Hall have shown that, given the test statistic is asymptotically normal, the one sided P-value follows a lognormal distribution with mean $-nc(\theta)$ and variance $n\tau^2(\theta)$, where $n$ is the within group sample size (1982).
 The parameter $c(\theta)$ is defined as half the Bahadur slope, given by $-\frac{1}{n}lim_{n\to\infty}logP_n=c(\theta)$, and is the exponential rate at which the significance level converges to zero under the alternative hypothesis.
-In addition, observe that the variance of the standardized P-value is $\frac {\tau^2(\theta)}{n}$.
+In addition, observe that the variance of the standardized P-value is $\frac {\tau^2(\theta)}{2n}$.
 We can approximate the two sided P-value by doubling the one sided one, which implies that $P_n\sim AlogN\lbrack-\frac{1}{2}nc(\theta), \frac {1}{4}n\tau^2(\theta)\rbrack$, and thus $logP_n\sim AN\lbrack-\frac{1}{2} nc(\theta), \frac{1}{4}n\tau^2(\theta)\rbrack$.
 Multiplying $logP_n$ by $-2$, we obtain $-2logP_n\sim AN\lbrack nc(\theta), n\tau^2(\theta)\rbrack$.
-Note that $-2logP_{n_j}$, $j=1,...,k$, are asymptotically independent random variables following identical distributions, therefore under $H_a$ $$X^2_n=-2\sum_{j=1}^{k} logP_{n_j} \sim AN(\sum_{j=1}^{k}n_j c_j(\theta), \sum_{j=1}^{k}n_j \tau^2_j(\theta))$$.
+Note that $-2logP_{n_j}$, $j=1,...,k$, are independent random variables asymptotically following Normal distributions, therefore under $H_a$ $$X^2_n=-2\sum_{j=1}^{k} logP_{n_j} \sim AN(\sum_{j=1}^{k}n_j c_j(\theta), \sum_{j=1}^{k}n_j \tau^2_j(\theta))$$.
 
-Examining the behavior of $X^2_n$ using simulated data, we find that the normal approximation is valid if the within study sample size $n_j$ is at least 1500, 250, and 100 for $\theta$ equal to 0.2, 0.5, and 0.8, respectively.
-Note that $\theta$ in this case can be interpreted as Cohen's d because $\sigma$ is assumed to be 1.
-When the within study sample sizes are smaller than indicated previously, the distribution of the test statistic is skewed to the right, and therefore the normal approximation is not valid.
+Now consider the two sample shift problem as an example.
+Let $Y_{11},Y_{21},...,Y_{n_11}$ denote a sample of i.i.d.
+observations from a normal distribution with mean $\mu$ and standard deviation 1.
+Let $Y_{12},Y_{22},...,Y_{n_22}$ denote a second sample, independent of the first, with i.i.d.
+observations from a normal $(\mu+\theta, 1)$.
+For simplicity, suppose that $n_1=n_2=n$.
+We test $H_0:\theta=0$ versus $H_a:\theta\neq0$ using the test statistic $\frac{\sqrt{n}(\overline{Y_2}-\overline{Y_1})}{\sqrt{2}}$.
+According to Lambert and Hall, $c(\theta)=\frac{1}{2}\lambda\overline{\lambda}\theta^2$ and $\tau^2(\theta)=\lambda\overline{\lambda}\theta^2$, where $\lambda$ denotes the fractional sample size, and $\overline{\lambda}=1-\lambda$.
+Since we are assuming $n_1=n_2$, both $\lambda$ and $\overline{\lambda}$ are $\frac{1}{2}$.
+Therefore, in the two-sample shift problem we have $$-2logP_{n} \sim AN(\frac{1}{4}n\theta^2, \frac{1}{2}n\theta^2).$$ Note that $\theta$ in this case can be interpreted as Cohen's d because $\sigma$ is assumed to be 1.
 
-```{r,echo=FALSE, message=FALSE, fig.show='hold', fig.align='center'}
+Examining the behavior of $-2log(p)$ using simulated data, we find that the normal approximation is only valid for impractical within study sample sizes.
+The following figure shows histograms of simulated data and the corresponding normal approximations for various values of $\theta$ and $n$.
+Since the normal approximation is a function of $n$ and $\theta$, it can also be considered a function of the power of a single test to detect the specified $\theta$.
+Values of $n$ were chosen to achieve power of 0.80, 0.90, and 0.9999 for each $\theta$.
+For example, when $\theta=0.2$, a within-group sample size of $n=394$ is needed to achieve 80% power, and the first plot in the grid shows that the normal appoximation is poor in this scenario.
+Note that in most cases, the lower tail of the normal approximation includes negative values, whereas we know the true values of $-2log(p)$ can never be negative.
+Unless the within study sample sizes are very large, the distribution of $-2log(p)$ is skewed to the right, and therefore the normal approximation is not valid.
+Even in scenarios with sample sizes large enough the achieve 99.99% power (the third column in the figure), the normal distribution is only a rough approximation.
 
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
+diane<-function(delta,n){
   p1<-c()
-dist_of_X2<-function(delta,n){
+  z <- c()
   for (j in 1:100000){ 
-    p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n + delta^2 / (4*n))))/sqrt(2/n + delta^2 / (4*n))))
+    #p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n + delta^2 / (4*n))))/sqrt(2/n + delta^2 / (4*n))))
+    #p1[j]<-2*(1 - pnorm(abs(rnorm(1, 1/2*2*n*1/4*delta^2, sqrt(2*n*1/4*delta^2)))/sqrt(2*n*1/4*delta^2)))
+    z[j] <- rnorm(1,delta, sqrt(2/n + delta^2/(4*n)))
+    p1[j] <- (1-pnorm(z[j]/sqrt(2/n + delta^2/(4*n))))
+    #p1[j] <- 2*(1-pnorm(abs(z[j])/sqrt(2/n + delta^2/(4*n)))) #two-sided
   }
-  X2<--2*log(p1)
+  D<--2*log(p1)
 }
 
-d1 <- as.data.frame(dist_of_X2(0.2,700))
+library(grid)
+library(gridExtra)
+
+plot <- function(delta, n){
+  d1 <- as.data.frame(diane(delta,n))
   colnames(d1)[1]<-"V1"
   ggplot(d1, aes(x = V1)) +
-  geom_density()
+    theme(text =element_text(size=10), plot.title = element_text(size = 11))+
+    scale_x_continuous(name = "-2log(p)", limits = c(-20,80), breaks = seq(-20,100,20)) +
+    geom_histogram(aes(y = ..density..) , bins = 100) +
+    stat_function(fun = dnorm, args = list(mean = 2*n*1/4*delta^2, sd = sqrt(2*n*delta^2))) +
+    #stat_function(fun = dnorm, args = list(mean = 1/2*2*n*1/4*delta^2, sd = sqrt(2*n*1/4*delta^2))) + #two-sided
+  labs(title = bquote(theta == .(delta) ~ ", n" == .(n)))
+}
 
-library(latex2exp)
-par(mfrow=c(3,2))
-
-hist(dist_of_X2(0.2,700), main=TeX('n=700, $\\theta$=0.2'), xlab=expression(-2*log(p_i)))
-x<-seq(0,60,0.1)
-curve(dnorm(x, mean = 0.5*700*2*1/4*0.2^2, sd = sqrt(2*700*1/4*0.2^2)), add = TRUE)
-hist(dist_of_X2(0.2,1500), main=TeX('n=1500, $\\theta$=0.2'), xlab=expression(-2*log(p_i)))
-
-hist(dist_of_X2(0.5,100), main=TeX('n=100, $\\theta$=0.5'),  xlab=expression(-2*log(p_i)))
-hist(dist_of_X2(0.5,250), main=TeX('n=250, $\\theta$=0.5'),  xlab=expression(-2*log(p_i)))
-
-hist(dist_of_X2(0.8,50), main=TeX('n=50, $\\theta$=0.8'),  xlab=expression(-2*log(p_i)))
-hist(dist_of_X2(0.8,100), main=TeX('n=100, $\\theta$=0.8'),  xlab=expression(-2*log(p_i)))
+quartz()
+p1 <- plot(0.2,394)
+p2 <- plot(0.2,528)
+p3 <- plot(0.2,1500)
+p4 <- plot(0.5,65)
+p5 <- plot(0.5,87)
+p6 <- plot(0.5,266)
+p7 <- plot(0.8,26)
+p8 <- plot(0.8,35)
+p9 <- plot(0.8,109)
+grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, nrow = 3, top = textGrob("Normal approximation of -2log(p)"))
 ```
 
-Now consider the two sample shift problem as an example.
-Let $Y_{11},Y_{21},...,Y_{n1}$ denote a sample of i.i.d.
-observations from a normal distribution with mean $\mu$ and standard deviation 1.
-Let $Y_{12},Y_{22},...,Y_{n2}$ denote a second sample, independent of the first, with i.i.d.
-observations from a normal $(\mu+\theta, 1)$.
-For simplicity, suppose that $n_1$ and $n_2$ are equal.
-We test $H_0:\theta=0$ versus $H_a:\theta\neq0$ by using $\frac{\sqrt{n}(\overline{Y_2}-\overline{Y_1})}{\sqrt{2}}$.
-According to Lambert and Hall, $c(\theta)=\frac{1}{2}\lambda\overline{\lambda}\theta^2$ and $\tau^2(\theta)=\lambda\overline{\lambda}\theta^2$, where $\lambda$ denotes the fractional sample size, and $\overline{\lambda}=1-\lambda$.
-Since we are assuming $n_1$ is equal to $n_2$, both $\lambda$ and $\overline{\lambda}$ are $\frac{1}{2}$.
-Then, if all k studies are testing the same hypothesis, it follows that $X^2_n$ is asymptotically normal with grand mean $\sum_{j=1}^{k}\frac{1}{2}n_j \lambda_j \overline{\lambda_j}\theta_j^2$ and grand variance $\sum_{j=1}^{k}n_j \lambda_j \overline{\lambda_j}\theta_j^2$.
-Consequently, the asymptotic power is:
+The asymptotic results can confirm that the power of Fisher's method approaches 1 as the within study sample sizes go to infinity, which is consistent with the power simulations presented in the body of the paper.
+To see this, note that if all $k$ studies are testing the same hypothesis, it follows that $X^2_n$ is asymptotically normal with grand mean $\sum_{j=1}^{k}\frac{1}{2}n_j \lambda_j \overline{\lambda_j}\theta_j^2$ and grand variance $\sum_{j=1}^{k}n_j \lambda_j \overline{\lambda_j}\theta_j^2$.
+Consequently, the asymptotic power of Fisher's test is:
 
@@
 &=1-\Phi\Big(\frac{\chi^2_{2k}}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}-\frac{1}{2}\Big)+\Phi\Big(\frac{-\chi^2_{2k}}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}-\frac{1}{2}\Big)\\
+&\xrightarrow{}1-\Phi(-\frac{1}{2})+\Phi(-\frac{1}{2}) ~~ as ~~ n\rightarrow \infty\\
 &=1
@@
 
-The following plots of the power function reveal that asymptotic power approaches 1 when the total within study sample size is greater than 600, 100, and 50 for a fixed $\theta$ of 0.2, 0.5, and 0.8, respectively.
+Shown for the the case when $k=64$, the following plots of the power function reveal that asymptotic power approaches 1 when the total within study sample sizes are approximately 600, 100, and 40 for fixed $\theta$'s of 0.2, 0.5, and 0.8, respectively.
 
-```{r prt 1, echo=FALSE, message=FALSE, fig.width=3, fig.height=5,fig.show='hold',fig.align='center'}
-power<-function(n,delta){
-  1-pnorm(155.4047, mean=64*(1/2)*n*(1/4)*(delta^2), sqrt(64*n*(1/4)*(delta^2)))+pnorm(-155.4047, mean=64*(1/2)*n*(1/4)*(delta^2), sqrt(64*n*(1/4)*(delta^2)))
+```{r prt 1, echo=FALSE, message=FALSE, warning = FALSE}
+power_asymp<-function(n,delta){
+  delta <- rep(delta,length(n))
+  power <- 1-pnorm(155.4047, mean=64*(1/2)*n*(1/4)*(delta^2), sqrt(64*n*(1/4)*(delta^2)))+pnorm(-155.4047, mean=64*(1/2)*n*(1/4)*(delta^2), sqrt(64*n*(1/4)*(delta^2)))
+  data <- as.data.frame(cbind(n,power, delta))
+  return(data)
 }
-require(latex2exp)
-par(mfrow=c(3,1))
-plot(power(1:700,0.2), type='l', xlab='n', ylab='Power', main=TeX('$\\theta$=0.2'))
-plot(power(1:700,0.5), type='l', xlab='n', ylab='Power', main=TeX('$\\theta$=0.5'))
-plot(power(1:700,0.8), type='l', xlab='n', ylab='Power', main=TeX('$\\theta$=0.8'))
+n <- seq(1,700,1)
+power_0.2 <- power_asymp(n, 0.2)
+power_0.5 <- power_asymp(n, 0.5)
+power_0.8 <- power_asymp(n, 0.8)
+power_all <- as.data.frame(rbind(power_0.2,power_0.5,power_0.8))
+
+ggplot(power_all, aes(x = n, y = power)) +
+ theme(panel.grid.minor = element_blank(),
+        axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) +
+  facet_grid(delta ~ . , labeller = label_bquote(rows = theta == .(delta)))+
+  scale_x_continuous(name = "Within group sample size (n)", limits = c(0,700), breaks = seq(0,700,100)) +
+    scale_y_continuous(name = "Power", limits = c(0,1), breaks = seq(0, 1, 0.2)) +
+  labs(title = "Aymptotic Power of Fisher's Method Based on Normal Approximation") +
+  geom_line()
 ```

### 2018-08-31 Fishers_method/k10 table.R

@@ -126,3 +126,24 @@ k10_n100_d.8$delta <- rep(0.8, dim(k10_n100_d.8)[1])
 
 k10 <- rbind(k10_n25_d.2,k10_n25_d.5,k10_n25_d.8,k10_n50_d.2,k10_n50_d.5,k10_n50_d.8,k10_n75_d.2, k10_n75_d.5, k10_n75_d.8,k10_n100_d.2,k10_n100_d.5,k10_n100_d.8)
 saveRDS(k10,"k10_data.RDS")
+
+#########NEW SIMS TO FILL IN GAPS######
+run <- power_sims(N= 10, M = c(6,7,8,9),delta = 0.2, n = rep(25,10), k =10)
+
+k10_n25_d.2_extra <- power_sims(100000, c(6,7,8,9), 0.2, rep(25,10), 10)
+k10_n25_d.5_extra <- power_sims(100000, c(6,7,8,9), 0.5, rep(25,10), 10)
+k10_n25_d.8_extra <- power_sims(100000, c(6,7,8,9), 0.8, rep(25,10), 10)
+
+k10_n50_d.2_extra <- power_sims(100000, c(6,7,8,9), 0.2, rep(50,10), 10)
+k10_n50_d.5_extra <- power_sims(100000, c(6,7,8,9), 0.5, rep(50,10), 10)
+k10_n50_d.8_extra <- power_sims(100000, c(6,7), 0.8, rep(50,10), 10) #check one at a time, stop once gets to .99 power
+
+k10_n75_d.2_extra <- power_sims(100000, c(6,7,8,9), 0.2, rep(75,10), 10)
+k10_n75_d.5_extra <- power_sims(100000, c(6,7,8,9), 0.5, rep(75,10), 10)
+k10_n75_d.8_extra <- power_sims(100000, 6, 0.8, rep(75,10), 10) #check one at a time, stop once gets to .99 power
+
+k10_n100_d.2_extra <- power_sims(100000, c(6,7,8,9), 0.2, rep(100,10), 10)
+k10_n100_d.5_extra <- power_sims(100000, c(6,7), 0.5, rep(100,10), 10) #check
+
[…]

### 2018-08-31 Fishers_method/k100 table.R

@@ -149,6 +149,33 @@ ggplot(k100, aes(x = M, y = power, color = factor(delta))) +
   geom_line()
 
 
+#####NEW SIMS TO FILL IN GAPS#####
+k100_n25_d.2_extra <- power_sims(100000, c(20,30,40,60,70,80,90), 0.2, rep(25, 100), 100)
+k100_n25_d.5_extra <- power_sims(100000, c(20,30,40), 0.5, rep(25, 100), 100)
+k100_n25_d.8_extra <- power_sims(100000, c(20,30,40), 0.8, rep(25, 100), 100)
+
+k100_n50_d.2_extra <- power_sims(100000, c(20,30,40,60,70,80,90), 0.2, rep(50, 100), 100)
+k100_n50_d.5_extra <- power_sims(100000, c(20,30,40), 0.5, rep(50, 100), 100)
+k100_n50_d.8_extra <- power_sims(100000, c(20,30,40), 0.8, rep(50, 100), 100) 
+
+k100_n75_d.2_extra <- power_sims(100000, c(20,30,40,60,70,80,90), 0.2, rep(75, 100), 100)
+k100_n75_d.5_extra <- power_sims(100000, c(20,30,40), 0.5, rep(75, 100), 100)
+k100_n75_d.8_extra <- power_sims(100000, c(20,30,40), 0.8, rep(75, 100), 100) 
+
+k100_n100_d.2_extra <- power_sims(100000, c(20,30,40,60,70,80,90), 0.2, rep(100, 100), 100)
+k100_n100_d.5_extra <- power_sims(100000, c(20,30,40), 0.5, rep(100, 100), 100) 
+k100_n100_d.8_extra <- power_sims(100000, c(20,30,40), 0.8, rep(100, 100), 100)
+
+k100 <- readRDS("./k100_data.RDS")
+k100_list <- list(k100, k100_n25_d.2_extra,k100_n25_d.5_extra, k100_n25_d.8_extra, k100_n50_d.2_extra, k100_n50_d.5_extra, k100_n50_d.8_extra,k100_n75_d.2_extra,k100_n75_d.5_extra,k100_n75_d.8_extra, k100_n100_d.2_extra,k100_n100_d.5_extra,k100_n100_d.8_extra)
+ChangeNames <- function(x) {
+  names(x)[1] <- "power"
+  return(x)
+}
+k100_list <- lapply(k100_list, ChangeNames)
[…]

### 2018-08-31 Fishers_method/k50 table.R

@@ -108,3 +108,31 @@ ggplot(k50, aes(x = M, y = power, color = factor(delta))) +
   labs(title = "Power of Fisher's method for k=50 studies", color = expression(delta))+
   geom_smooth(method = "loess", se = FALSE, size = 0.5)
   #geom_line()
+
+
+####NEW SIMS TO FILL IN GAPS###
+k50_n25_d.2_extra <- power_sims(100000, c(15,20,30,35,40,45), 0.2, rep(25, 50), 50)
+k50_n25_d.5_extra <- power_sims(100000, c(15,20,30,35,40,45), 0.5, rep(25, 50), 50)
+k50_n25_d.8_extra <- power_sims(100000, c(15,20), 0.8, rep(25, 50), 50)
+
+k50_n50_d.2_extra <- power_sims(100000, c(15,20,30,35,40,45), 0.2, rep(50, 50), 50)
+k50_n50_d.5_extra <- power_sims(100000, c(15,20), 0.5, rep(50, 50), 50)
+k50_n50_d.8_extra <- power_sims(100000, c(15,20), 0.8, rep(50, 50), 50) 
+
+k50_n75_d.2_extra <- power_sims(100000, c(15,20,30,35,40,45), 0.2, rep(75, 50), 50)
+k50_n75_d.5_extra <- power_sims(100000, c(15,20), 0.5, rep(75, 50), 50)
+k50_n75_d.8_extra <- power_sims(100000, c(15,20), 0.8, rep(75, 50), 50) 
+
+k50_n100_d.2_extra <- power_sims(100000, c(15,20,30,35,40,45), 0.2, rep(100, 50), 50)
+k50_n100_d.5_extra <- power_sims(100000, c(15,20), 0.5, rep(100, 50), 50) 
+k50_n100_d.8_extra <- power_sims(100000, c(15), 0.8, rep(100, 50), 50) 
+
+k50 <- readRDS("./k50_data.RDS")
+k50_list <- list(k50, k50_n25_d.2_extra,k50_n25_d.5_extra, k50_n25_d.8_extra, k50_n50_d.2_extra, k50_n50_d.5_extra, k50_n50_d.8_extra,k50_n75_d.2_extra,k50_n75_d.5_extra,k50_n75_d.8_extra, k50_n100_d.2_extra,k50_n100_d.5_extra,k50_n100_d.8_extra)
+ChangeNames <- function(x) {
+  names(x)[1] <- "power"
[…]

### 2018-08-31 Fishers_method/power_sims.R

@@ -1,48 +1,51 @@
-power_sims<-function(N,M,delta,n,K){
+power_sims<-function(N,M,delta,n,k){
   #########################################################################################
   # TAKES: N; number of simulations 
   #        M; vector of number of non-null effects
   #        delta; effect size under alternative hypothesis, on scale of cohen's d
   #        n; vector of treatment/control sample size across studies (total sample size/2)
-  #        K; number of studies
-  # RETURNS: power of Fisher's test to reject
+  #        k; number of studies
+  # RETURNS: 5 column dataframe of results [Power N k M delta]
   # Assumes 2-sided p-values, throws away p-values<=0.05 to match OSC methods
   #########################################################################################
   
   T<-c() #empty list to store Fisher's test statistic 
-  Power<-matrix() #empty matrix to store results
-  
-  for(k in 1:length(M)){
+  power<-matrix() #empty matrix to store results
+ 
+  for(l in 1:length(M)){
     
     for (i in 1:N){
       
-      print(i) # can uncomment to show progress for lengthy simulations
-    
-      p0<-runif(K - M[k], 0.05, 1) #draws p-values for the true null effects
+      #print(i) # can uncomment to show progress for lengthy simulations
+      
+      p0<-runif(k - M[l], 0.05, 1) #draws p-values for the true null effects
       
       p1<-c() #create list to store p-values drawn for non-null effects
       
-      for (j in 1:M[k]){ 
+      for (j in 1:M[l]){ 
         
         p1[j]<-0
         
         while (p1[j] <= 0.05) { #throw away p-values<=0.05
-          p1[j]<-2*(1 - pnorm(abs(rnorm(1, delta, sqrt(2/n[j] + delta^2/(4*n[j]))))/sqrt(2/n[j] + delta^2/(4*n[j]))))
+          z[j] <- rnorm(1,delta, sqrt(2/n[j] + delta^2/(4*n[j])))
+          p1[j] <- 2*(1-pnorm(abs(z[j])/sqrt(2/n[j] + delta^2/(4*n[j]))))
         }
         
-        print(n[j]) # can uncomment to show progress for lengthy simulations 
-        }
+        #print(n[j]) # can uncomment to show progress for lengthy simulations 
+      }
       
       #test statistic for Fisher's method, with transformation for truncating p-values
[…]

### 2018-10-15 Fishers_method/Fishers_method_101518.Rmd
updated asymptotic results in appendix

--- 2018-08-31 Fishers_method/Fishers_method_082118.Rmd
@@
 As a consequence of this, we rely on long known results from asymptotic theory.
-Lambert and Hall have shown that, given the test statistic is asymptotically normal, the one sided P-value follows a lognormal distribution with mean $-nc(\theta)$ and variance $n\tau^2(\theta)$, where $n$ is the within group sample size (1982).
+Lambert and Hall have shown that, given the test statistic is asymptotically normal, the one sided P-value follows a lognormal distribution with mean $-nc(\theta)$ and variance $n\tau^2(\theta)$, where $n$ is the within study sample size (1982).
 The parameter $c(\theta)$ is defined as half the Bahadur slope, given by $-\frac{1}{n}lim_{n\to\infty}logP_n=c(\theta)$, and is the exponential rate at which the significance level converges to zero under the alternative hypothesis.
-In addition, observe that the variance of the standardized P-value is $\frac {\tau^2(\theta)}{2n}$.
-We can approximate the two sided P-value by doubling the one sided one, which implies that $P_n\sim AlogN\lbrack-\frac{1}{2}nc(\theta), \frac {1}{4}n\tau^2(\theta)\rbrack$, and thus $logP_n\sim AN\lbrack-\frac{1}{2} nc(\theta), \frac{1}{4}n\tau^2(\theta)\rbrack$.
-Multiplying $logP_n$ by $-2$, we obtain $-2logP_n\sim AN\lbrack nc(\theta), n\tau^2(\theta)\rbrack$.
-Note that $-2logP_{n_j}$, $j=1,...,k$, are independent random variables asymptotically following Normal distributions, therefore under $H_a$ $$X^2_n=-2\sum_{j=1}^{k} logP_{n_j} \sim AN(\sum_{j=1}^{k}n_j c_j(\theta), \sum_{j=1}^{k}n_j \tau^2_j(\theta))$$.
-
+In addition,the variance of the standardized P-value is $\frac {\tau^2(\theta)}{2n}$.Now, given that $P_n$ is asymptotically lognormal, then $logP_n\sim AN(-nc(\theta), n\tau^2(\theta))$, and thus $-2logP_n\sim AN(2nc(\theta), 4n\tau^2(\theta))$.
+Note that $-2logP_{n_j}$, $j=1,...,k$, are asymptotically independent random variables, therefore $$X^2_n=-2\sum_{j=1}^{k} logP_{n_j} \sim AN(2\sum_{j=1}^{k}n_j c_j(\theta), 4\sum_{j=1}^{k}n_j \tau^2_j(\theta))$$ under $P_\theta$.
 Now consider the two sample shift problem as an example.
@@
 Since we are assuming $n_1=n_2$, both $\lambda$ and $\overline{\lambda}$ are $\frac{1}{2}$.
-Therefore, in the two-sample shift problem we have $$-2logP_{n} \sim AN(\frac{1}{4}n\theta^2, \frac{1}{2}n\theta^2).$$ Note that $\theta$ in this case can be interpreted as Cohen's d because $\sigma$ is assumed to be 1.
+Therefore, in the two-sample shift problem we have $$-2logP_{n} \sim AN(\frac{1}{4}n\theta^2, n\theta^2).$$ Note that $\theta$ in this case can be interpreted as Cohen's d because $\sigma$ is assumed to be 1.
 
@@
 The asymptotic results can confirm that the power of Fisher's method approaches 1 as the within study sample sizes go to infinity, which is consistent with the power simulations presented in the body of the paper.
-To see this, note that if all $k$ studies are testing the same hypothesis, it follows that $X^2_n$ is asymptotically normal with grand mean $\sum_{j=1}^{k}\frac{1}{2}n_j \lambda_j \overline{\lambda_j}\theta_j^2$ and grand variance $\sum_{j=1}^{k}n_j \lambda_j \overline{\lambda_j}\theta_j^2$.
+To see this, note that if all $k$ studies are testing the same hypothesis, it follows that $X^2_n$ is asymptotically normal with grand mean $\sum_{j=1}^{k} 4n_j \lambda_j \overline{\lambda_j}\theta_j^2$ and grand variance $\sum_{j=1}^{k}n_j \lambda_j \overline{\lambda_j}\theta_j^2$.
 Consequently, the asymptotic power of Fisher's test is:
@@
 &=1-Pr(|X^2_n| \ < \ \chi^2_{2k} \ | \ H_a \ is \ true)\\
-&=1-\Phi\Big(\frac{\chi^2_{2k}-\sum_{j=1}^{k} \frac{1}{2} n_j\lambda_j \overline{\lambda}_j\theta^2_j}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}\Big)+\Phi\Big(\frac{-\chi^2_{2k}-\sum_{j=1}^{k} \frac{1}{2} n_j\lambda_j \overline{\lambda}_j\theta^2_j}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}\Big)\\
-&=1-\Phi\Big(\frac{\chi^2_{2k}}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}-\frac{1}{2}\Big)+\Phi\Big(\frac{-\chi^2_{2k}}{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}-\frac{1}{2}\Big)\\
-&\xrightarrow{}1-\Phi(-\frac{1}{2})+\Phi(-\frac{1}{2}) ~~ as ~~ n\rightarrow \infty\\
-&=1
+&=1-\Phi\Big(\frac{\chi^2_{2k}-\sum_{j=1}^{k}  n_j\lambda_j \overline{\lambda}_j\theta^2_j}{\sqrt{\sum_{j=1}^{k}4n_j\lambda_j \overline{\lambda}_j\theta^2_j}} \Big)+\Phi\Big(\frac{-\chi^2_{2k}-\sum_{j=1}^{k}  n_j\lambda_j \overline{\lambda}_j\theta^2_j}{\sqrt{\sum_{j=1}^{k}4n_j\lambda_j \overline{\lambda}_j\theta^2_j}}\Big)\\
+&=1-\Phi\Big(\frac{\chi^2_{2k}}{\sqrt{\sum_{j=1}^{k}4n_j\lambda_j \overline{\lambda}_j\theta^2_j}}-\frac{\sqrt{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}}{2}\Big)+\Phi\Big(\frac{-\chi^2_{2k}}{\sqrt{\sum_{j=1}^{k}4n_j\lambda_j \overline{\lambda}_j\theta^2_j}}-\frac{\sqrt{\sum_{j=1}^{k}n_j\lambda_j \overline{\lambda}_j\theta^2_j}}{2}\Big)\\
+&\xrightarrow{}1 ~~ as ~~ n\rightarrow \infty\\
 \end{aligned}.$$

### 2019-01-14 Create Table 1 for REP and RPP data.R
rrita's simulation results (rpe + rpp_delta_0.2)

@@ -4,3 +4,9 @@ subset_RPP_replicate <- full_data_rep_fisher[full_data_rep_fisher$data == "RPP" 
 subset$t <- subset$d / subset$vd
 subset$significance <- ifelse(abs(subset$t)>1.96, 1, 0)
 subset$n <- (8+subset$d^2)/(4*subset$vd)
+
+rpp_data_cleaned <- read_csv("Fishers_method/rpp_data_cleaned.csv")
+rpp$z.test <- rpp$d/sqrt(rpp$vd)
+rpp_fishers <- subset(rpp_data_cleaned, rpp_data_cleaned$replicate == 1 & rpp_data_cleaned$pvalr > 0.05)
+
+rpp_fishers$n_effective <- (8+(rpp_fishers$d)^2)/(4*rpp_fishers$vd)

### 2019-01-14 Create Table 1 for REP and RPP data.R
update code and plot for new rpp and rpe simulations

@@ -10,3 +10,46 @@ rpp$z.test <- rpp$d/sqrt(rpp$vd)
 rpp_fishers <- subset(rpp_data_cleaned, rpp_data_cleaned$replicate == 1 & rpp_data_cleaned$pvalr > 0.05)
 
 rpp_fishers$n_effective <- (8+(rpp_fishers$d)^2)/(4*rpp_fishers$vd)
+
+#01-14-19
+rpp_power_results_delta_0.2 <- readRDS("Fishers_method/rpp_power_results_delta_0.2.RDS")
+rpp_power_results_delta_0.5 <- readRDS("Fishers_method/rpp_power_results_delta_0.5.RDS")
+rpp_delta_0.5_fillin_power100 <- rpp_power_results_delta_0.2 %>% filter(M > max(rpp_power_results_delta_0.5$M)) %>% mutate(power = 1, delta = 0.5)
+rpp_power_results_delta_0.8 <- readRDS("Fishers_method/rpp_power_results_delta_0.8.RDS")
+rpp_delta_0.8_fillin_power100 <- rpp_power_results_delta_0.2 %>% filter(M > max(rpp_power_results_delta_0.8$M)) %>% mutate(power = 1, delta = 0.8)
+rpp_power_results <- rbind(rpp_power_results_delta_0.2, rpp_power_results_delta_0.5, rpp_delta_0.5_fillin_power100, rpp_power_results_delta_0.8, rpp_delta_0.8_fillin_power100)
+
+library(ggplot2)
+n_labels <- c('25' = "n=25", '50' = "n=50", '75' = "n=75", '100'= "n=100")
+k_labels <- c( '10' = "k=10", '50' = "k=50", '100' = "k=100")
+rpp_power_results$prop <- rpp_power_results$M / rpp_power_results$k
+
+library(wesanderson) # color palettes
+zs1 = wes_palette("Zissou1")
+dj1 = wes_palette("Darjeeling1")
+dj2 = wes_palette("Darjeeling2")
+
+# standard theme for plots
+stdtheme = theme(panel.grid.minor = element_blank(),
+                 axis.ticks = element_blank(),
+                 axis.text = element_text(size = 16),
+                 axis.title = element_text(size = 18),
+                 legend.title = element_text(size = 16),
+                 legend.text = element_text(size = 16),
+                 legend.text.align = 0)
+
+plot_rpp <- ggplot(rpp_power_results, aes(x = prop, y = power, color = factor(delta))) +
[…]

### 2019-01-14 Fishers_method/power_sims.R

@@ -1,4 +1,4 @@
-power_sims<-function(N,M,delta,n,k){
+power_sims<-function(N, M, delta, n, k){
   #########################################################################################
   # TAKES: N; number of simulations 
   #        M; vector of number of non-null effects
@@ -44,7 +44,7 @@ power_sims<-function(N,M,delta,n,k){
     if(power[l]>0.99) break
   }
   
-  data <- as.data.frame(cbind(power = power, n = n[1:length(power)], k = rep(k,length(power)), M = M[1:length(power)], delta = rep(delta,length(power))))
+  data <- as.data.frame(cbind(power = power, k = rep(k,length(power)), M = M[1:length(power)], delta = rep(delta,length(power))))
   return(data)
 }
 
@@ -52,6 +52,7 @@ power_sims<-function(N,M,delta,n,k){
 ### RPP DATA SIMULATIONS             ###
 ########################################
 
+library(tidyverse)
 
 #test upper bounds of power (by largest sample size study having false negatives)
 rpp_data_cleaned <- read_csv("Fishers_method/rpp_data_cleaned.csv")

### 2019-01-16 Fishers_method/Fishers method tables and plots.Rmd
updated files for rpp and rpe plots and tables with new simulation data

---
title: "Fisher's method tables and plots"
author: "Katie Fitzgerald"
date: "1/15/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Table: RPE- Power of Fisher's method given large sample sizes for varying $\delta$ and true # of non-null effects (i.e. "Best-case scenario")
```{r, echo=FALSE, warning= FALSE, message = FALSE}
library(knitr)
library(tidyverse)
RPP_RPE_power_results <- readRDS("RPP_RPE_power_results.RDS")
table <- spread(RPP_RPE_power_results, key = delta, value = power)
table_rpp <- table[table$study == "RPP", -c(1,3, 4)]
table_rpe <- table[table$study == "RPE", -c(1,3, 4)]
kable(table_rpe, format="markdown", digits=4, align=c('l','r','r','r'),row.names = FALSE, col.names=c("# of false negatives","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="RPE: Power to reject H0 with large n")
```

Table: RPP- Power of Fisher's method given large sample sizes for varying $\delta$ and true # of non-null effects (i.e. "Best-case scenario")
```{r, echo=FALSE, warning= FALSE, message = FALSE}
kable(table_rpp, format="markdown", digits=4, align=c('l','r','r','r'),row.names = FALSE, col.names=c("# of false negatives","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="RPP: Power to reject H0 with large n")
```

```{r echo = FALSE, warning = FALSE, message = FALSE , fig.cap = "Power of Fisher's method, best-case scenario using RPE and RPP sample sizes \\label{fig:sig:plot_all}"}
library(ggplot2)
library(wesanderson) # color palettes
zs1 = wes_palette("Zissou1")
dj1 = wes_palette("Darjeeling1")
dj2 = wes_palette("Darjeeling2")

# standard theme for plots
stdtheme = theme(panel.grid.minor = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text = element_text(size = 16),
                 axis.title = element_text(size = 18),
                 legend.title = element_text(size = 16),
                 legend.text = element_text(size = 16),
                 legend.text.align = 0)

study_labels <- c( 'RPE' = "RPE, k = 7", 'RPP' = "RPP, k = 64")

plot_all <- ggplot(RPP_RPE_power_results, aes(x = prop, y = power, color = factor(delta)))+
  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
  stdtheme +
  theme(strip.text = element_text(size = 16))+
  scale_x_continuous(name = "Proportion of false negatives") +
  scale_y_continuous(name = "Power")+
  labs(color = expression(delta))+
  geom_line()
plot_all
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Power of Fisher's method, general case \\label{fig:sig:plot_general_case}"}

###CREATE PLOT FOR GENERAL SETTING####
k10 <- readRDS("./k10_data.RDS")
k50 <- readRDS("./k50_data.RDS")
k100 <- readRDS("./k100_data.RDS")

n_labels <- c('25' = "n=25", '50' = "n=50", '75' = "n=75", '100'= "n=100")
k_labels <- c( '10' = "k=10", '50' = "k=50", '100' = "k=100")

k10$prop <- k10$M/k10$k
k50$prop <- k50$M/k50$k
k100$prop <- k100$M/k100$k

all <- rbind(k10,k50,k100)
plot_general_case <- ggplot(all, aes(x = prop, y = power, color = factor(delta))) +
  facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
  stdtheme + 
  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 16)) +
  scale_x_continuous(name = "Proportion of false negatives" ) +
  scale_y_continuous(name = "Power") +
  labs(color = expression(delta))+
  geom_line()

plot_general_case
```

### 2019-01-16 Fishers_method/Fishers_method_101518.Rmd
updated files for rpp and rpe plots and tables with new simulation data

@@
 kable(table,format="markdown",digits=4,align=c('l','r','r','r'),col.names=c("# of non-null effects","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="Power to reject H0 with large n")
+
+table <- spread(all_power_results, key = delta, value = power)
+table_rpp <- table[table$study == "RPP", -c(1,3, 4)]
+table_rpe <- table[table$study == "RPE", -c(1,3, 4)]
+kable(table_rpp, format="markdown", digits=4, align=c('l','r','r','r'),row.names = FALSE, col.names=c("# of false negatives","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="RPP: Power to reject H0 with large n")
+
+kable(table_rpe, format="markdown", digits=4, align=c('l','r','r','r'),row.names = FALSE, col.names=c("# of false negatives","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="RPE: Power to reject H0 with large n")
 ```

### 2019-01-16 Create Table 1 for REP and RPP data.R

@@ -39,17 +39,29 @@ stdtheme = theme(panel.grid.minor = element_blank(),
                  legend.text.align = 0)
 
 plot_rpp <- ggplot(rpp_power_results, aes(x = prop, y = power, color = factor(delta))) +
-  #facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
-  #theme(panel.grid.minor = element_blank(),
-        #axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5)) +
   stdtheme +
-  #scale_color_brewer(palette="Dark2")+
   scale_x_continuous(name = "Proportion of false negatives") +
   scale_y_continuous(name = "Power")+
-  #geom_point(size = 3) +
-  #facet_grid(.~n, labeller = labeller(n = labels)) +
   labs(title = "Power of Fisher's method: RPP", color = expression(delta))+
-  #geom_smooth(method = "loess", se = FALSE, size = 0.5)
   geom_line()
 
 rpe_power_results <- readRDS("Fishers_method/rpe_power_results.RDS")
+rpe_delta_0.8_fillin_power100 <- rpe_power_results %>% filter(M > M[dim(rpe_power_results)[1]] & delta == 0.2) %>% mutate(power = 1, delta = 0.8)
+rpe_power_results <- rbind(rpe_power_results, rpe_delta_0.8_fillin_power100)
+rpe_power_results$prop <- rpe_power_results$M / rpe_power_results$k
+
+rpe_power_results$study <- "RPE"
+rpp_power_results$study <- "RPP"
+all_power_results <- rbind(rpp_power_results, rpe_power_results)
+saveRDS(all_power_results, "RPP_RPE_power_results.RDS")
+
+study_labels <- c( 'RPE' = "RPE, k = 7", 'RPP' = "RPP, k = 64")
+
+plot_all <- ggplot(all_power_results, aes(x = prop, y = power, color = factor(delta))) +
+  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
+  stdtheme +
+  theme(strip.text = element_text(size = 16))+
+  scale_x_continuous(name = "Proportion of false negatives") +
+  scale_y_continuous(name = "Power")+
+  labs(color = expression(delta))+
+  geom_line()

### 2019-01-16 Ttest-pvals/projectseval.R
finished prject have both tables

@@ -8,20 +8,60 @@ fulldat<-readRDS("C:/Users/Mena Whalen/Downloads/full_data_rep_fisher.RDS")
 #                         vd = variance of effect size estimate
 #                        replicated = 0 if authors determined if replication failed, 1 if succeeded
 
+# t test function for the same theta
+ttestsameFunc<-function(theta){
+  
+fulldat %>% 
+  mutate(theta = rep(theta, length(data))) %>%
+  mutate(pval = pval_fun(theta, vd) )%>%
+  group_by(data,experiment) %>%
+  spread(site, pval) %>%
+  mutate(orig = sum(original, na.rm=T), rep=sum(replicate, na.rm=T)) %>%
+  mutate(diff = orig - rep) %>%
+  distinct(data, experiment, orig, rep, diff) %>%
+  ungroup() %>%
+  group_by(data) %>%
+  summarise( n = length(data),
+             mean = mean(diff),
+             sd = sd(diff),
+             power = power.t.test(n=n, delta = mean(diff), sd = sd(diff), sig.level = 0.05, type = "paired", alternative = "two.sided")$power
+             )
+}
 
-ttestF<-function(theta){
+ttestsameFunc(0.2)
+ttestsameFunc(0.5)
+ttestsameFunc(0.8)
+
+
+
+####################################################################################
+# t test function for different thetas
+##bigger effects go into theta1
+
+
+ttestdiffFunc<-function(theta1, theta2){
+  fulldat %>% 
+    mutate(theta = ifelse(site == "original", theta1, theta2)) %>%
+    mutate(pval = pval_fun(theta, vd)) %>%
+    group_by(data,experiment) %>%
+    spread(site, pval) %>%
+    mutate(orig = sum(original, na.rm=T), rep=sum(replicate, na.rm=T)) %>%
+    mutate(diff = orig - rep) %>%
+    distinct(data, experiment, orig, rep, diff) %>%
+    ungroup() %>%
+    group_by(data) %>%
+    summarise( n = length(data),
+               mean = mean(diff),
+               sd = sd(diff),
+               power = 1-power.t.test(n=n, delta = mean(diff), sd = sd(diff), sig.level = 0.05, type = "paired", alternative = "two.sided")$power
+    )
   
[…]

### 2019-01-16 Ttest-pvals/simttest.R
updated fucntion to include changing samples size n

@@ -21,10 +21,11 @@ source("./src/metrics_funs.R")
 ## T-Test Function
 ##-----------------------------------------------##
 
-ttestsim<-function(N, theta1,theta2, theta3, powern1, powern2, powern3){
+ttestsim<-function(N, n, theta1,theta2, theta3, powern1, powern2, powern3){
   
   #############################################
   # TAKES: N; Number of simulations
+  #       n; number of pairs
   #       theta1; theta coulmn 1
   #       theta2; theta coulmn 2
   #       theta3; theta coulmn 3
@@ -38,11 +39,11 @@ ttestsim<-function(N, theta1,theta2, theta3, powern1, powern2, powern3){
   tmatrix<-matrix(data = NA, nrow = N, ncol = 3)
   for (k in 1:N){
     
-    t1<-rep(theta1, 100) # theta values column 1
+    t1<-rep(theta1, n) # theta values column 1
     sim_1<-simulate_studies(t1, 2/(powern1)) # sample size needed for given power depending on theta1 
-    t2<-rep(theta2, 100) # theta values column 2
+    t2<-rep(theta2, n) # theta values column 2
     sim_2<-simulate_studies(t2, 2/(powern2)) # samples size needed for given power depending on theta2
-    t3<-rep(theta3,100) # theta values column 3
+    t3<-rep(theta3, n) # theta values column 3
     sim_3<-simulate_studies(t3, 2/(powern3)) # samples size needed for given power depending on theta3
     
     #difference in each pvalue
@@ -52,9 +53,9 @@ ttestsim<-function(N, theta1,theta2, theta3, powern1, powern2, powern3){
     
     # find t statistics
     # single t statistics for given disparity
-    ts1<- mean(diff_1)/(sd(diff_1)/10)
-    ts2<- mean(diff_2)/(sd(diff_2)/10)
-    ts3<- mean(diff_3)/(sd(diff_3)/10)
+    ts1<- mean(diff_1)/(sd(diff_1)/sqrt(n))
+    ts2<- mean(diff_2)/(sd(diff_2)/sqrt(n))
+    ts3<- mean(diff_3)/(sd(diff_3)/sqrt(n))
     
     tmatrix[k,1]<-ts1
     tmatrix[k,2]<-ts2
[…]

### 2019-01-16 Fishers_method/Fishers method tables and plots.Rmd

@@
 library(tidyverse)
-RPP_RPE_power_results <- readRDS("RPP_RPE_power_results.RDS")
+RPP_RPE_power_results <- readRDS("./RPP_RPE_power_results.RDS")
 table <- spread(RPP_RPE_power_results, key = delta, value = power)

### 2019-01-17 Ttest-pvals/projectseval.R
updated to full rpp data

@@ -1,6 +1,9 @@
 source("./src/metrics_funs.R")
 library(tidyverse)
+library(readr)
+rpp<- read_csv("Ttest-pvals/rpp_data_cleaned.csv")
 fulldat<-readRDS("C:/Users/Mena Whalen/Downloads/full_data_rep_fisher.RDS")
+fulldat<- readRDS("C:/Users/Mena/Documents/Replication/replication_metrics/Ttest-pvals/full_data_rep_fisher.RDS")
 #data = paper from which results were pulled
 #experiment = experiment name
 #site = site name ("original" or "replicate")
@@ -9,6 +12,17 @@ fulldat<-readRDS("C:/Users/Mena Whalen/Downloads/full_data_rep_fisher.RDS")
 #                        replicated = 0 if authors determined if replication failed, 1 if succeeded
 
 # t test function for the same theta
+rpp$site<-ifelse(rpp$replicate == 0, "original", "replicate")
+newrpp<-data.frame(cbind(data = rep("RPP",length(rpp$site)),experiment =rpp$experiment, site= rpp$site, d=rpp$d, 
+                         vd=rpp$vd, replicated=rpp$replicated))
+
+fulldat<-fulldat[!(fulldat$data == "RPP"),]
+fulldat<-rbind(fulldat, newrpp)
+fulldat$d<-as.numeric(fulldat$d)
+fulldat$vd<-as.numeric(fulldat$vd)
+fulldat$replicated<-as.integer(fulldat$replicated)
+
+
 ttestsameFunc<-function(theta){
   
 fulldat %>% 

### 2020-05-19 Fishers_method/Figures 3 and 4.Rmd
previously uncommitted changes to Fisher's method

---
title: "Figures 3 and 4"
author: "Katie Fitzgerald"
date: "6/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Figure 3, warning = F, message = F, echo = F}
library(ggplot2)
library(wesanderson) # color palettes
zs1 = wes_palette("Zissou1")
dj1 = wes_palette("Darjeeling1")
dj2 = wes_palette("Darjeeling2")

# standard theme for plots
stdtheme = theme(panel.grid.minor = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text = element_text(size = 16),
                 axis.title = element_text(size = 18),
                 legend.title = element_text(size = 16),
                 legend.text = element_text(size = 16),
                 legend.text.align = 0)

#Figure 3
power_fishers <- readRDS("./power_fishers.RDS")

ggplot(power_fishers, aes(x = prop, y = power, linetype = factor(power_original))) +
  facet_wrap(~k, labeller = labeller(k = r_labels), ncol = 1) +
  stdtheme + 
  theme(axis.text = element_text(size = 16), strip.text = element_text(size = 16)) +
  scale_x_continuous(name = "Proportion of false negatives" ) +
  scale_y_continuous(name = "Power of Fisher's method", breaks = seq(0, 1, .25)) +
  labs(linetype = "Power of \nexperiments")+
  geom_line()
#geom_smooth(se = F, color = "black")

```

```{r Figure 4, warning = F, message = F, echo = F}
#Figure 4
RPP_RPE_power_results <- readRDS("./RPP_RPE_power_results.RDS")
study_labels <- c( 'RPE' = "RPE, r = 7", 'RPP' = "RPP, r = 64")

ggplot(RPP_RPE_power_results, aes(x = prop, y = power, linetype = factor(delta)))+
  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
  stdtheme +
  theme(strip.text = element_text(size = 16))+
  scale_x_continuous(name = "Proportion of false negatives", ) +
  scale_y_continuous(name = "Power of Fisher's method")+
  labs(linetype = expression(delta))+
  geom_line()
#geom_smooth(se = FALSE, color = "black")
```

### 2020-05-19 Fishers_method/Fishers method tables and plots.Rmd
previously uncommitted changes to Fisher's method

@@
 
-study_labels <- c( 'RPE' = "RPE, k = 7", 'RPP' = "RPP, k = 64")
+study_labels <- c( 'RPE' = "RPE, r = 7", 'RPP' = "RPP, r = 64")
 
-plot_all <- ggplot(RPP_RPE_power_results, aes(x = prop, y = power, color = factor(delta)))+
+plot_all <- ggplot(RPP_RPE_power_results, 
+                   aes(x = prop, y = power, linetype = factor(delta)))+
   facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
@@
   scale_x_continuous(name = "Proportion of false negatives") +
-  scale_y_continuous(name = "Power")+
-  labs(color = expression(delta))+
+  scale_y_continuous(name = "Power of Fisher's method")+
+  labs(linetype = expression(delta))+
   geom_line()
 plot_all
+
+plot_worst <- ggplot(RPP_RPE_worst, aes(x = prop, y = power, linetype = factor(delta)))+
+  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
+  stdtheme +
+  theme(strip.text = element_text(size = 16))+
+  scale_x_continuous(name = "Proportion of false negatives") +
+  scale_y_continuous(name = "Power of Fisher's method")+
+  labs(linetype = expression(delta))+
+  geom_line()
+plot_worst
 ```
@@
 all <- rbind(k10,k50,k100)
-plot_general_case <- ggplot(all, aes(x = prop, y = power, color = factor(delta))) +
-  facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
+plot_general_case <- ggplot(all, aes(x = prop, y = power, linetype = factor(delta))) +
+  #facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
   stdtheme + 
@@
   scale_y_continuous(name = "Power") +
-  labs(color = expression(delta))+
+  labs(linetype = expression(delta))+
   geom_line()
@@
 plot_general_case
+
+power_80_k10 <- dplyr::filter(all, k == 10 & ((delta == .8 & n == 25 ) | (delta == 0.5 & (n == 50 | n == 75)) | (delta == 0.2 & n == 100)))
+
 ```

### 2020-05-19 Fishers_method/Create Table 1 for REP and RPP data.R

@@ -1,0 +1,67 @@
+subset <- full_data_rep_fisher[full_data_rep_fisher$data == "RPP" & full_data_rep_fisher$site == "replicate" & full_data_rep_fisher$replicated == 0,]
+subset_RPP <- full_data_rep_fisher[full_data_rep_fisher$data == "RPP",]
+subset_RPP_replicate <- full_data_rep_fisher[full_data_rep_fisher$data == "RPP" & full_data_rep_fisher$site == "replicate" ,]
+subset$t <- subset$d / subset$vd
+subset$significance <- ifelse(abs(subset$t)>1.96, 1, 0)
+subset$n <- (8+subset$d^2)/(4*subset$vd)
+
+rpp_data_cleaned <- read_csv("Fishers_method/rpp_data_cleaned.csv")
+rpp$z.test <- rpp$d/sqrt(rpp$vd)
+rpp_fishers <- subset(rpp_data_cleaned, rpp_data_cleaned$replicate == 1 & rpp_data_cleaned$pvalr > 0.05)
+
+rpp_fishers$n_effective <- (8+(rpp_fishers$d)^2)/(4*rpp_fishers$vd)
+
+#01-14-19
+rpp_power_results_delta_0.2 <- readRDS("Fishers_method/rpp_power_results_delta_0.2.RDS")
+rpp_power_results_delta_0.5 <- readRDS("Fishers_method/rpp_power_results_delta_0.5.RDS")
+rpp_delta_0.5_fillin_power100 <- rpp_power_results_delta_0.2 %>% filter(M > max(rpp_power_results_delta_0.5$M)) %>% mutate(power = 1, delta = 0.5)
+rpp_power_results_delta_0.8 <- readRDS("Fishers_method/rpp_power_results_delta_0.8.RDS")
+rpp_delta_0.8_fillin_power100 <- rpp_power_results_delta_0.2 %>% filter(M > max(rpp_power_results_delta_0.8$M)) %>% mutate(power = 1, delta = 0.8)
+rpp_power_results <- rbind(rpp_power_results_delta_0.2, rpp_power_results_delta_0.5, rpp_delta_0.5_fillin_power100, rpp_power_results_delta_0.8, rpp_delta_0.8_fillin_power100)
+
+library(ggplot2)
+n_labels <- c('25' = "n=25", '50' = "n=50", '75' = "n=75", '100'= "n=100")
+k_labels <- c( '10' = "k=10", '50' = "k=50", '100' = "k=100")
[…]

### 2020-05-19 Fishers_method/Figures 3 and 4.R

@@ -1,0 +1,41 @@
+library(ggplot2)
+library(wesanderson) # color palettes
+zs1 = wes_palette("Zissou1")
+dj1 = wes_palette("Darjeeling1")
+dj2 = wes_palette("Darjeeling2")
+
+# standard theme for plots
+stdtheme = theme(panel.grid.minor = element_blank(),
+                 axis.ticks = element_blank(),
+                 axis.text = element_text(size = 16),
+                 axis.title = element_text(size = 18),
+                 legend.title = element_text(size = 16),
+                 legend.text = element_text(size = 16),
+                 legend.text.align = 0)
+
+#Figure 3
+power_fishers <- readRDS("power_fishers.RDS")
+
+ggplot(power_fishers, aes(x = prop, y = power, linetype = factor(power_original))) +
+  facet_wrap(~k, labeller = labeller(k = r_labels), ncol = 1) +
+  stdtheme + 
+  theme(axis.text = element_text(size = 16), strip.text = element_text(size = 16)) +
+  scale_x_continuous(name = "Proportion of false negatives" ) +
+  scale_y_continuous(name = "Power of Fisher's method", breaks = seq(0, 1, .25)) +
+  labs(linetype = "Power of \nexperiments")+
+  geom_line()
+  #geom_smooth(se = F, color = "black")
+
+#Figure 4
+RPP_RPE_power_results <- readRDS("RPP_RPE_power_results.RDS")
+study_labels <- c( 'RPE' = "RPE, r = 7", 'RPP' = "RPP, r = 64")
+
+ggplot(RPP_RPE_power_results, aes(x = prop, y = power, linetype = factor(delta)))+
+  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
+  stdtheme +
+  theme(strip.text = element_text(size = 16))+
+  scale_x_continuous(name = "Proportion of false negatives" ) +
+  scale_y_continuous(name = "Power of Fisher's method")+
+  labs(linetype = expression(delta))+
+  geom_line()
+  #geom_smooth(se = FALSE, color = "black")

### 2020-05-19 Fishers_method/power_fishers_plot.R

@@ -1,0 +1,88 @@
+##find sample sizes needed to get 40, 60, 80% power
+calculate_v <- function(n, delta){
+  v <- 2/n + delta^2/(4*n)
+  return(v)
+}
+power_fun(.8, calculate_v(27,.8)) #.81 power
+power_fun(.5, calculate_v(65,.5)) #.8 power
+power_fun(.2, calculate_v(400,.2))#.8 power
+power_fun(.8, calculate_v(17,.8))#.61
+power_fun(.5, calculate_v(40,.5))#.6
+power_fun(.2, calculate_v(250,.2))#.6
+power_fun(.8, calculate_v(10,.8))#.4
+power_fun(.5, calculate_v(24,.5))#.4
+power_fun(.2, calculate_v(150,.2))#.4
+
+##k=10##
+M<-c(1,2,3,4,5,10)
+power_by_k <- function(k, M){
+  #M <- c(1, seq(.1*k,k,.1*k))  
+  
+  #40% power
+  n_0.5 <- rep(24, k)
+  power_40_delta_0.5 <- power_sims(100000,M,0.5,n_0.5,k)
+  library(dplyr)
+  power_40 <- mutate(power_40_delta_0.5, prop = M/k)
+  power_40 <- mutate(power_40, power_original = .4)
+  
+  #60% power
+  n_0.5 <- rep(40, k)
+  power_60_delta_0.5 <- power_sims(100000,M,0.5,n_0.5,k)
+  power_60 <- mutate(power_60_delta_0.5, prop = M/k)
+  power_60 <- mutate(power_60, power_original = 0.6)
+  
+  
+  #80% power
+  n_0.5 <- rep(65, k)
+  power_80_delta_0.5 <- power_sims(100000,M,0.5,n_0.5,k)
+  power_80 <- mutate(power_80_delta_0.5, prop = M/k)
+  power_80 <- mutate(power_80, power_original = 0.8)
+  
+  power_fishers <- rbind(power_40, power_60, power_80)
+  return(power_fishers)
+}
+k_10 <- power_by_k(10, seq(1,10,1))
+k_50 <- power_by_k(50, seq(1,50,1))
+k_100 <- power_by_k(100, seq(2,100,2))
+
+power_fishers <- rbind(k_10, k_50, k_100)
+extra <- data.frame(power = rep(1,31), 
+                              k = c(rep(50,13), rep(100, 18)),
+                              M = c(45, 50, 
+                                    30, 35, 40, 45, 50,
[…]

### 2020-05-19 Fishers_method/power_sims.R

@@ -63,6 +63,7 @@ rpp_fishers$n_effective <- (8+(rpp_fishers$d)^2)/(4*rpp_fishers$vd)
 rpp_fishers$n_effective <- ifelse(rpp_fishers$experiment == "Koo", max(rpp_fishers$n_effective,na.rm =TRUE), rpp_fishers$n_effective)
 n<-sort(rpp_fishers$n_effective,decreasing=TRUE, na.last = TRUE)
 
+
 source("./src/metrics_funs.R")
 replicate_power<-data.frame(n,power_delta_0.2=power_fun(0.2,2/n + 0.2^2/(4*n)),power_delta_0.5=power_fun(0.5,2/n + 0.5^2/(4*n)),power_delta_0.8=power_fun(0.8,2/n + 0.8^2/(4*n)))
 #set largest to be the largest sample size among the studies for which there was at most 99.99% power for a given delta
@@ -72,11 +73,27 @@ n_delta_0.8<-c(rep(replicate_power$n[12],11),n[12:64]) #11 largest studies had >
 
 M <- c(1,2,3,4,5,round(64*seq(0.1,1,0.1))) #want simulations for 1-5 studies and at each 10th percentile
 delta_0.2_large_n<-power_sims(100000, M, 0.2, n_delta_0.2, 64) #starts at n=745
-delta_0.5_large_n<-power_sims(100000, M, 0.5, n_delta_0.5, 64) #starts at n=159
+delta_0.5_large_n<-power_sims(10000, M, 0.5, n_delta_0.5, 64) #starts at n=159
 delta_0.8_large_n<-power_sims(100000, M, 0.8, n_delta_0.8, 64) #starts at n=100
 
 rpp_power_results <- rbind(delta_0.2_large_n, delta_0.5_large_n, delta_0.8_large_n)
 saveRDS(rpp_power_results ,"rpp_power_results.RDS")
+
+n_worst <- sort(rpp_fishers$n_effective,decreasing=FALSE, na.last = TRUE)
+source("./src/metrics_funs.R")
+replicate_power<-data.frame(n_worst,power_delta_0.2=power_fun(0.2,2/n_worst + 0.2^2/(4*n_worst)),power_delta_0.5=power_fun(0.5,2/n_worst + 0.5^2/(4*n_worst)),power_delta_0.8=power_fun(0.8,2/n_worst + 0.8^2/(4*n_worst)))
[…]
